\documentclass[oneside,a5paper]{book}
\title{\Huge \textbf{The Danetian Language}}
\author{Javier Castro}
\date{\today}

% DOCUMENT MARGINS %
%\usepackage[top=4em,bottom=5em,left=3em,right=3em]{geometry}
\usepackage{geometry}

% GENERAL SETUP %
\usepackage{multicol} % For multi-column in-text
\usepackage{booktabs} % For tabular horizontal lines
\usepackage{multirow} % For multi-column within the tabular environment
\usepackage{tabularx} % For fixed width tables
\usepackage{enumerate} % For Roman-numeral enumerated lists
\usepackage{microtype} % Supposedly improves spacing
\usepackage{hyperref} % For in-document hyperlinks
\newcommand{\qt}[1]{«#1»} % French-style quotation marks
\usepackage[toc,title,page]{appendix} % For appendices

% Set up the date format %
\usepackage{datetime}
\renewcommand{\dateseparator}{-}
\yyyymmdddate

% For centered multicolumns inside the tabularx environment
\newcommand{\multicolumnxcentered}[3]{
	\multicolumn{#1}{
		>{\centering\hsize=\dimexpr#1\hsize
		+#1\tabcolsep+\arrayrulewidth\relax}#2}{#3}}

% Use a linguistics font under the appropriate environment
\usepackage{fontspec}
\usepackage{xunicode}
\usepackage{xltxtra}
\newfontfamily\ipafont[Scale=1.0]{DoulosSIL}
%\newfontfamily\ipafont[Scale=0.9]{CharisSIL}
%\newcommand{\lan}[1]{{\ipafont \textit{#1}}}
\newcommand{\lan}[1]{\textit{#1}}
\newcommand{\mlan}[1]{\mbox{\lan{#1}}}
\newcommand{\ipa}[1]{{\ipafont #1}}
%\newcommand{\ipa}[1]{#1}

% SET MAIN FONT %
\usepackage{ebgaramond}

% To signal an incomplete section
\newcommand{\INCOMPLETE}[0]{(THIS SECTION IS INCOMPLETE)}

% Define linguistic abbreviations
\usepackage[mcolblock]{leipzig}
\makeglossaries
\usepackage{gb4e} % For typesetting interlinear glosses
\newleipzig{Num}{num}{numeral}
\newleipzig{Int}{int}{interrogative}
\newleipzig{Uni}{uni}{universal}
\newleipzig{Exist}{exist}{existential}
\newleipzig{Impf}{impf}{imperfect}
\newleipzig{Plup}{plup}{pluperfect}
\newleipzig{V}{v}{verb}

\begin{document}
\frontmatter
\maketitle
\tableofcontents
\printglossaries
\mainmatter

\chapter*{Introduction}

The Danetian language is an Indo-European constructed language. I started
this project in 2016, when I decided to take the Proto-Indo-European lexicon
and apply whichever sound changes satisfied me.

At that time, I had been obsessed with Latin starting a few years prior.
However, there were several small details that bothered me about the language.
Therefore, I wanted to have my own personal classical language that I could
have full creative control over.

I decided that I wanted to emulate an Ancient Greek and Latin aesthetic. Thus,
my first iteration of Danetian had a phonology resembling that of Ancient
Greek: long and short vowels, and a three-way contrast between voiced,
unvoiced, and aspirated plosives. But as I kept writing more things in that
version of Danetian, I realized that long vowels were really cumbersome to
pronounce, and aspirated plosives made the language sound ugly to my ears.
Hence, I decided to revisit the sound changes that I originally applied to
Proto-Indo-European.

The result is this current iteration of Danetian, which I am more than
satisfied with. It has a very beautiful European aesthetic both in writing and
in speech. If you'd like to see a modernized version of Proto-Indo-European,
this is the conlang for you!

In this book, I intend to give a comprehensive description of Danetian. I
assume that the reader is as much of a language nerd as I am, or at least that
they are familiar with basic terms in linguistics.

\paragraph{The name of the language}

In Europe, there are many rivers that have similar names: Danube, Don, Dniepr,
Dniestr, etc. In Proto-Indo-European, the term \ipa{*déh₂nu} referred to a river
goddess. Since civilizations throughout history have commonly formed near a
river, I imagined that whoever spoke my conlang would come from a civilization
near a river. I envisioned them worshiping a river goddess called
\textit{Daneta}, so they would call themselves the \textit{Danetians}, or
\lan{danetiskoi}. Therefore, their native tongue is called \textit{Danetian},
or \lan{danetiska}.

\paragraph{Classification}

Danetian forms its own branch in the Indo-European language family. It belongs
to the \textit{satem} side of the family, but it shares some properties
with the \textit{centum} languages.

\INCOMPLETE


\chapter{Phonology}

\section{Consonants}

Danetian has 20 consonant phonemes: six plosives, six fricatives, two
affricates, and six sonorants. These phonemes are organized in the table
below.

\begin{center}
\begin{tabular}{l|cccc}
	\toprule
	~ & Labial & Dental & Palatal & Velar \\
	\midrule
	Plosive & \ipa{p~b} & \ipa{t~d} & ~ & \ipa{k~g} \\
	Affricate & ~ & ~ & \ipa{tʃ~dʒ} & ~ \\
	Fricative & \ipa{f~v} & \ipa{s~z} & \ipa{ʃ} & \ipa{x} \\
	Nasal & \ipa{m} & \ipa{n} & \ipa{ɲ} & ~ \\
	Rhotic & ~ & \ipa{r} & ~ & ~ \\
	Approximant & ~ & \ipa{l} & \ipa{j} & ~ \\
	\bottomrule
\end{tabular}
\end{center}

The main dialect of Danetian has several allophonical variations for
consonants, summarized below. However, the reader is encouraged to speak the
language with their own regional accent, and to pronounce words in whichever
way feels more natural to them. The points below are intended as a guideline,
and are not strictly authoritative.
\begin{itemize}
	\item Whenver \ipa{/s/} appears between vowels or before a voiced
		consonant, it is pronounced as \ipa{[z]}:
		\subitem \lan{nefesos} \ipa{['nɛ.fɛ.zos]} \qt{of the cloud}
		\subitem \lan{fesor} \ipa{[fɛ'zor]} \qt{hand}
		\subitem \lan{tusontis} \ipa{['tu.zon.tis]} \qt{thousand}
		\subitem \lan{esvos} \ipa{['ɛz.vos]} \qt{horse}
		\subitem \lan{donvos} \ipa{['don.vos]} \qt{number}
	\item \ipa{/r/} is pronounced as \ipa{[r]} by default, but as \ipa{[ɾ]}
		between vowels:
		\subitem \lan{pater} \ipa{[pa'tɛr]} \qt{father}
		\subitem \lan{kruros} \ipa{['kru.ɾos]} \qt{bloody}
	\item \ipa{/l/} is pronounced as \ipa{[l]} by default, but as \ipa{[ɫ]}
		at the end of a word, before a consonant, or before \ipa{/o/}
		or \ipa{/u/}:
		\subitem \lan{lana} \ipa{['la.na]} \qt{wool},
		\subitem \lan{luvos} \ipa{['ɫu.vos]} \qt{love},
		\subitem \lan{lokus} \ipa{['ɫo.kus]} \qt{lake},
		\subitem \lan{Michael} \ipa{[mi.xa'ɛɫ]} \qt{Michael},
		\subitem \lan{Vilmos} \ipa{['viɫ.mos]} \qt{William}.
	\item \ipa{/x/} is pronounced as \ipa{[x~h]}.
\end{itemize}

\section{Vowels}

Like many other languages around the world, Danetian has five vowel phonemes.
There is no vowel length distinction.

\begin{center}
\begin{tabular}{l|ccc}
	\toprule
	~ & Front & Central & Back \\
	\midrule
	Close & i & ~ & u \\
	Mid & e & ~ & o \\
	Open & ~ & a & ~ \\
	\bottomrule
\end{tabular}
\end{center}

The vowels \ipa{/a~e~o/} can also be combined with \ipa{/i~u/} to form the
following falling diphthongs:

\begin{center}
\begin{tabular}{c|cc}
	\toprule
	~ & \ipa{i} & \ipa{u} \\
	\midrule
	\ipa{a} & \ipa{ai̯} & \ipa{au̯} \\
	\ipa{e} & \ipa{ei̯} & \ipa{eu̯} \\
	\ipa{o} & \ipa{oi̯} & ~ \\
	\bottomrule
\end{tabular}
\end{center}

In the main dialect of Danetian, vowels undergo a couple of allophonical
variations. Once again, each speaker will have their own unique pronounciation
based on their regional accent, so the following points are merely a guideline.
\begin{itemize}
	\item \ipa{/e/} is pronounced as \ipa{[ɛ]} by default, but as \ipa{[e]}
		whenever the following syllable contains \ipa{/i/} or
		\ipa{/j/}:
		\subitem \lan{preseti} \ipa{['prɛ.ze.ti]} \qt{he asks}
		\subitem \lan{kreva} \ipa{['krɛ.va]} \qt{raw blood}
		\subitem \lan{drageti} \ipa{['dra.ge.ti]} \qt{he drags}
	\item \ipa{/o/} is always pronounced as \ipa{[o]}.
\end{itemize}

%\chapter{Syllables}

%\chapter{Stress}

%\chapter{Phonotactics}



\section{Writing system}

Danetian orthography makes the language look just like any other European
language. The writing system uses the 26 letters of the standard Latin
alphabet, along with the accute accent (\lan{áéíóúý}) and a couple of
digraphs.

Most letters in the Danetian alphabet represent a single sound. The table
below sumarizes the names of every letter, and the phonemic value.

\begin{center}
\begin{tabular}{c|cc}
	\toprule
	letter & name & sound \\
	\midrule
	a & \lan{a} & \ipa{/a/} \\
	b & \lan{be} & \ipa{/b/} \\
	c & \lan{ce} & \ipa{/k/} \\
	d & \lan{de} & \ipa{/d/} \\
	e & \lan{e} & \ipa{/e/} \\
	f & \lan{ef} & \ipa{/f/} \\
	g & \lan{ge} & \ipa{/g/} \\
	h & \lan{hasj} & \ipa{/x/} \\
	i & \lan{i} & \ipa{/i/} \\
	j & \lan{ja} & \ipa{/j/} \\
	k & \lan{ka} & \ipa{/k/} \\
	l & \lan{el} & \ipa{/l/} \\
	m & \lan{em} & \ipa{/m/} \\
	n & \lan{en} & \ipa{/n/} \\
	o & \lan{o} & \ipa{/o/} \\
	p & \lan{pe} & \ipa{/p/} \\
	q & \lan{qu} & \ipa{/k/} \\
	r & \lan{er} & \ipa{/r/} \\
	s & \lan{es} & \ipa{/s/} \\
	t & \lan{te} & \ipa{/t/} \\
	u & \lan{u} & \ipa{/u/} \\
	v & \lan{ve} & \ipa{/v/} \\
	w & \lan{dublevé} & \ipa{/v/} \\
	x & \lan{ix} & \ipa{/s/, /ks/} \\
	y & \lan{igrek} & \ipa{/i/, /j/} \\
	z & \lan{zed} & \ipa{/z/} \\
	\bottomrule
\end{tabular}
\end{center}

The letter \ipa{⟨c⟩} always sounds like a hard \ipa{/k/}, and
is never palatalized as it is in other European languages. The letter
\ipa{⟨x⟩} sounds like /ʃ/ at the beginning of a word (usually Greek or Spanish
loanwords).
The letter \ipa{⟨y⟩} can take on a vocalic or consonantal value, depending on
context, but it's very intuitive to figure out which one is the case.

The native digraphs used are the following:
\begin{itemize}
	\item \ipa{⟨cj⟩}: \ipa{/tʃ/}
	\item \ipa{⟨gj⟩}: \ipa{/dʒ/}
	\item \ipa{⟨nj⟩}: \ipa{/ɲ/}
	\item \ipa{⟨sj⟩}: \ipa{/ʃ/}
\end{itemize}

In addition to these native digraphs, Danetian also has Greco-Roman digraphs:
\begin{itemize}
	\item \ipa{⟨ch⟩}: \ipa{/x/}
	\item \ipa{⟨ph⟩}: \ipa{/f/}
	\item \ipa{⟨th⟩}: \ipa{/t/}
\end{itemize}

Moreover, if any extraneous Greek consonant cluster appears in the
beginning of a word and cannot be pronounced, such as
\ipa{⟨pt⟩}, \ipa{⟨pn⟩}, \ipa{⟨ps⟩}, or any of the like,
then the first consonant is assumed to be silent. Examples:
\begin{itemize}
	\item \lan{psycha} \ipa{/'si.xa/} \qt{soul}
	\item \lan{pneumonia} \ipa{/new.mo'ni.a/} \qt{pneumonia}
	\item \lan{Ptolemaios} \ipa{/to.le'ma.jos/} \qt{Ptolemy}
\end{itemize}



\section{Stress}

Danetian has a stress accent, much like Modern Greek, Spanish, and Russian.
Every word has exactly one stressed syllable, pronounced stronger than the
rest. By the nature of the language, the stressed syllable is restricted to be
either
\begin{itemize}
	\item the last syllable, in which case the word is called oxytone; or
	\item the penultimate syllable, in which case the word is called
		paroxytone; or finally
	\item the antepenultimate syllable, in which case the word is called
		proparoxytone.
\end{itemize}

In most words, the stress falls as further back as it can. For example, in
\lan{respublica}, \qt{republic}, the \lan{-pu-} is stressed; but in
\lan{respublicamos}, \qt{of the republics}, the \lan{-bli-} is stressed,
because that's as far back as the stress can fall.

In writing, the stress is marked by an acute accent, as in
\lan{natúra} \ipa{[na'tu.ɾa]} \qt{nature},
or it can simply be inferred, as in
\lan{anapreseti} \ipa{[a.na'prɛ.ze.ti]} \qt{he prays}. In order to determine
which words should be marked and which ones should be unmarked, we lay down the
following rules:
\begin{enumerate}
	\item In an unmarked word that ends in a vowel, or a vowel plus
		\lan{-m}, \lan{-n}, \lan{-s}, or \lan{-t}, the stress falls
		on the antepenultimate orthographic
		\textit{vowel} (not syllable) if it can,
		or as far back as possible if the word is shorter.
		\subitem \lan{viros} \ipa{/'vi.ros/} \qt{man}
		\subitem \lan{bena} \ipa{/'be.na/} \qt{woman}
		\subitem \lan{donom} \ipa{/'do.nom} \qt{gift}
		\subitem \lan{praman} \ipa{/'pra.man/} \qt{beginning}
		\subitem \lan{vepti} \ipa{/'vep.ti} \qt{he speaks}
		\subitem \lan{ministeriom} \ipa{/mi.nis'te.rjom/} \qt{ministry}
		\subitem \lan{Mercurios} \ipa{/mer'ku.rjos/} \qt{Mercury}
		\subitem \lan{Demetrios} \ipa{/de'me.trjos/} \qt{Dmitri}
	\item Unmarked words that end in \lan{-ia} constitute an exception to
		the previous rule. In these words, the stress falls on the
		\lan{i}.
		\subitem \lan{Anastasia} \ipa{/a.nas.ta'si.a/} \qt{Anastasia}
		\subitem \lan{vidia} \ipa{/vi'di.a/} \qt{wisdom}
		\subitem \lan{rodia} \ipa{/ro'di.a/} \qt{law}
	\item If an unmarked word ends in a consonant other than
		\lan{-m}, \lan{-n}, \lan{-s}, or \lan{-t}, then the stress
		falls on the last syllable.
		\subitem \lan{Daniel} \ipa{/da'njel/} \qt{Daniel}
		\subitem \lan{Gabriel} \ipa{/ga'brjel/} \qt{Gabriel}
		\subitem \lan{Sadok} \ipa{/sa'dok/} \qt{Zadok} (Biblical name)
		\subitem \lan{Achaz} \ipa{/a'xaz/} \qt{Ahaz} (Biblical name)
	\item In an unmarked word, if these previous rules result in the
		a hiatus after the stressed vowel or possibly including it,
		then the stress is moved to the vowel immediately following.
		\subitem \lan{Joanna} \ipa{/jo'a.na/} \qt{John}
		\subitem \lan{Andrea} \ipa{/an'dre.a/} \qt{Andrew}
\end{enumerate}




\chapter{Nominal morphology}

\section{General properties}

Nouns inflect according to the following categories:
\begin{itemize}
	\item \textbf{gender:} masculine, feminine, neuter
	\item \textbf{number:} singular, plural
	\item \textbf{case:}
		nominative,
		accusative,
		genitive,
		ablative,
		dative,
		locative,
		instrumental,
\end{itemize}

\subsection{Gender}

Gender is a grammatical property inherent to each noun. Sometimes it's
correlated with biological sex:
\begin{multicols}{2}
\begin{itemize}
	\item \lan{viros} (\M): man
	\item \lan{bena} (\F): woman
	\item \lan{hertogos} (\M): duke
	\item \lan{hertoga} (\F): duchess
\end{itemize}
\end{multicols}

Most of the time, however, it has nothing to do with sex:
\begin{multicols}{2}
\begin{itemize}
	\item \lan{abol} (\M): apple
	\item \lan{anfros} (\M): rain
	\item \lan{barda} (\F): beard
	\item \lan{nas} (\F): nose
	\item \lan{imperiom} (\N): empire
	\item \lan{dranom} (\N): grain
\end{itemize}
\end{multicols}

\subsection{Number}

Nouns inflect according to number. If a noun refers to a single entity, it's
called \textit{singular}. If it refers to more than one entity, it's called
\textit{plural}.

\subsection{Case}

Nouns can also inflect according to case. Case refers to the grammatical role
that a noun plays in a phrase or sentence. For instance, consider the following
two sentences:

\begin{exe}
	\ex
	\gll
	\lan{Svon}
	\lan{smerdeti}
	\lan{virom}.
	\\
	Dog-{\Nom} bites man-{\Acc}. \\
	\glt The dog bites the man.
\end{exe}

\begin{exe}
	\ex
	\gll
	\lan{Viros}
	\lan{smerdeti}
	\lan{svonam}.
	\\
	Man-{\Nom} bites dog-{\Acc}. \\
	\glt The man bites the dog.
\end{exe}

Clearly, the word for \qt{man} plays a different role in each sentence. In the
first example, the man is receiving the action. In the second one, he is
performing the action. This difference is marked in the different endings:
\lan{viros} when the man performs the action, \lan{virom} when the man receives
the action.

There are seven cases in Danetian. They are listed below, along with their most
common functions.
\begin{itemize}
	\item \textbf{nominative:} Refers to the subject of a sentence. This is
		the standard form of a noun.
		\begin{exe}
			\ex
			\textbf{\lan{Viros}}
			\lan{lufjeti}
			\lan{svonam.}
			\\
			Man-{\Nom} loves dog-{\Acc}.
			\glt The \textbf{man} loves the dog.
		\end{exe}

	\item \textbf{accusative:} Refers to the direct object of a sentence.
		\begin{exe}
			\ex
			\lan{Svon}
			\lan{lufjeti}
			\textbf{\lan{virom}}.
			\\
			Dog-{\Nom} loves man-{\Acc}
			\glt The dog loves the \textbf{man}.
		\end{exe}
	\item \textbf{genitive:} Denotes possession. Equivalent to the
		phrase \qt{of + [something]}.
		\begin{exe}
			\ex
			\lan{Mjanka}
			\textbf{\lan{virosjo}}
			\lan{lufjeti}
			\lan{svonam}.
			\\
			Kitten-{\Nom} man-{\Gen} loves dog-{\Acc}.
			\glt The \textbf{man's} kitten loves the dog.
		\end{exe}
	\item \textbf{ablative:} Can refer to a source of something, or denote
		movement away from something. Equivalent to the phrase
		\qt{from + [something]}.
		\begin{exe}
			\ex
			\lan{Mjanka}
			\lan{peperet}
			\lan{apo}
			\textbf{\lan{virad}}
			\lan{svonei}.
			\\
			Kitten-{\Nom} ran from man-{\Abl} dog-{\Dat}.
			\glt The kitten ran \textbf{from the man} to the dog.
		\end{exe}
	\item \textbf{dative:} Can refer to the indirect object of a sentence,
		or denote movement towards something. Equivalent to the phrase
		\qt{to + [something]}.
		\begin{exe}
			\ex
			\lan{Svon}
			\lan{dedót}
			\lan{fastom}
			\textbf{\lan{viroi}}.
			\\
			Dog-{\Nom} gave stick-{\Acc} man-{\Dat}.
			\glt The dog gave the stick \textbf{to the man}.
		\end{exe}
	\item \textbf{locative:} Refers to a location. Roughly equivalent to the
		phrase \qt{in + [location]}.
		\begin{exe}
			\ex
			\lan{Mjanka}
			\lan{sedjeto}
			\lan{ana}
			\textbf{\lan{viri}}.
			\\
			Kitten-{\Nom} was-sitting on man-{\Loc}.
			\glt The kitten was sitting \textbf{on the man}.
		\end{exe}
	\item \textbf{instrumental:} Can denote a tool or instrument with which
		an action is performed, or can refer to an entity that gives
		company to another. Equivalent to the phrase
		\qt{with + [something]}.
		\begin{exe}
			\ex
			\lan{Svon}
			\lan{igret}
			\lan{som}
			\textbf{\lan{vire}}.
			\\
			Dog-{\Nom} was-playing with man-{\Ins}.
			\glt The dog was playing \textbf{with the man}.
		\end{exe}
\end{itemize}


\section{Strong and weak cases}

The seven cases can be divided into two types: the strong cases ({\Nom} and
\Acc), and the weak cases (\Gen, \Abl, \Dat, \Loc, and \Ins). There are two
main reasons for this.
\begin{enumerate}[i]
	\item For neuter nouns, the strong cases are always identical, while
		for all other nouns, the strong cases are always distinct.
		Thus, it makes sense to group the strong cases together. For
		example,

		\begin{center}
		\begin{tabular}{llll}
			\toprule
			~ & gift (\N) & man (\M) & woman (\F) \\
			\midrule
			\Nom & \lan{donom} & \lan{viros} & \lan{bena} \\
			\Acc & \lan{donom} & \lan{virom} & \lan{benam} \\
			\bottomrule
		\end{tabular}
		\end{center}

	\item There are some nouns that use one stem for the strong cases, and
		a different stem for the weak cases:

		\begin{center}
		\begin{tabular}{llll}
			\toprule
			~ & thought & dog & lake \\
			\midrule
			\Nom & \lan{mentis} & \lan{svon} & \lan{lokus} \\
			\Gen & \lan{manties} & \lan{sunes} & \lan{alkves} \\
			\bottomrule
		\end{tabular}
		\end{center}
\end{enumerate}


\section{Inflection classes}
By examining how each noun inflects, we can come up with several broad
categories:
\begin{multicols}{2}
\begin{itemize}
	\item \textbf{I}: \lan{a}-stem nouns
	\item \textbf{II}: \lan{o}-stem nouns
	\item \textbf{III-i}: \lan{i}-stem nouns
	\item \textbf{III-u}: \lan{u}-stem nouns
	\item \textbf{III-s}: \lan{es}-stem nouns
	\item \textbf{III-n}: \lan{en}-stem nouns
	\item \textbf{III-L}: liquid-stem nouns
	\item \textbf{III-F}: fleeting \lan{s} nouns
	\item \textbf{III-C}: consonant-stem nouns
\end{itemize}
\end{multicols}

Historically, these categories have been grouped into the \textit{first
declension} (\textbf{I}), \textit{second declension} (\textbf{II}), and
\textit{third declension} (\textbf{III}). As can be seen from the list above,
the third declension contains many subcategories.

\subsection{\lan{a}-stem nouns}

Another of the biggest classes of nouns are the \lan{a}-stems. The vast
majority of these are feminine, but there are some masculines here and there.
Some familiar examples include:
\begin{multicols}{2}
\begin{itemize}
	\item \lan{barda} (\F): beard
	\item \lan{bena} (\F): woman
	\item \lan{danva} (\F): tongue; language
	\item \lan{lana} (\F): wool
	\item \lan{luksna} (\F): moon
	\item \lan{redia} (\F): law, right
	\item \lan{pisda} (\F): vulva
	\item \lan{pletia} (\F): square, plaza
	\item \lan{pluksna} (\F): feather
	\item \lan{poina} (\F): punishment
	\item \lan{videva} (\F): widow
\end{itemize}
\end{multicols}

Since all \lan{a}-stem nouns inflect regularly, we can choose one of them to
represent them all.

\begin{center}
\begin{tabular}{l|ll}
	\toprule
	~ & \Sg & \Pl \\
	\midrule
	\Nom & \lan{bena} & \lan{benai} \\
	\Acc & \lan{benam} & \lan{benans} \\
	\Gen & \lan{benas} & \lan{benóm} \\
	\Abl & \lan{benas} & \lan{benamos} \\
	\Dat & \lan{benai} & \lan{benamos} \\
	\Loc & \lan{beni} & \lan{benasu} \\
	\Ins & \lan{bene} & \lan{benafi} \\
	\bottomrule
\end{tabular}
\end{center}

\subsection{\lan{o}-stem nouns}

One of the biggest classes of nouns are the \lan{o}-stem nouns. The vast
majority of nouns in this class are masculine and neuter, but there are a few
feminines here and there. Some familiar examples include:
\begin{multicols}{2}
\begin{itemize}
	\item \lan{adros} (\M): field
	\item \lan{Aigyptos} (\F): Egypt
	\item \lan{artos} (\M): bear
	\item \lan{divos} (\M): god
	\item \lan{donom} (\N): gift
	\item \lan{dranom} (\N): grain
	\item \lan{fumos} (\M): smoke
	\item \lan{jugom} (\N): yoke
	\item \lan{kapros} (\M): goat
	\item \lan{nefros} (\M): kidney
	\item \lan{nisdos} (\M): nest
	\item \lan{orfos} (\M): orphan; slave
	\item \lan{peplos} (\M): wheel, circle
	\item \lan{snoifos} (\M): snow
	\item \lan{svennos} (\M): sleep, slumber
	\item \lan{tauros} (\M): bull
	\item \lan{valpos} (\M): wolf
	\item \lan{verdom} (\N): word
	\item \lan{viros} (\M): man
\end{itemize}
\end{multicols}

Notice that all \lan{o}-stem neuters end in \lan{-om}, and the rest of the
\lan{o}-stems end in \lan{-os}. Since they all inflect regularly, we can
simply choose two nouns (one in \lan{-os} and one in \lan{-om}) to serve as
role models for all the \lan{o}-stems.

\begin{center}
\begin{tabular}{l|ll}
	\toprule
	~ & \Sg & \Pl \\
	\midrule
	\Nom & \lan{viros} & \lan{viroi} \\
	\Acc & \lan{virom} & \lan{virons} \\
	\Gen & \lan{virosjo} & \lan{viróm} \\
	\Abl & \lan{virad} & \lan{viromos} \\
	\Dat & \lan{viroi} & \lan{viromos} \\
	\Loc & \lan{viri} & \lan{virosu} \\
	\Ins & \lan{vire} & \lan{virois} \\
	\bottomrule
\end{tabular}
\quad
\quad
\begin{tabular}{l|ll}
	\toprule
	~ & \Sg & \Pl \\
	\midrule
	\Nom & \lan{donom} & \lan{dona} \\
	\Acc & \lan{donom} & \lan{dona} \\
	\Gen & \lan{donosjo} & \lan{donóm} \\
	\Abl & \lan{donad} & \lan{donomos} \\
	\Dat & \lan{donoi} & \lan{donomos} \\
	\Loc & \lan{doni} & \lan{donosu} \\
	\Ins & \lan{done} & \lan{donois} \\
	\bottomrule
\end{tabular}
\end{center}



\subsection{\lan{i}-stem nouns}

The \lan{i}-stem nouns include both masculine and feminine nouns, and exactly
one neuter noun. Many nouns in this class have one stem for the strong cases
and another stem for the weak cases. Therefore, when listing them out on a
dictionary, it is important to list the {\Nom-\Sg} and the {\Gen-\Pl} to
represent the strong and weak stems, respectively. \footnote{These are called
the \textit{principal parts} of a noun.} If only the {\Nom-\Sg} is listed on a
dictionary entry, then we can assume that the noun doesn't change its stem.
Some familiar examples of \lan{i}-stem nouns include:
\begin{multicols}{2}
\begin{itemize}
	\item \lan{avis}\footnote{Irregular noun.} (\F): bird
	\item \lan{dotis, daties} (\F): act of giving
	\item \lan{hostis} (\M): guest
	\item \lan{mentis, manties} (\F): thought
	\item \lan{mertis, marties} (\F): death
	\item \lan{mori, maries} (\N): sea
	\item \lan{ovis, avies} (\F): sheep
	\item \lan{statis} (\F): standing; position
	\item \lan{vestis, usties} (\F): clothing
\end{itemize}
\end{multicols}

Since the vast majority of \lan{i}-stem nouns are regular \footnote{A noun
that has one stem for the strong cases and another stem for the weak cases is
still considered regular, because its inflection table can be inferred from
just two principal parts: the {\Nom-\Sg} and the {\Gen-\Sg}.}, we can just
choose one word to represent all regular nouns in this class, and then treat
the irregular nouns separately.

\begin{center}
\begin{tabular}{l|ll}
	\toprule
	~ & \Sg & \Pl \\
	\midrule
	\Nom & \lan{hostis} & \lan{hosties} \\
	\Acc & \lan{hostim} & \lan{hostins} \\
	\Gen & \lan{hosties} & \lan{hostióm} \\
	\Abl & \lan{hosties} & \lan{hostimos} \\
	\Dat & \lan{hostiei} & \lan{hostimos} \\
	\Loc & \lan{hosti} & \lan{hostisu} \\
	\Ins & \lan{hostie} & \lan{hostifi} \\
	\bottomrule
\end{tabular}
\end{center}

The only neuter \lan{i}-stem noun is \lan{mori}, and it's completely regular,
albeit with a change of stem.

\begin{center}
\begin{tabular}{l|ll}
	\toprule
	~ & \Sg & \Pl \\
	\midrule
	\Nom & \lan{mori} & \lan{moria} \\
	\Acc & \lan{mori} & \lan{moria} \\
	\Gen & \lan{maries} & \lan{marióm} \\
	\Abl & \lan{maries} & \lan{marimos} \\
	\Dat & \lan{mariei} & \lan{marimos} \\
	\Loc & \lan{mari} & \lan{marisu} \\
	\Ins & \lan{marie} & \lan{marifi} \\
	\bottomrule
\end{tabular}
\end{center}

One notable irregular noun is \lan{avis}. It is considered irregular because it
has three distinct stems, instead of the usual one or two. Moreover, since one
of the stems is purely vocalic, an epenthetic consonant is sometimes inserted
between the stem and the ending. The inflection table is shown below.

\begin{center}
\begin{tabular}{l|ll}
	\toprule
	~ & \Sg & \Pl \\
	\midrule
	\Nom & \lan{avis} & \lan{avies} \\
	\Acc & \lan{avim} & \lan{avins} \\
	\Gen & \lan{ujes} & \lan{ujóm} \\
	\Abl & \lan{ujes} & \lan{vimos} \\
	\Dat & \lan{ujei} & \lan{vimos} \\
	\Loc & \lan{uvi} & \lan{visu} \\
	\Ins & \lan{uje} & \lan{vifi} \\
	\bottomrule
\end{tabular}
\end{center}

The inflection tables for all irregular nouns can be found in
\ref{sec:irregular-nouns}.

\subsection{\lan{u}-stem nouns}

The \lan{u}-stem nouns include nouns from all three genders. Since this class
is a relic from PIE, there are relatively many irregular nouns here. Some
familiar examples include:
\begin{multicols}{2}
\begin{itemize}
	\item \lan{befrus} (\M): beaver
	\item \lan{brus}\footnote{Irregular noun.} (\F): eyebrow
	\item \lan{dakru} (\N): tear (of the eye)
	\item \lan{danu}\footnotemark[\value{footnote}] (\N): river
	\item \lan{denus}\footnotemark[\value{footnote}] (\F): jaw; cheek
	\item \lan{djus}\footnotemark[\value{footnote}] (\M): sky, heaven
	\item \lan{doru}\footnotemark[\value{footnote}] (\N): tree
	\item \lan{dustus} (\M): taste
	\item \lan{lokus, alkues} (\M): pond, lake
	\item \lan{naus, naves} (\F): boat, ship
	\item \lan{smerus, smarues} (\M): grease
	\item \lan{sunus} (\M): son
\end{itemize}
\end{multicols}

We can choose two of these to represent the regular \lan{u}-stem nouns: one for
the masculine and feminine nouns, and one for the neuters.

\begin{center}
\begin{tabular}{l|ll}
	\toprule
	~ & \Sg & \Pl \\
	\midrule
	\Nom & \lan{sunus} & \lan{sunues} \\
	\Acc & \lan{sunum} & \lan{sununs} \\
	\Gen & \lan{sunues} & \lan{sunuóm} \\
	\Abl & \lan{sunues} & \lan{sunumos} \\
	\Dat & \lan{sunuei} & \lan{sunumos} \\
	\Loc & \lan{sunui} & \lan{sunusu} \\
	\Ins & \lan{sunue} & \lan{sunufi} \\
	\bottomrule
\end{tabular}
\quad\quad
\begin{tabular}{l|ll}
	\toprule
	~ & \Sg & \Pl \\
	\midrule
	\Nom & \lan{danu} & \lan{danú} \\
	\Acc & \lan{danu} & \lan{danú} \\
	\Gen & \lan{danues} & \lan{danuóm} \\
	\Abl & \lan{danues} & \lan{danumos} \\
	\Dat & \lan{danuei} & \lan{danumos} \\
	\Loc & \lan{danui} & \lan{danusu} \\
	\Ins & \lan{danue} & \lan{danufi} \\
	\bottomrule
\end{tabular}
\end{center}

All the irregular nouns can be found in the tables in
\ref{sec:irregular-nouns}.

\subsection{\lan{es}-stem nouns}

The class of \lan{es}-stem nouns includes all three genders. All of these nouns
are regular. Some familiar examples are:
\begin{multicols}{2}
\begin{itemize}
	\item \lan{ankos} (\N): curve, bend
	\item \lan{denos} (\N): race, lineage
	\item \lan{jugos} (\N): pair of yoked animals
	\item \lan{menos} (\N): mind
	\item \lan{nefos} (\N): cloud
	\item \lan{odés} (\M): axe, hatchet
	\item \lan{pletos} (\N): breadth
	\item \lan{rebos} (\N): darkness
	\item \lan{slevos} (\N): fame; glory
	\item \lan{tenos} (\N): extension
\end{itemize}
\end{multicols}

Since all \lan{es}-stem nouns are regular, we can simply choose two nouns
to represent them all: one for the masculine and feminine nouns, and one for
the neuters.

\begin{center}
\begin{tabular}{l|ll}
	\toprule
	~ & \Sg & \Pl \\
	\midrule
	\Nom & \lan{odés} & \lan{odesos} \\
	\Acc & \lan{odesam} & \lan{odesans} \\
	\Gen & \lan{odesos} & \lan{odesóm} \\
	\Abl & \lan{odesos} & \lan{odesmos} \\
	\Dat & \lan{odesei} & \lan{odesmos} \\
	\Loc & \lan{odesi} & \lan{odesu} \\
	\Ins & \lan{odese} & \lan{odesfi} \\
	\bottomrule
\end{tabular}
\quad\quad
\begin{tabular}{l|ll}
	\toprule
	~ & \Sg & \Pl \\
	\midrule
	\Nom & \lan{menos} & \lan{menós} \\
	\Acc & \lan{menos} & \lan{menós} \\
	\Gen & \lan{menesos} & \lan{menesóm} \\
	\Abl & \lan{menesos} & \lan{menesmos} \\
	\Dat & \lan{menesei} & \lan{menesmos} \\
	\Loc & \lan{menesi} & \lan{menesu} \\
	\Ins & \lan{menese} & \lan{menesfi} \\
	\bottomrule
\end{tabular}
\end{center}

\subsection{\lan{en}-stem nouns}

The \lan{en}-stem nouns are all neuter, and most of them are regular. Some
common examples are:
\begin{multicols}{2}
\begin{itemize}
	\item \lan{denaman, danamens} (\N): offspring; generation (genealogical)
	\item \lan{noman, nemens} (\N): name
	\item \lan{numan, panvamens} (\N): spirit, soul
	\item \lan{petan, patens} (\N): wing
	\item \lan{seman, samens} (\N): seed
	\item \lan{sovan, savens} (\N): sun
	\item \lan{terman, tarmens} (\N): boundary; end
	\item \lan{vodan, vedens} (\N): water
	\item \lan{vosan}\footnote{Irregular noun.} (\N): spring (season)
\end{itemize}
\end{multicols}

We can choose one of these to represent all \lan{en}-stem nouns, keeping in
mind that there may or may not be a change of stem.

\begin{center}
\begin{tabular}{l|ll}
	\toprule
	~ & \Sg & \Pl \\
	\midrule
	\Nom & \lan{noman} & \lan{nomó} \\
	\Acc & \lan{noman} & \lan{nomó} \\
	\Gen & \lan{nemens} & \lan{nemenóm} \\
	\Abl & \lan{nemens} & \lan{nemenmos} \\
	\Dat & \lan{nemenei} & \lan{nemenmos} \\
	\Loc & \lan{nemeni} & \lan{nemensu} \\
	\Ins & \lan{nemene} & \lan{nemenfi} \\
	\bottomrule
\end{tabular}
\end{center}

\subsection{Liquid-stem nouns}

The liquid consonants are \lan{l} and \lan{r}. If the {\Nom-\Sg} of a noun ends
in a liquid consonant, then it belongs to the class of liquid-stem nouns. This
class includes both masculine and feminine nouns. Some familiar examples are:
\begin{multicols}{2}
\begin{itemize}
	\item \lan{abol}\footnotemark[\value{footnote}] (\M): apple
	\item \lan{bor}\footnotemark[\value{footnote}] (\M): thief
	\item \lan{brater} (\M): brother
	\item \lan{denator, danatres} (\M): parent
	\item \lan{dotor, datres} (\M): giver
	\item \lan{fugater, fugtres} (\F): daughter
	\item \lan{for, fures} (\M): door
	\item \lan{mater} (\F): mother
	\item \lan{pater} (\M): father
	\item \lan{sper, aspres} (\M): sparrow
	\item \lan{ster, astres} (\M): star
	\item \lan{svesor}\footnotemark[\value{footnote}] (\F): sister
	\item \lan{vistor} (\M): witness
\end{itemize}
\end{multicols}

The vowel right before the liquid is usually elided in all its forms except
the {\Nom-\Sg},
like \lan{pater} > \lan{patres}, or \lan{abol} > \lan{ables}. However, if
the {\Nom-\Sg} is a monosyllable, then no elision occurs, e.g.
\lan{bor} > \lan{bores}. With this rule in mind, we can choose one of the
liquid-stem nouns to serve as a role model for the rest.

\begin{center}
\begin{tabular}{l|ll}
	\toprule
	~ & \Sg & \Pl \\
	\midrule
	\Nom & \lan{pater} & \lan{patres} \\
	\Acc & \lan{patram} & \lan{patrans} \\
	\Gen & \lan{patres} & \lan{patróm} \\
	\Abl & \lan{patres} & \lan{patrimos} \\
	\Dat & \lan{patrei} & \lan{patrimos} \\
	\Loc & \lan{patri} & \lan{patrisu} \\
	\Ins & \lan{patre} & \lan{patrifi} \\
	\bottomrule
\end{tabular}
\end{center}

\subsection{Fleeting \lan{s} nouns}

There are some nouns that end in \lan{-s} in the {\Nom-\Sg}, but this \lan{-s}
gets dropped in all other inflections. This is called the fleeting \lan{s}.
Some nouns like this include:
\begin{multicols}{2}
\begin{itemize}
	\item \lan{anats} (\F): duck
	\item \lan{donts, dantes} (\M): tooth
	\item \lan{fans} (\F): goose
	\item \lan{frags} (\M): dregs, sediment (of wine, oil, fat)
	\item \lan{nogs, neges} (\M): nail (of the finger or toe)
	\item \lan{oms, ames} (\M): shoulder
	\item \lan{ops} (\N): eye
	\item \lan{ponts, pantes} (\M): path
	\item \lan{rads} (\F): root
	\item \lan{reds} (\M): king
	\item \lan{sals} (\M): salt
	\item \lan{vents, vantes} (\M): wind
	\item \lan{vops} (\F): voice, speech
\end{itemize}
\end{multicols}

We will choose two nouns to represent the fleeting \lan{s} nouns: one to
represent the masculines and feminines, and the other to represent the neuters.

\begin{center}
\begin{tabular}{l|ll}
	\toprule
	~ & \Sg & \Pl \\
	\midrule
	\Nom & \lan{vops} & \lan{vopes} \\
	\Acc & \lan{vopam} & \lan{vopes} \\
	\Gen & \lan{vopes} & \lan{vopóm} \\
	\Abl & \lan{vopes} & \lan{vopimos} \\
	\Dat & \lan{vopei} & \lan{vopimos} \\
	\Loc & \lan{vopi} & \lan{vopisu} \\
	\Ins & \lan{vope} & \lan{vopifi} \\
	\bottomrule
\end{tabular}
\quad\quad
	\begin{tabular}{l|ll}
	\toprule
	~ & \Sg & \Pl \\
	\midrule
	\Nom & \lan{ops} & \lan{opa} \\
	\Acc & \lan{ops} & \lan{opa} \\
	\Gen & \lan{opes} & \lan{opóm} \\
	\Abl & \lan{opes} & \lan{opimos} \\
	\Dat & \lan{opei} & \lan{opimos} \\
	\Loc & \lan{opi} & \lan{opisu} \\
	\Ins & \lan{ope} & \lan{opifi} \\
	\bottomrule
\end{tabular}
\end{center}

\subsection{Consonant-stem nouns}

If a noun doesn't fall into any of the categories mentioned above, then we can
usually take the stem and just append the inflectional endings directly to it.
Some nouns that don't fall into the previous classes include:
\begin{multicols}{2}
\begin{itemize}
	\item \lan{dom, dems} (\M): house, home
	\item \lan{famón}\footnote{Irregular noun.} (\M): earthling
	\item \lan{fedóm, fames} (\F): earth
	\item \lan{fejóm, fimes} (\F): winter
	\item \lan{frags} (\M): dregs, sediment (of wine, oil, fat)
	\item \lan{gerán}\footnotemark[\value{footnote}] (\M): crane, heron
\end{itemize}
\end{multicols}

The inflection table of consonant-stem nouns is similar to the liquid-stem and
fleeting \lan{s} nouns.

\begin{center}
\begin{tabular}{l|ll}
	\toprule
	~ & \Sg & \Pl \\
	\midrule
	\Nom & \lan{dom} & \lan{domes} \\
	\Acc & \lan{domam} & \lan{domes} \\
	\Gen & \lan{dems} & \lan{demóm} \\
	\Abl & \lan{dems} & \lan{demimos} \\
	\Dat & \lan{demei} & \lan{demimos} \\
	\Loc & \lan{demi} & \lan{demisu} \\
	\Ins & \lan{deme} & \lan{demifi} \\
	\bottomrule
\end{tabular}
\end{center}

\subsection{Irregular nouns}
\label{sec:irregular-nouns}

\INCOMPLETE


\section{Adjectives}
\INCOMPLETE


\chapter{Numerals}

\section{Cardinals}

Cardinals are used for counting things: \textit{one}, \textit{two},
\textit{three}, \textit{four}... They can be simple, like \textit{twenty},
or compound, like \textit{four hundred}. A cardinal number governs a noun or
a noun phrase, forming a numeral phrase as a whole.

\begin{exe}
	\ex
	\gll
	\lan{trifi}
	\lan{berdantifi}
	\lan{virois} \\
	three-{\Ins} tall-{\Ins}-{\Pl} men-{\Ins}-{\Pl} \\
	\glt \qt{with the three tall men}
\end{exe}

\begin{exe}
	\ex
	\gll
	\lan{nevandsont}
	\lan{pempe}
	\lan{thesióm}
	\lan{Martines}
	\lan{Lutheres} \\
	ninety five theses-{\Gen}-{\Pl} Martin-{\Gen} Luther-{\Gen} \\
	\glt \qt{the ninety five theses of Martin Luther}
\end{exe}

\begin{exe}
	\ex
	\gll
	\lan{tusontis}
	\lan{petvóres}
	\lan{satones}
	\lan{octodesam}
	\lan{pavenjóm}
	\lan{dinóm}
	\lan{i}
	\lan{neptóm}
	\lan{Medanosjo}
	\lan{Patriotiskosjo}
	\lan{Urusjes} \\
	thousand four hundred eighteen fiery-{\Gen}-{\Pl}
	days-{\Gen}-{\Pl} and nights-{\Gen}-{\Pl}
	Great-{\Gen} Patriotic-{\Gen} War-{\Gen} \\
	\glt \qt{one thousand four hundred and eighteen fiery days and nights
	of the Great Patriotic War}
\end{exe}

\subsection{One}
\label{sec:one}

The numeral for 1 is \lan{oinos}, and it is the only cardinal that can be
both singular and plural. It inflects just like a regular adjective.

\begin{center}
\begin{tabular}{l|lll}
	\toprule
	\multicolumn{4}{c}{Singular} \\
	\toprule
	~ & \M & \F & \N \\
	\midrule
	\Nom & \lan{oinos} & \lan{oina} & \lan{oinom} \\
	\Acc & \lan{oinom} & \lan{oinam} & \lan{oinom} \\
	\Gen & \lan{oinosjo} & \lan{oinas} & \lan{oinosjo} \\
	\Abl & \lan{oinad} & \lan{oinas} & \lan{oinad} \\
	\Dat & \lan{oinoi} & \lan{oinai} & \lan{oinoi} \\
	\Loc & \lan{oini} & \lan{oini} & \lan{oini} \\
	\Ins & \lan{oine} & \lan{oine} & \lan{oine} \\
	\toprule
	\multicolumn{4}{c}{Plural} \\
	\toprule
	~ & \M & \F & \N \\
	\midrule
	\Nom & \lan{oinoi} & \lan{oinai} & \lan{oina} \\
	\Acc & \lan{oinons} & \lan{oinans} & \lan{oina} \\
	\Gen & \lan{oinóm} & \lan{oinóm} & \lan{oinóm} \\
	\Abl & \lan{oinomos} & \lan{oinamos} & \lan{oinomos} \\
	\Dat & \lan{oinomos} & \lan{oinamos} & \lan{oinomos} \\
	\Loc & \lan{oinosu} & \lan{oinasu} & \lan{oinosu} \\
	\Ins & \lan{oinois} & \lan{oinafi} & \lan{oinois} \\
	\bottomrule
\end{tabular}
\end{center}

The cardinal \lan{oinos}, like adjectives, must always agree with the gender,
number, and case of the noun phrase that it governs.

\subsection{Two, three, and four}
\label{sec:two-three-and-four}

The cardinals for 2, 3, and 4 are \lan{dvo}, \lan{tres}, and
\lan{petvóres}, respectively. They are all similar because they all inflect
according to the three genders, and they're all inherently plural.

\begin{center}
\begin{tabular}{l|lll}
	\toprule
	\multicolumn{4}{c}{Two} \\
	\toprule
	~ & \M & \F & \N \\
	\midrule
	\Nom & \lan{dvo} & \lan{dva} & \lan{dvoi} \\
	\Acc & \lan{dvons} & \lan{dvans} & \lan{dvoi} \\
	\Gen & \lan{dvóm} & \lan{dvóm} & \lan{dvóm} \\
	\Abl & \lan{dvamos} & \lan{dvamos} & \lan{dvamos} \\
	\Dat & \lan{dvamos} & \lan{dvamos} & \lan{dvamos} \\
	\Loc & \lan{dvasu} & \lan{dvasu} & \lan{dvasu} \\
	\Ins & \lan{dvafi} & \lan{dvafi} & \lan{dvafi} \\
	\toprule
	\multicolumn{4}{c}{Three} \\
	\toprule
	~ & \M & \F & \N \\
	\midrule
	\Nom & \lan{tres} & \lan{tistres} & \lan{tri} \\
	\Acc & \lan{trins} & \lan{tistrans} & \lan{tri} \\
	\Gen & \lan{trióm} & \lan{tistróm} & \lan{trióm} \\
	\Abl & \lan{trimos} & \lan{tisarmos} & \lan{trimos} \\
	\Dat & \lan{trimos} & \lan{tisarmos} & \lan{trimos} \\
	\Loc & \lan{trisu} & \lan{tisarsu} & \lan{trisu} \\
	\Ins & \lan{trifi} & \lan{tisarfi} & \lan{trifi} \\
	\toprule
	\multicolumn{4}{c}{Four} \\
	\toprule
	~ & \M & \F & \N \\
	\midrule
	\Nom & \lan{petvóres} & \lan{petestres} & \lan{petvor} \\
	\Acc & \lan{petvórans} & \lan{petestrans} & \lan{petvor} \\
	\Gen & \lan{peturóm} & \lan{petestróm} & \lan{peturóm} \\
	\Abl & \lan{petvarmos} & \lan{petesarmos} & \lan{petvarmos} \\
	\Dat & \lan{petvarmos} & \lan{petesarmos} & \lan{petvarmos} \\
	\Loc & \lan{petvarsu} & \lan{petesarsu} & \lan{petvarsu} \\
	\Ins & \lan{petvarfi} & \lan{petesarfi} & \lan{petvarfi} \\
	\bottomrule
\end{tabular}
\end{center}

Since these three cardinals are inherently plural, the noun phrase that they
govern must also be plural. Moreover, these three cardinals must agree with
the case of their noun phrase.

\subsection{Five to nineteen}
\label{sec:five-to-nineteen}

The cardinals 5-19 inflect according to case, but not gender.
Moreover, they're all inherently plural, so the noun phrase that they govern
must also be plural. However, they may not necessarily agree with the case of
the noun phrase. Case agreement follows a simple set of rules, discussed in
section \ref{sec:case-agreement-for-simple-cardinals}.

\begin{center}
\footnotesize
\begin{tabular}{l|llllll}
	\toprule
	~ & five & six & seven & eight & nine & ten \\
	\midrule

	\Nom & \lan{pempe} & \lan{sveks} & \lan{septam}
	& \lan{okto} & \lan{nevan} & \lan{desam} \\

	\Acc & \lan{pempans} & \lan{svekans} & \lan{septans}
	& \lan{oktons} & \lan{nevans} & \lan{desans} \\

	\Gen & \lan{pempóm} & \lan{svekóm} & \lan{septóm}
	& \lan{októm} & \lan{nevóm} & \lan{desóm} \\

	\Abl & \lan{pempamos} & \lan{svekamos} & \lan{septamos}
	& \lan{oktomos} & \lan{nevamos} & \lan{desamos} \\

	\Dat & \lan{pempamos} & \lan{svekamos} & \lan{septamos}
	& \lan{oktomos} & \lan{nevamos} & \lan{desamos} \\

	\Loc & \lan{pempasu} & \lan{svekasu} & \lan{septasu}
	& \lan{oktosu} & \lan{nevasu} & \lan{desasu} \\

	\Ins & \lan{pempafi} & \lan{svekafi} & \lan{septafi}
	& \lan{oktofi} & \lan{nevafi} & \lan{desafi} \\

	\bottomrule
\end{tabular}
\end{center}

The cardinals 11-19 inflect just like \textit{desam}, so it is not
necessary to show their inflection tables.
\begin{multicols}{2}
\begin{itemize}
	\item \lan{oinodesam} (\Num): eleven
	\item \lan{dvoidesam} (\Num): twelve
	\item \lan{tridesam} (\Num): thirteen
	\item \lan{petvordesam} (\Num): fourteen
	\item \lan{pempedesam} (\Num): fifteen
	\item \lan{svesdesam} (\Num): sixteen
	\item \lan{septandesam} (\Num): seventeen
	\item \lan{oktodesam} (\Num): eighteen
	\item \lan{nevandesam} (\Num): nineteen
\end{itemize}
\end{multicols}

\subsection{Simple cardinals beyond twenty}
\label{sec:simple-cardinals-beyond-twenty}

Every simple cardinal 20 and above inflects like any Danetian noun. They're
all inherently plural, so the noun phrase that they govern must also be in the
plural. However, they may not necessarily agree in case; the case agreement
rules are discussed in section \ref{sec:case-agreement-for-simple-cardinals}.
\begin{multicols}{2}
\begin{itemize}
	\item \lan{vidsant} (\Num): twenty
	\item \lan{tridsont} (\Num): thirty
	\item \lan{petvardsont} (\Num): forty
	\item \lan{pempedsont} (\Num): fifty
	\item \lan{svedsont} (\Num): sixty
	\item \lan{septandsont} (\Num): seventy
	\item \lan{oktodsont} (\Num): eighty
	\item \lan{nevandsont} (\Num): ninety
	\item \lan{satón} (\Num): hundred
	\item \lan{tusontis} (\Num): thousand
\end{itemize}
\end{multicols}

The simple cardinals between 20 and 90 inflect like a consonantal-stem
noun, with the caveat that they're always in the singular.
The numerals \lan{satón} and \lan{tusontis} inflect like a consonantal-stem and
\lan{i}-stem noun, respectively, and they can be either singular or plural.

\section{Case agreement for simple cardinals}
\label{sec:case-agreement-for-simple-cardinals}

As we saw in sections \ref{sec:one} and \ref{sec:two-three-and-four}, the
cardinals 1-4 must agree with the gender, number, and case of the noun phrase
that they govern. Of course, the numerals 2-4 are inherently plural, so the
noun phrase must be cast into the plural.

However, the simple cardinals greater than or equal to 5 follow a slightly
more complicated set of rules:
\begin{enumerate}[i]
	\item The case of the cardinal determines the case of the entire
		numeral phrase.
	\item If we wish to inflect the numeral phrase into one of the
		strong cases ({\Nom} and {\Acc}), then the noun phrase
		must be inflected in the genitive plural.

		\begin{exe}
			\ex
			\gll
			\lan{pempe}
			\lan{viróm}
			\lan{i}
			\lan{benóm} \\
			five-{\Nom} men-{\Gen}-{\Pl} and women-{\Gen}-{\Pl} \\
			\glt \qt{five men and women}
		\end{exe}

		\begin{exe}
			\ex
			\gll
			\lan{tusontim}
			\lan{dranóm}
			\lan{sales} \\
			thousand-{\Acc} grains-{\Gen}-{\Pl}
			of-salt-{\Gen}-{\Sg} \\
			\glt \qt{one thousand grains of salt}
		\end{exe}

	\item If we wish to inflect the numeral phrase into one of the weak
		cases ({\Gen}, {\Abl}, {\Dat}, {\Loc}, {\Ins}), then the noun
		phrase also agrees with that case.

		\begin{exe}
			\ex
			\gll
			\lan{som}
			\lan{petvardsonte}
			\lan{virois} \\
			with forty-{\Ins} men-{\Ins}-{\Pl} \\
			\glt with the forty men
		\end{exe}

		\begin{exe}
			\ex
			\gll
			\lan{apo}
			\lan{oktomos}
			\lan{stadimos} \\
			from eight-{\Abl} locations-{\Abl}-{\Pl} \\
			\glt from eight locations
		\end{exe}
\end{enumerate}

A brief table should illustrate these simple rules.

\begin{center}
\begin{tabular}{l|lll}
	\toprule
	~ & one man & two men & five men \\
	\midrule
	\Nom
	& \lan{oinos viros} & \lan{dvo viroi} & \lan{pempe viróm} \\
	\Acc
	& \lan{oinom virom} & \lan{dvons virons} & \lan{pempans viróm} \\
	\Gen
	& \lan{oinosjo virosjo} & \lan{dvóm viróm} & \lan{pempóm viróm} \\
	\Abl
	& \lan{oinad virad} & \lan{dvamos viromos} & \lan{pempamos viromos} \\
	\Dat
	& \lan{oinoi viroi} & \lan{dvamos viromos} & \lan{pempamos viromos} \\
	\Loc
	& \lan{oini viri} & \lan{dvasu virosu} & \lan{pempasu virosu} \\
	\Ins
	& \lan{oine vire} & \lan{dvafi virois} & \lan{pempafi virois} \\
	\bottomrule
\end{tabular}
\end{center}

\section{Compound numerals}

Several simple numerals can be arranged to form a compound numeral. In English,
an example of this would be \textit{sixty nine}, or \textit{four hundred and
twenty}. We can also do this in Danetian:
\begin{multicols}{2}
\begin{itemize}
	\item 21: \lan{vidsant oinos}
	\item 42: \lan{petvardsont dvo}
	\item 69: \lan{svedsont nevan}
	\item 96: \lan{nevandsont sveks}
	\item 101: \lan{satón oinos}
	\item 144: \lan{satón petvardsont petvóres}
	\item 200: \lan{dvo satones}
	\item 210: \lan{dvo satones desam}
	\item 365: \lan{tres satones svedsont pempe}
	\item 420: \lan{petvóres satones vidsant}
	\item 500: \lan{pempe satonóm}
	\item 507: \lan{pempe satonóm septam}
	\item 1001: \lan{tusontis oinos}
	\item 1200: \lan{tusontis dvo satones}
	\item 1418: \lan{tusontis petvóres satones octodesam}
	\item 1999: \lan{tusontis nevan satones nevandsont nevan}
	\item 2000: \lan{dvo tusonties}
	\item 2021: \lan{dvo tusonties vidsant oinos}
	\item 3000: \lan{tres tusonties}
	\item 5000: \lan{pempe tusontióm}
	\item 9999: \lan{nevan tusontióm nevan satonóm nevandsont nevan}
\end{itemize}
\end{multicols}

Notice that a numeral can govern another numeral, as in \textit{two thousand}.
In this case, the subordinate numeral follows the case agreement rules
discussed in section \ref{sec:case-agreement-for-simple-cardinals}. The
gender of this subordinate numeral is assumed to be masculine.

\section{Ordinals}

Below is a list of the simple ordinals, comprised of just one word.
\begin{multicols}{2}
\begin{itemize}
	\item \lan{pramos} (\Adj): first
	\item \lan{viteros} (\Adj): second
	\item \lan{tritos} (\Adj): third
	\item \lan{petvártos} (\Adj): fourth
	\item \lan{pemptos} (\Adj): fifth
	\item \lan{svektos} (\Adj): sixth
	\item \lan{septamos} (\Adj): seventh
	\item \lan{oktamos} (\Adj): eighth
	\item \lan{nevanos} (\Adj): ninth
	\item \lan{desamos} (\Adj): tenth
	\item \lan{oinodesamos} (\Adj): eleventh
	\item \lan{dvoidesamos} (\Adj): twelfth
	\item \lan{tridesamos} (\Adj): thirteenth
	\item \lan{petvardesamos} (\Adj): fourteenth
	\item \lan{pempedesamos} (\Adj): fifteenth
	\item \lan{svekdesamos} (\Adj): sixteenth
	\item \lan{septandesamos} (\Adj): seventeenth
	\item \lan{oktodesamos} (\Adj): eighteenth
	\item \lan{nevandesamos} (\Adj): nineteenth
	\item \lan{vidsantos} (\Adj): twentieth
	\item \lan{tridsontos} (\Adj): thirtieth
	\item \lan{petvardsontos} (\Adj): fortieth
	\item \lan{pempedsontos} (\Adj): fiftieth
	\item \lan{svekdsontos} (\Adj): sixtieth
	\item \lan{septandsontos} (\Adj): seventieth
	\item \lan{oktodsontos} (\Adj): eightieth
	\item \lan{nevandsontos} (\Adj): ninetieth
	\item \lan{satenjos} (\Adj): hundredth
	\item \lan{tusontjos} (\Adj): thousandth(\Adj): hundredth
	\item \lan{tusontjos} (\Adj): thousandth
\end{itemize}
\end{multicols}

To form compound numerals, simply take the corresponding cardinal, and replace
the last word with its ordinal form. For example, sixty ninth would be
\lan{svekdsont nevanos}.

\chapter{Pronouns and pro-adverbs}

\section{Personal pronouns}
The personal pronouns are classified according to three persons
(\textit{first}, \textit{second}, and \textit{third}), and two numbers
(\textit{singular} and \textit{plural}). The first and second person pronouns
do not inflect according to gender, but the third person pronouns do. There is
also a reflexive pronoun, which does not inflect by number.

\begin{center}
\begin{tabular}{l|lllll}
	\toprule
	~ & \Fsg & \Ssg & \Fpl & \Spl & \Refl \\
	\midrule

	\Nom & \lan{edo} & \lan{tu}
	& \lan{vi} & \lan{ju} & - \\

	\Acc & \lan{me} & \lan{tve}
	& \lan{nos} & \lan{vos} & \lan{sve} \\

	\Gen & \lan{mene} & \lan{teve}
	& \lan{ánser} & \lan{júser} & \lan{seve} \\

	\Abl & \lan{med} & \lan{tved}
	& \lan{ansmed} & \lan{usmed} & \lan{sved} \\

	\Dat & \lan{medjo} & \lan{tevjo}
	& \lan{ansmi} & \lan{usmi} & \lan{sevjo} \\

	\Loc & \lan{moi} & \lan{toi}
	& \lan{ansmi} & \lan{usmi} & \lan{soi} \\

	\Ins & \lan{moi} & \lan{toi}
	& \lan{ansa} & \lan{usma} & \lan{soi} \\

	\bottomrule
\end{tabular}
\end{center}

The third person pronoun inflects according to the table below.
\begin{center}
\begin{tabular}{l|lll}
	\toprule
	\multicolumn{4}{c}{Singular} \\
	\toprule
	~ & \M & \F & \N \\
	\midrule
	\Nom & \lan{eno} & \lan{ena} & \lan{enod} \\
	\Acc & \lan{enom} & \lan{enam} & \lan{enod} \\
	\Gen & \lan{enosjo} & \lan{enosjas} & \lan{enosjo} \\
	\Abl & \lan{enosmad} & \lan{enosjas} & \lan{enosmad} \\
	\Dat & \lan{enosmoi} & \lan{enosjai} & \lan{enosmoi} \\
	\Loc & \lan{enosmi} & \lan{enosji} & \lan{enosmi} \\
	\Ins & \lan{enosme} & \lan{enosje} & \lan{enosme} \\
	\toprule
	\multicolumn{4}{c}{Plural} \\
	\toprule
	~ & \M & \F & \N \\
	\midrule
	\Nom & \lan{enoi} & \lan{enai} & \lan{ena} \\
	\Acc & \lan{enons} & \lan{enans} & \lan{ena} \\
	\Gen & \lan{enosóm} & \lan{enasóm} & \lan{enosóm} \\
	\Abl & \lan{enoimos} & \lan{enamos} & \lan{enoimos} \\
	\Dat & \lan{enoimos} & \lan{enamos} & \lan{enoimos} \\
	\Loc & \lan{enoisu} & \lan{enasu} & \lan{enoisu} \\
	\Ins & \lan{enois} & \lan{enafi} & \lan{enois} \\
	\bottomrule
\end{tabular}
\end{center}

\section{Possessive pronouns}

The possessive pronouns for first and second person inflect just like regular
adjectives, so we will simply list them without showing their inflection
tables.
\begin{multicols}{2}
\begin{itemize}
	\item \Fsg: \lan{menjos}
	\item \Ssg: \lan{tevjos}
	\item \Fpl: \lan{nosjos}
	\item \Spl: \lan{vosjos}
\end{itemize}
\end{multicols}

However, the third person possessive pronouns are equivalent to the genitive of
the third person personal pronouns. That is,
\lan{enosjo, enosjas, enosjo, enosóm, enasóm, enosóm}. These possessives
do not inflect at all.

\section{Pro-adverbs}

A pro-adverb is a word that substitutes an adverb, such as
\textit{when}, \textit{where}, \textit{somewhere}, \textit{like this}, among
others. In Danetian, pro-adverbs refer to the following categories:
\begin{multicols}{2}
\begin{itemize}
	\item location: where?
	\item source: where from?
	\item destination: where to?
	\item time: when?
	\item manner: how?
	\item amount: how much?
\end{itemize}
\end{multicols}

Moreover, every pro-adverb has its own type, and we can tell what type it is
based on the prefix attached to it:
\begin{multicols}{2}
\begin{itemize}
	\item interrogative: \lan{po-}
	\item relative: \lan{jo-}
	\item proximal: \lan{so-}
	\item distal: \lan{to-}
	\item universal: \lan{viso-}
	\item negative: \lan{nipo-}
	\item existential: \lan{poto-}
\end{itemize}
\end{multicols}

Every possible combination is shown in the following table:
\begin{center}
\footnotesize
\begin{tabular}{l|llllll}
	\toprule
	~ & location
	& source
	& destination
	& time
	& manner
	& amount \\
	\midrule

	\Int & \lan{podi} & \lan{posmad} & \lan{poimos}
	& \lan{poives} & \lan{posme} & \lan{poti} \\

	\Rel & \lan{jodi} & \lan{josmad} & \lan{joimos}
	& \lan{joives} & \lan{josme} & \lan{joti} \\

	\Prox & \lan{sodi} & \lan{sosmad} & \lan{soimos}
	& \lan{soives} & \lan{sosme} & \lan{soti} \\

	\Dist & \lan{todi} & \lan{tosmad} & \lan{toimos}
	& \lan{toives} & \lan{tosme} & \lan{toti} \\

	\Uni & \lan{visodi} & \lan{visosmad} & \lan{visoimos}
	& \lan{visoives} & \lan{visosme} & \lan{visoti} \\

	\Neg & \lan{nipodi} & \lan{niposmad} & \lan{nipoimos}
	& \lan{nipoives} & \lan{niposme} & \lan{nipoti} \\

	\Exist & \lan{potodi} & \lan{potosmad} & \lan{potoimos}
	& \lan{potoives} & \lan{potosme} & \lan{pototi} \\

	\bottomrule
\end{tabular}
\end{center}

When using a pro-adverb that refers to the amount of something, the genitive
case should be used.

(e.g. combien de grand)\INCOMPLETE

proximal referse to objects near the speaker
medial refers to objects near the listener
distal refers to objects far from both

to: that (from pie so, with analogy from its inflected cases)

poti: From PIE *kwoti, adverb from kwos, kwis. Cognate with AG posos and Lat
quot.

For poimos: compare AG peemos (when). Even though the meanings are difrent.

\section{Demonstrative pronouns}

\INCOMPLETE

\section{Interrogative and relative pronouns}

\INCOMPLETE

\chapter{Verbal Morphology}

\section{General properties}

Verbs inflect according to the following categories:
\begin{itemize}
	\item \textbf{person}: first, second, third
	\item \textbf{number}: singular, plural
	\item \textbf{mood}: indicative, subjunctive, imperative
	\item \textbf{voice}: active, mediopassive
	\item \textbf{aspect}: perfective, imperfective
\end{itemize}

The starting point for analyzing the morphology of Danetian verbs is the
stem. The different endings appended at the end of the
verb stem mark the different
inflections of the verb. If the endings are appended directly after a
consonant, as in \lan{vepti}, then the verb is called \textit{athematic}. But
if an intervening vowel is placed between the stem and the ending, as in
\lan{vopeti}, then the verb is called \textit{thematic}.

In the dictionary, verbs are placed in the active indicative present third
person singular, continuing the tradition of Proto-Indo-European.

\section{Thematic verbs}
\label{sec:thematic-verbs}

Thematic verbs are divided into three inflectional
classes for convenience: T1, T2, and T3.

\subsection{Class T1}

The following are examples of class T1 verbs.
\begin{multicols}{2}
\begin{itemize}
	\item \lan{anapreseti} (\V): to pray
	\item \lan{anasedeti} (\V): to be victorious over
	\item \lan{bereti} (\V): to carry
	\item \lan{drageti} (\V): to drag
	\item \lan{flugeti} (\V): to fly
	\item \lan{lekseti} (\V): to protect
	\item \lan{preseti} (\V): to ask, to question
	\item \lan{somjugeti} (\V): to unite, to gather
	\item \lan{sugeti} (\V): to sow
	\item \lan{tendeti} (\V): to stretch, to extend
	\item \lan{trosbereti} (\V): to finish, to complete
	\item \lan{verteti} (\V): to turn over; to spin
	\item \lan{vopeti} (\V): to call
\end{itemize}
\end{multicols}

We choose \lan{vopeti} as a model verb for the inflection table below.
\input{compiled/vopeti.table}

\subsection{Class T2}

The following are examples of class T2 verbs.
\begin{multicols}{2}
\begin{itemize}
	\item \lan{gladjeti} (\V): to make happy, to satisfy
	\item \lan{lufjeti} (\V): to love
	\item \lan{numenjeti} (\V): to breathe
	\item \lan{opigladjeti} (\V): to comfort, to console
	\item \lan{opisagjeti} (\V): to celebrate
	\item \lan{pekaputjeti} (\V): to lead
	\item \lan{pemanjeti} (\V): to remind; to remember
	\item \lan{pisjeti} (\V): to write
	\item \lan{protislabjeti} (\V): to understand
	\item \lan{sagjeti} (\V): to mark, to signify
	\item \lan{satjeti} (\V): to place down
	\item \lan{slabjeti} (\V): to take
	\item \lan{vardjeti} (\V): to work
	\item \lan{vedenjeti} (\V): to water (plants)
	\item \lan{vojasjeti} (\V): to travel
\end{itemize}
\end{multicols}

The inflection table for class T2 verbs is shown below.
\input{compiled/lufjeti.table}

\textit{Note:} There are four sounds that are prone to yod-coalescence:
\begin{itemize}
	\item \lan{k + j = cj}
	\item \lan{g + j = gj}
	\item \lan{n + j = nj}
	\item \lan{s + j = sj}
\end{itemize}

In verbs whose stems end in one of these four palatal sounds, when an ending
such as \lan{-íte} is appended, the palatal sound remains. For example,
contrast the following two examples:
\begin{itemize}
	\item \lan{lufj- + -íte = lufíte}
	\item \lan{sagj- + -íte = sagjíte}
\end{itemize}
Notice that the orthographic \lan{j} is preserved in the second example, but
not the first.


\subsection{Class T3}

Below are some examples of class T3 verbs.
\begin{multicols}{2}
\begin{itemize}
	\item \lan{anapelvíeti} (\V): to multiply
	\item \lan{artíeti} (\V): to taunt, to provoke
	\item \lan{atifestríeti} (\V): to shape
	\item \lan{augíeti} (\V): to show, to display
	\item \lan{bredíeti} (\V): to roam, to wander freely
	\item \lan{donfíeti} (\V): to count
	\item \lan{garaidíeti} (\V): to prepare, to make ready
	\item \lan{opisvomíeti} (\V): to set free
	\item \lan{pleníeti} (\V): to fill
	\item \lan{potíeti} (\V): to govern
	\item \lan{sontíeti} (\V): to send
	\item \lan{svenjíeti} (\V): to sanctify
	\item \lan{vesverdíeti} (\V): to bless
\end{itemize}
\end{multicols}

The conjugation of class T3 verbs is exemplified in the table below.
\input{compiled/sontieti.table}


\section{Athematic verbs}

Athematic verbs are divided into two inflectional classes for convenience:
A1 and A2.

\subsection{Class A1}

Class A1 encompasses thematic verbs that end in \lan{-mi} in the
first person present, like \lan{denati}. These verbs are grouped with athematic
verbs because the endings are identical to athematic verbs, and distinct from
all the thematic verbs discussed in section \ref{sec:thematic-verbs}. Some
examples of A1 verbs are:
\begin{multicols}{2}
\begin{itemize}
	\item \lan{airáti} (\V): to protect
	\item \lan{andjóti} (\V): to end, to finish
	\item \lan{armáti} (\V): to pity
	\item \lan{arnóti} (\V): to harvest, to reap
	\item \lan{badóti} (\V): to frighten
	\item \lan{bloti} (\V): to bloom
	\item \lan{futi} (\V): to run
	\item \lan{groti} (\V): to grow
	\item \lan{opistanati} (\V): to remain
	\item \lan{piboti} (\V): to drink
	\item \lan{pluti} (\V): to rain
	\item \lan{stati} (\V): to stand
	\item \lan{trostanati} (\V): to persevere
\end{itemize}
\end{multicols}

The inflection table for these verbs is shown below.
\input{compiled/stati.table}

\textit{Note:} For verbs in this class, the stress \textit{always} falls on the
stem whenever reduplication or prefixation happens.
For example, we say \lan{fefút} instead of
\lan{fefut}; \lan{gegróme} instead of \lan{gegrome}; \lan{sestáte} instead of
\lan{sestate}; and so on.


\subsection{Class A2}

This class includes regular athematic verbs, such as the ones below.
\begin{multicols}{2}
\begin{itemize}
	\item \lan{dukti} (\V): to pull
	\item \lan{fekti} (\V): to do; to put
	\item \lan{merti} (\V): to die
	\item \lan{vepti} (\V): to speak, to say
\end{itemize}
\end{multicols}

As can be seen, this class is very small, and only includes relic verbs. The
inflection table is shown below.
\input{compiled/merti.table}

\section{Irregular verbs}

All verbs discussed so far have been regular. Irregular verbs are classified
separately, and must be memorized one by one. They are listed below.
\begin{multicols}{2}
\begin{itemize}
	\item \lan{doti} (\V): to give
	\item \lan{esti} (\V): to be
	\item \lan{esti} (\V): to eat
	\item \lan{iti} (\V): to go
	\item \lan{linepti} (\V): to leave
	\item \lan{opíti} (\V): to come
	\item \lan{proíti} (\V): to occur, to pass
	\item \lan{upoíti} (\V): to descend
\end{itemize}
\end{multicols}

The conjugations for \lan{opíti}, \lan{proíti}, and \lan{upoíti} are identical
to that of \lan{iti}, because they are derived from it.

Several tables of irregular verbs are provided as reference.

\input{compiled/esti-be.table}
\input{compiled/iti.table}
\input{compiled/doti.table}
\input{compiled/esti-eat.table}
\input{compiled/linepti.table}



%\chapter{Prepositions}

%NOTE: Perhaps treat adverb-like prepositions as if they were prepositions with a null object

%\section{General overview}

%A preposition accepts a noun or noun phrase.

%\INCOMPLETE

%\chapter{Conjunctions}
%\INCOMPLETE


%\chapter{Derivational morphology}

%\section{Standard word order}

%Danetian is an SVO language, like many other languages in Europe. However,
%thanks to its rich inflectional morphology, this word order is fairly flexible.

%\INCOMPLETE

\chapter{General syntax}

\section{Copula omission}
\label{sec:copula-omission}

The copula can be optionally omitted if it is in the present tense.

\begin{exe}
	\ex
	\gll
	\lan{Aicaterína}
	\lan{en}
	\lan{Gallí}. \\
	Catherine in France-{\Loc} \\
	\glt Catherine is in France.
\end{exe}

\begin{exe}
	\ex
	\gll
	\lan{Petros}
	\lan{mechaniskos}
	\lan{insjenier}. \\
	Peter mechanical engineer \\
	\glt Peter is a mechanical engineer.
\end{exe}

\begin{exe}
	\ex
	\gll
	\lan{Nestor}
	\lan{i}
	\lan{Antonios}
	\lan{physicoi}. \\
	Nestor and Anthony physicists \\
	\glt Nestor and Anthony are physicists.
\end{exe}

\begin{exe}
	\ex
	\gll
	\lan{Benjamín}
	\lan{viros}
	\lan{de}
	\lan{Alexandra}
	\lan{bena}. \\
	Benjamin man and Alexandra woman \\
	\glt Benjamin is a man, and Alexandra is a woman.
\end{exe}

If it's not in the present tense, then the copula can still be omitted,
provided that it was used in the previous sentence. This way,
the tense of the omitted copula is inferred from the copula in the previous
sentence.

\begin{exe}
	\ex
	\gll
	\lan{En}
	\lan{prameni}
	\lan{but}
	\lan{Verdom},
	\lan{da}
	\lan{Verdom}
	\lan{opi}
	\lan{Divosjo},
	\lan{da}
	\lan{Divos}
	\lan{Verdom}. \\
	in beginning-{\Loc} was word and word next-to God-{\Gen} and
	God word \\
	\glt In the beginning was the Word, and the Word was with God, and the
	Word was God.
\end{exe}

\section{Possession}

Unlike most European languages, Danetian does not typically use a verb
\textit{to have} to express possession. Rather, it uses a grammatical
construction: Possessor-{\Dat} copula object-{\Nom}.

\begin{exe}
	\ex
	\gll
	\lan{Medjo}
	\lan{iti}
	\lan{sonts}
	\lan{rufros}
	\lan{automobilis}. \\
	me-{\Dat} it-goes being-{\M} red-{\M} automobile-{\M} \\
	\glt I will have a red car.
\end{exe}

\begin{exe}
	\ex
	\gll
	\lan{Petroi}
	\lan{i}
	\lan{Nestorei}
	\lan{esti}
	\lan{bibliom}
	\lan{physicas}. \\
	Peter-{\Dat} and Nestor-{\Dat} is book physics-{\Gen}. \\
	\glt Peter and Nestor have a physics book.
\end{exe}

\begin{exe}
	\ex
	\gll
	\lan{Aicaterínai}
	\lan{bunt}
	\lan{tres}
	\lan{bles}. \\
	Catherine-{\Dat} had three apples \\
	\glt Catherine had three apples.
\end{exe}

Of course, the copula can be omitted according to the
rules in section \ref{sec:copula-omission}.

\chapter{Time expressions}

\section{Hours and minutes}

\section{Days of the week}

The days of the week are the following:
\begin{itemize}
	\item \lan{din Soles} (\M): Sunday
	\item \lan{din Lunas} (\M): Monday
	\item \lan{din Marties} (\M): Tuesday
	\item \lan{din Mercuriosjo} (\M): Wednesday
	\item \lan{din Jovies} (\M): Thursday
	\item \lan{din Venesos} (\M): Friday
	\item \lan{din Saturnosjo} (\M): Saturday
\end{itemize}

In order to express that an event occured on one of the days of the week,
the dative case is used.
\begin{exe}
	\ex
	\gll
	\lan{Imi}
	\lan{enam}
	\lan{vidonts}
	\lan{dinei}
	\lan{Saturnosjo}. \\
	I-will-go her-{\Acc} seeing-{\M} day-{\Dat} Saturn-{\Gen} \\
	\glt I will see her on Saturday.
\end{exe}

When saying that an event will happen \textit{this week}, or
\textit{that week}, or \textit{next week}, or any
similar phrase, we use the construction \lan{ana} + accusative. A rough
equivalent in English would be to say \textit{upon this week}, for instance.
\begin{exe}
	\ex
	\gll
	\lan{Vojasj}
	\lan{pepramenjet}
	\lan{ana}
	\lan{probanvesam}
	\lan{soliskam}. \\
	Trip started on past-{\Acc} week-{\Acc} \\
	\glt The trip started last week.
\end{exe}

\section{Days of the month}

When saying that something will happen \textit{on that day}, or on any
particular numeric day of the month, we use the genitive case.
\begin{exe}
	\ex
	\gll
	\lan{Urusj}
	\lan{anandjóto}
	\lan{tosjo}
	\lan{memorabilies}
	\lan{dines}. \\
	War concluded that-{\Gen} memorable-{\Gen} day-{\Gen} \\
	\glt The war ended on that memorable day.
\end{exe}
\begin{exe}
	\ex
	\gll
	\lan{Vi}
	\lan{dedamímeda}
	\lan{oktamosjo}
	\lan{juniosjo}. \\
	We were-married eighth-{\Acc} June-{\Acc} \\
	\glt We got married on the eight of June.
\end{exe}

\section{Months and years}

Below is a list of all twelve months of the Western calendar.
\begin{multicols}{2}
\begin{itemize}
	\item \lan{januarios} (\M): January
	\item \lan{februarios} (\M): February
	\item \lan{martios} (\M): March
	\item \lan{aprilis} (\M): April
	\item \lan{majos} (\M): May
	\item \lan{junios} (\M): June
	\item \lan{julios} (\M): July
	\item \lan{augustos} (\M): August
	\item \lan{september} (\M): September
	\item \lan{october} (\M): October
	\item \lan{november} (\M): November
	\item \lan{december} (\M): December
\end{itemize}
\end{multicols}

When saying that something will happen on a particular month or year, we use
the construction \lan{en} + locative. Years are expressed with cardinal
numbers.
\begin{exe}
	\ex
	\gll
	\lan{Dedenáto}
	\lan{en}
	\lan{tusonti}
	\lan{nevani}
	\lan{satonisu}
	\lan{nevandsonti}
	\lan{nevani}. \\
	I-was-born in thousand-{\Loc} nine-{\Loc} hundreds-{\Loc} ninety-{\Loc}
	nine-{\Loc} \\
	\glt I was born in 1999.
\end{exe}
\begin{exe}
	\ex
	\gll
	\lan{Schola}
	\lan{pramenjeti}
	\lan{en}
	\lan{septembri}. \\
	School-{\Nom} starts in September-{\Loc} \\
	\glt School starts in September.
\end{exe}

%\chapter{Negation}

%\chapter{Questions}
%\section{Binary questions}
%\section{Open-ended questions}

%\chapter{Emphasis and word order}



\appendix
\renewcommand{\thesection}{\Alph{section}}
\setcounter{section}{0}

\begin{appendices}

\section{Verbal morphology tables}

\input{compiled/bideti.table}
\input{compiled/bloti.table}
\input{compiled/denati.table}
\input{compiled/doti.table}
\input{compiled/esti-be.table}
\input{compiled/esti-eat.table}
\input{compiled/fekti.table}
\input{compiled/iti.table}
\input{compiled/jekti.table}
\input{compiled/linepti.table}
\input{compiled/lufjeti.table}
\input{compiled/merti.table}
\input{compiled/stati.table}
\input{compiled/sontieti.table}
\input{compiled/vepti.table}
\input{compiled/vopeti.table}


\pagebreak
\section{Western personal names}

\begin{multicols}{2}
\begin{itemize}
	\item \lan{Abrahám} (\M): Abraham
	\item \lan{Abrám} (\M): Abram
	\item \lan{Adám} (\M): Adam
	\item \lan{Aicaterína} (\F): Catherine
	\item \lan{Albértos} (\M): Albert
	\item \lan{Alexandros, -dra} (\M, \F): Alexander; Alexandra
	\item \lan{Alexios, -xia} (\M, \F): Alexius; Alexia
	\item \lan{Anania} (\M): Ananias
	\item \lan{Anastasios, -sia} (\M, \F): Anastasius; Anastasia
	\item \lan{Anatolios} (\M): Anatolius, Anatoly
	\item \lan{Anaximandros, -dra} (\M, \F): Anaximander; Anaximandra
	\item \lan{Andréa} (\M, \F): Andrew, Andres, Andrei; Andrea
	\item \lan{Andromeda} (\F): Andromeda
	\item \lan{Andronicos} (\M): Andronicus
	\item \lan{Angelos, -la} (\M, \F): Angelus; Angela
	\item \lan{Anna} (\F): Anna, Anne
	\item \lan{Antiochos} (\M): Antiochus
	\item \lan{Antonios} (\M): Anthony
	\item \lan{Apollodoros} (\M): Apollodorus, Apollodore
	\item \lan{Apollonios, -nia} (\M, \F): Apollonius; Apollonia
	\item \lan{Arcadios} (\M): Arcadius
	\item \lan{Aristarchos} (\M): Aristarchus
	\item \lan{Aristotelis} (\M): Aristotle
	\item \lan{Arón} (\M): Aaron
	\item \lan{Artemios} (\M): Artemius, Artyom
	\item \lan{Athanasios, -sia} (\M, \F): Athanasius; Athanasia
	\item \lan{Augustínos} (\M): Augustine
	\item \lan{Barbara} (\F): Barbara
	\item \lan{Barnaba} (\M): Barnabas
	\item \lan{Bartholomaios} (\M): Bartholomew
	\item \lan{Basilios} (\M): Basilius, Basil, Vasily
	\item \lan{Benjamín} (\M): Benjamin
	\item \lan{Braján} (\M): Bryan, Brian
	\item \lan{Carlos, -la} (\M, \F): Charles, Carlos, Karl; Carla, Carol
	\item \lan{Cassandros, -dra} (\M; \F): Cassander; Cassandra
	\item \lan{Christiános} (\M): Christian
	\item \lan{Christína} (\F): Christine
	\item \lan{Christophoros} (\M): Christopher
	\item \lan{Clement} (\M): Clement
	\item \lan{Cleopatra} (\F): Cleopatra
	\item \lan{Condrados} (\M): Conrad, Konrad
	\item \lan{Constantínos} (\M): Constantine
	\item \lan{Cynthia} (\F): Cynthia
	\item \lan{Cyriacos} (\M): Cyriacus
	\item \lan{Cyrillos} (\M): Cyril, Kirill
	\item \lan{Cyros} (\M): Cyrus
	\item \lan{Daniel, -ella} (\M, \F): Daniel; Daniella
	\item \lan{David} (\M): David
	\item \lan{Demetrios, -tria} (\M, \F): Demetrius, Dmitri; Demetria
	\item \lan{Dionysios, -sia} (\M, \F): Dionysius, Denis; Dionysia
	\item \lan{Dioscuros} (\M): Dioscurus
	\item \lan{Donald} (\M): Donald
	\item \lan{Dylán} (\M): Dylan
	\item \lan{Edvárdos} (\M): Edward
	\item \lan{Elia} (\M): Elijah, Elias, Ilya
	\item \lan{Elisabet} (\F): Elisabeth
	\item \lan{Emma} (\F): Emma
	\item \lan{Emmanuel} (\M): Immanuel, Emmanuel
	\item \lan{Ephrém} (\M): Ephraim
	\item \lan{Erik} (\M): Eric
	\item \lan{Ernéstos} (\M): Ernest, Ernst, Ernesto
	\item \lan{Esaía} (\M): Isaiah
	\item \lan{Esther} (\F): Esther
	\item \lan{Eugenia} (\F): Eugenia
	\item \lan{Európa} (\F): Europa
	\item \lan{Eusebios} (\M): Eusebius
	\item \lan{Eva} (\F): Eve
	\item \lan{Evangelos} (\M): Evangelus
	\item \lan{Fernándos} (\M): Ferdinand
	\item \lan{Francis} (\M): Franciscus, Francis
	\item \lan{Frederik} (\M): Frederick, Friedrich, Federico
	\item \lan{Gabriel} (\M): Gabriel
	\item \lan{Georgios} (\M): George
	\item \lan{Gervaldína} (\F): Geraldine
	\item \lan{Gervaldos} (\M): Gerrald
	\item \lan{Gilbértos} (\M): Gilbert
	\item \lan{Glen} (\M): Glen
	\item \lan{Gloria} (\M): Gloria
	\item \lan{Gregorios} (\M): Gregory
	\item \lan{Hadriános} (\M): Adrian, Hadrian
	\item \lan{Heléna} (\F): Helen, Ellen
	\item \lan{Homéros} (\M): Homer
	\item \lan{Hoséa} (\M): Hosea
	\item \lan{Hendrik} (\M): Henry, Heinrich, Enrique
	\item \lan{Ignatios} (\M): Ignatius
	\item \lan{Iréna} (\F): Irene
	\item \lan{Isak} (\M): Isaac
	\item \lan{Isdrael} (\M): Israel
	\item \lan{Isidoros} (\M): Isidore
	\item \lan{Jacobos} (\M): Jacob, James, Jack
	\item \lan{Jasón} (\M): Jason
	\item \lan{Jeremia} (\M): Jeremiah
	\item \lan{Jeronymos} (\M): Jerome
	\item \lan{Jesus} (\M): Jesus, Joshua
	\item \lan{Joakím} (\M): Joachim
	\item \lan{Joanna} (\M, \F): John; Joanne
	\item \lan{Joel} (\M): Joel
	\item \lan{Jona} (\M): Jonah
	\item \lan{Jonathán} (\M): Jonathan
	\item \lan{Josephos} (\M): Joseph
	\item \lan{Jotfredos} (\M): Jeffrey, Geoffrey, Gottfried
	\item \lan{Juda} (\M): Judas, Jude
	\item \lan{Julios, -lia} (\M, \F): Julius, Joules; Julia
	\item \lan{Justínos} (\M): Justin
	\item \lan{Laurentios} (\M): Lawrence
	\item \lan{Lazaros} (\M): Lazarus
	\item \lan{León} (\M): Leo
	\item \lan{Leonida} (\M): Leonidas, Leonid
	\item \lan{Luca} (\M): Luke
	\item \lan{Ludvigos} (\M): Ludwig, Louis, Luis
	\item \lan{Lydia} (\F): Lydia
	\item \lan{Magdalena} (\M): Magdalene
	\item \lan{Marcos} (\M): Marcus, Mark
	\item \lan{Margaríta} (\F): Margarita, Margareth
	\item \lan{Marios, ria} (\M, \F): Marius, Mario; Mary
	\item \lan{Martha} (\F): Martha
	\item \lan{Matthaios} (\M): Matthew
	\item \lan{Mauritios} (\M): Maurice
	\item \lan{Maximos} (\M): Maximus, Maksim
	\item \lan{Methodios} (\M): Methodius
	\item \lan{Moýsis} (\M): Moses
	\item \lan{Nectarios} (\M): Nectarius
	\item \lan{Nestor} (\M): Nestor
	\item \lan{Nicanor} (\M): Nicanor
	\item \lan{Nicator} (\M): Nicator
	\item \lan{Nicodemos} (\M): Nicodemus
	\item \lan{Nicola} (\M): Nicholas
	\item \lan{Nikephoros} (\M): Nikephor
	\item \lan{Nikéta} (\M): Niketas, Nikita
	\item \lan{Nils} (\M): Nils
	\item \lan{Noel} (\M): Noel
	\item \lan{Omar} (\M): Omar
	\item \lan{Paulos} (\M): Paul
	\item \lan{Petros} (\M): Peter
	\item \lan{Philemón} (\M): Philemon
	\item \lan{Philippos} (\M): Philip
	\item \lan{Platón} (\M): Plato
	\item \lan{Procopios} (\M): Procopius
	\item \lan{Prolemaios} (\M): Ptolemy
	\item \lan{Raimondos} (\M): Raymond, Ramón
	\item \lan{Raján} (\M): Ryan
	\item \lan{Raphael} (\M): Raphael
	\item \lan{Renvaldos} (\M): Reginald, Ronald, Reinhold
	\item \lan{Richárdos} (\M): Richard
	\item \lan{Robértos} (\M): Robert
	\item \lan{Romános} (\M): Roman
	\item \lan{Ruth} (\F): Ruth
	\item \lan{Samuel} (\M): Samuel
	\item \lan{Sebastián} (\M): Sebastian
	\item \lan{Seléna} (\F): Selena
	\item \lan{Seleucos} (\M): Seleucus
	\item \lan{Sergios} (\M): Sergius
	\item \lan{Severos} (\M): Severus
	\item \lan{Sila} (\M): Silas, Saul
	\item \lan{Simón} (\M): Simon
	\item \lan{Sophia} (\F): Sophia
	\item \lan{Stephanos} (\M): Stephen, Steven
	\item \lan{Theodoros} (\M): Theodore
	\item \lan{Theodosios, -sia} (\M, \F): Theodosius; Theodosia
	\item \lan{Theophilos} (\M): Theophilus
	\item \lan{Thoma} (\M): Thomas
	\item \lan{Timotheos} (\M): Timothy
	\item \lan{Titos} (\M): Titus
	\item \lan{Tobia} (\M): Tobias
	\item \lan{Valentínos} (\M): Valentinus
	\item \lan{Valerios, -ria} (\M, \F): Valerius; Valeria
	\item \lan{Victor} (\M): Victor
	\item \lan{Vilmos} (\M): William, Guillermo, Wilhelm
	\item \lan{Xavier} (\M): Javier, Xavier
\end{itemize}
\end{multicols}




\end{appendices}

\end{document}
