
	

	
		\key{a.}  n. letter \textit{a}.
	

	
		\key{abasjur.} {\footnotesize (From Fr. \ipa{abat-jour}).} m. lampshade.
	

	
		\key{abjón.} {\footnotesize (From PG \ipa{*abjǭ}).} f.
			\textbf{1.} ebb, the receding movement of the tide.
			\textbf{2.} a gradual decline.
	

	\key{abol,}
		\textit{bles, balmos.} {\footnotesize (From PIE \ipa{*h₂ébōl}).} m. apple.
	

	
		\key{Abrahám.} {\footnotesize (From Lat. \ipa{Ābrahām}).} m. Abraham.
	

	
		\key{Abrám.}  m. Abram.
	

	
		\key{actiniom.}  n.{\footnotesize \textit{ chemistry.}} actinium.
	

	
		\key{adalos.} {\footnotesize (From PG \ipa{*adisô}).} m.
			\textbf{1.} liquid manure.
			\textbf{2.} filth.
			\textbf{3.} mire, mud.
			\textbf{4.} urine.
	

	
		\key{Adám.} {\footnotesize (From AG \gr{Ᾱ̓δᾱ́μ}).} m. Adam.
	

	
		\key{adisa.} {\footnotesize (From PG \ipa{*adisô}).} m.
			\textbf{1.} adze.
			\textbf{2.} axe.
	

	
		\key{administratión.} {\footnotesize (From Lat. \ipa{administrātiō}).} f. administration.
	

	
		\key{adrjos.} {\footnotesize (From PIE \ipa{*h₂éǵr̥yos}).} adj. campestral, pertaining to fields.
	

	
		\key{adros.} {\footnotesize (From PIE \ipa{*h₂éǵros}).} m. field.
	

	
		\key{aerodromos.} {\footnotesize (From Fr. \ipa{aérodrome}, by analogy with \ipa{hippodrome}, from AG \gr{ἱππόδρομος}).} m. airfield, aerodrome.
	

	
		\key{aeroportos.} {\footnotesize (From Fr. \ipa{aéroport}, with a latinized ending).} m. airport.
	

	
		\key{afísja.} {\footnotesize (From Fr. \ipa{affiche}).} f.
			\textbf{1.} poster.
			\textbf{2.} placard.
			\textbf{3.} sign.
	

	
		\key{aflom.} {\footnotesize (From PG \ipa{*aflą}).} n. power, strength, might.
	

	
		\key{aflos.} {\footnotesize (From PG \ipa{*aflaz}).} m. hearth.
	

	
		\key{Africa.} {\footnotesize (From Lat. \ipa{Āfrica}).} f.{\footnotesize \textit{ continent.}} Africa.
	

	
		\key{ageti.} {\footnotesize (From PG \ipa{*aganą}).} v. to fear, to be afraid.
	

	
		\key{agida.} {\footnotesize (From PG \ipa{*agiþō}).} f. harrow.
	

	
		\key{agja.} {\footnotesize (From PG \ipa{*agjō}).} f. edge, corner.
	

	
		\key{agjostainos.} {\footnotesize (From PG \ipa{*agjōstainaz}).} m. sharp-stone, cornerstone.
	

	
		\key{agjóti.} {\footnotesize (From PG \ipa{*agjōną}).} v. to harrow.
	

	
		\key{aglatiom.} {\footnotesize (From PG \ipa{*agalaitiją}).} n.
			\textbf{1.} licentiousness.
			\textbf{2.} lewdness, lasciviousness.
			\textbf{3.} impudence, importunity.
	

	
		\key{aglíeti.} {\footnotesize (From PG \ipa{*aglijaną}).} v.
			\textbf{1.} to damage.
			\textbf{2.} to offend.
	

	
		\key{aglos.} {\footnotesize (From PG \ipa{*agraz}).} m. flood.
	

	
		\key{Agnis.} {\footnotesize (From PIE \ipa{*h₁n̥gʷnís}).} m. Danetian god of fire, oaths, promises, and moral justice.
	

	
		\key{agnos.} {\footnotesize (From PIE \ipa{*h₂egʷnós}).} m. lamb.
	

	
		\key{agos.} {\footnotesize (From PIE \ipa{*h₂égʰos}).} n.
			\textbf{1.} (emotional) distress.
			\textbf{2.} fear, dread, terror.
	

	
		\key{ahana.} {\footnotesize (From PG \ipa{*ahanō}).} f. chaff, awns.
	

	
		\key{ahla.} {\footnotesize (From PG \ipa{*anhulō}).} f. strap, band, thong.
	

	
		\key{ahos.} {\footnotesize (From PG \ipa{*ahaz}).} n. ear of grain.
	

	
		\key{ahurnos.} {\footnotesize (From PG \ipa{*ahurnaz}).} m. maple.
	

	
		\key{Aicaterína.} {\footnotesize (From AG \gr{Αἰκατερῑ́νη}).} f. Catherine, Catalina, Katrina, Caitlin.
	

	
		\key{Aicaterinapolis.} {\footnotesize (From Dan. \ipa{Aicaterina + -polis}).} f. Yekaterinburg.
	

	
		\key{aidom.} {\footnotesize (From PG \ipa{*aidą}).} n. isthmus, strait.
	

	
		\key{aidos.} {\footnotesize (From PG \ipa{*aidaz}).} m. fire, pyre.
	

	
		\key{aigineti.} {\footnotesize (From PG \ipa{*aigināną}).} v. to appropriate, to take possession by force.
	

	
		\key{aiglos.} {\footnotesize (From PG \ipa{*aiglaz}).} m. shoot, barb.
	

	
		\key{Aigyptos.} {\footnotesize (From Lat. \ipa{Aegyptus}, from AG \gr{Αἴγυπτος}, from Egyptian \ipa{ḥwt-kꜣ-ptḥ}).} f. Egypt.
	

	
		\key{aikila.} {\footnotesize (From PG \ipa{*aikilǭ}).} f. acorn.
	

	
		\key{ailíeti.} {\footnotesize (From PG \ipa{*ailijaną}).} v. to kindle, to light on fire.
	

	
		\key{aimos.} {\footnotesize (From PG \ipa{*aimaz}).} m. steam, smoke.
	

	
		\key{aimurja.} {\footnotesize (From PG \ipa{*aimuzjǭ}).} f. ember, hot ash.
	

	
		\key{ainacja.} {\footnotesize (From PG \ipa{*ainakjǭ}).} f. widow.
	

	
		\key{ainios.} {\footnotesize (From PG \ipa{*ainijaz}).} m. juniper.
	

	
		\key{ainstainiom.}  n.{\footnotesize \textit{ chemistry.}} einsteinium.
	

	
		\key{aiqual.} {\footnotesize (From Lat. \ipa{aequālis}).} adj. equal.
	

	
		\key{aira.} {\footnotesize (From PG \ipa{*airō}).} f. oar.
	

	
		\key{aira.} {\footnotesize (From PG \ipa{*aizō}).} f.
			\textbf{1.} reverence.
			\textbf{2.} honour.
	

	
		\key{airáti.} {\footnotesize (From PG \ipa{*aizāną}).} v. to honor.
	

	
		\key{airi.} {\footnotesize (From PIE \ipa{*h₂éyeri}).} adv. early in the morning.
	

	
		\key{airístom.} {\footnotesize (From PIE \ipa{*h₂éyeri-h₁d-to-m}).} n. breakfast.
	

	
		\key{airos.} {\footnotesize (From PG \ipa{*airuz}).} m.
			\textbf{1.} messenger.
			\textbf{2.} servant.
			\textbf{3.} herald.
	

	
		\key{airundiom.} {\footnotesize (From PG \ipa{*airundiją}).} n.
			\textbf{1.} message.
			\textbf{2.} errand.
	

	
		\key{aisa.} {\footnotesize (From PG \ipa{*aisǭ}).} f. fireplace.
	

	
		\key{aisóti.} {\footnotesize (From PG \ipa{*aisōną}).} v. to rush.
	

	
		\key{aistáti.} {\footnotesize (From PG \ipa{*aistāną}).} v. to respect.
	

	
		\key{aitos.} {\footnotesize (From PG \ipa{*aitaz}).} m. ulcer.
	

	
		\key{aitrom.} {\footnotesize (From PG \ipa{*aitrą}).} n.
			\textbf{1.} poison.
			\textbf{2.} pus.
	

	
		\key{aive.} {\footnotesize (From Dan. \ipa{aivos}).} adv. forever, always.
	

	
		\key{aivíeti.} {\footnotesize (From PG \ipa{*aiwijaną}).} v. to graze.
	

	
		\key{aivos.} {\footnotesize (From PG \ipa{*aiwaz}).} m. eternity.
	

	
		\key{ajos.} {\footnotesize (From PIE \ipa{*h₂éyos}).} n.{\footnotesize \textit{ chemistry.}} copper.
	

	
		\key{aketi.} {\footnotesize (From PG \ipa{*akaną}).} v. to drive, sail, navigate.
	

	
		\key{akis.} {\footnotesize (From PG \ipa{*akiz}).} m. ache, pain.
	

	
		\key{akros.} {\footnotesize (From PG \ipa{*akraz}).} m. farm.
	

	
		\key{aksa.} {\footnotesize (From PG \ipa{*ahsō}).} f. axle.
	

	
		\key{aksla.} {\footnotesize (From PG \ipa{*ahslō}).} f. shoulder.
	

	
		\key{alasna.} {\footnotesize (From PG \ipa{*alasnō}).} f. awl.
	

	
		\key{Albania.} {\footnotesize (From Lat. \ipa{Albania}, from BG \gr{Ἀλβανία}).} f. Albania.
	

	
		\key{Albertos Ainstain.}  m. Albert Einstein.
	

	
		\key{aldíeti.} {\footnotesize (From PG \ipa{*aldijaną}).} v. to delay.
	

	
		\key{aldis.} {\footnotesize (From PG \ipa{*aldiz}).} f. age, generation, lifetime.
	

	
		\key{aldrom.} {\footnotesize (From PG \ipa{*aldrą}).} n. lifetime, age, generation.
	

	
		\key{aleti.} {\footnotesize (From PG \ipa{*alaną}).} v.
			\textbf{1.} to grow.
			\textbf{2.} to grow old, to mature.
			\textbf{3.} to breed.
	

	
		\key{Alexandros.} {\footnotesize (From AG \gr{Ἀλέξανδρος}).} m. Alexander, Alejandro.
	

	\key{alfit,}
		\textit{alfites, alfitmos.} {\footnotesize (From PIE \ipa{*h₂élbʰit}).} n. barley.
	

	
		\key{alfos.} {\footnotesize (From PIE \ipa{*albʰós}).} adj. white.
	

	
		\key{algis.} {\footnotesize (From PG \ipa{*algiz}).} m. elk.
	

	
		\key{algóti.} {\footnotesize (From PG \ipa{*algōną}).} v. to defend, to protect.
	

	
		\key{alina.} {\footnotesize (From PG \ipa{*alinō}).} f. elbow.
	

	
		\key{alisa.} {\footnotesize (From PIE \ipa{*h₂élis-}).} f. alder.
	

	
		\key{alisketi.} {\footnotesize (From PG \ipa{*aliskōną}).} v. to love, to cherish.
	

	
		\key{aljanom.} {\footnotesize (From PG \ipa{*aljaną}).} n.
			\textbf{1.} zeal, eagerness.
			\textbf{2.} strength, courage.
	

	
		\key{aljos.} {\footnotesize (From PIE \ipa{*h₂élyos}).} pron. other, another.
	

	
		\key{alka.} {\footnotesize (From PG \ipa{*alkǭ}).} f. a kind of sea bird, auk.
	

	\key{alks,}
		\textit{alkes.} {\footnotesize (From PG \ipa{*alhs}).} m.
			\textbf{1.} shelter.
			\textbf{2.} building.
			\textbf{3.} temple, sanctuary.
			\textbf{4.} settlement, residence.
	

	
		\key{alnóti.} {\footnotesize (From PG \ipa{*aljanōną}).} v. to be zealous for, to strive for.
	

	
		\key{alpha.}  f. Greek letter alpha.
	

	
		\key{alsóti.} {\footnotesize (From PG \ipa{*aliskōną}).} v. to tend to, to care for.
	

	
		\key{altin.} {\footnotesize (From a Turkic language. Compare Turkmen \ipa{altyn} and Uzbek \ipa{oltin}).} m.{\footnotesize \textit{ chemistry.}} gold.
	

	
		\key{altjeti.} {\footnotesize (From PG \ipa{*alatjaną}).} v. to drive, to move.
	

	
		\key{alu.} {\footnotesize (From PG \ipa{*alu}).} n. beer, ale.
	

	
		\key{aluminiom.}  n.{\footnotesize \textit{ chemistry.}} aluminium, aluminum.
	

	
		\key{alvis.} {\footnotesize (From PG \ipa{*albiz}).} m. elf, fairy.
	

	
		\key{amara.} {\footnotesize (From PG \ipa{*amazǭ}).} f.
			\textbf{1.} blackbird.
			\textbf{2.} yellowhammer.
	

	
		\key{amba.} {\footnotesize (From PG \ipa{*ambô}).} m. belly.
	

	
		\key{ambaktos.} {\footnotesize (From PG \ipa{*ambahtaz}).} m. servant.
	

	
		\key{ambindeti.} {\footnotesize (From PG \ipa{*andabindaną}).} v. to unbind, to loosen, to set free.
	

	
		\key{amblis.} {\footnotesize (From PG \ipa{*amliz}).} m.
			\textbf{1.} spelt, emmer.
			\textbf{2.} starch produced from it.
	

	
		\key{ambudeti.} {\footnotesize (From PG \ipa{*anabeudaną}).} v. to command, to order, to instruct.
	

	
		\key{ambúsnis.} {\footnotesize (From PG \ipa{*anabūsniz}).} f.
			\textbf{1.} commandment.
			\textbf{2.} precept.
	

	
		\key{America.} {\footnotesize (From Lat. \ipa{America}).} f. America.
	

	
		\key{americaniskos.} {\footnotesize (From Dan. \ipa{America}).} adj. American.
	

	
		\key{americiom.}  n.{\footnotesize \textit{ chemistry.}} americium.
	

	
		\key{amos.} {\footnotesize (From PIE \ipa{*h₂eh₃mós}).} adj.
			\textbf{1.} raw, uncooked.
			\textbf{2.} bitter.
	

	
		\key{Amsterdám.}  m. Amsterdam, capital of the Netherlands.
	

	
		\key{ana.} {\footnotesize (From PIE \ipa{*h₂en-h₂}).} prep.
			\textbf{1.} (+loc) on.{\footnotesize \ipa{Enosmi pulnis ana enosjo kaputi}. \textit{He has hair on his haid}.}
			\textbf{2.} (+acc) onto.
			\textbf{3.} (+acc) along.
	

	
		\key{anabideti.} {\footnotesize (From Dan. \ipa{ana- + bideti}).} v.
			\textbf{1.} to share, to exchange.
			\textbf{2.} to allot, to give, to provide.
			\textbf{3.} to confide, to tell.
			\textbf{4.} to endow.
	

	
		\key{analysis.} {\footnotesize (From Lat. \ipa{analysis}, from AG \gr{ᾰ̓νᾰ́λῠσῐς}).} f. analysis.
	

	
		\key{anamos.} {\footnotesize (From PIE \ipa{*h₂enh₁mos}).} m. breath.
	

	
		\key{anapelvíeti.} {\footnotesize (From Dan. \ipa{ana- + pelv- + -íeti}).} v. to multiply.
	

	
		\key{anapreseti.} {\footnotesize (From Dan. \ipa{preseti}).} v. to pray (Use the dative case on the person being prayed to.).
	

	
		\key{anapresetis.} {\footnotesize (From Dan. \ipa{anapreseti}).} f. prayer.
	

	
		\key{anasedeti.} {\footnotesize (From Dan. \ipa{anasedos}. $\blacklozenge$~This verb is usually paired with \ipa{+ úper X}, to denote the meaning of being victorious over something or someone.).} v. to triumph, to be victorious over, to win.
	

	
		\key{anasedos.} {\footnotesize (From Dan. \ipa{ana- + sedos}).} n. victory.
	

	
		\key{anats.} {\footnotesize (From PIE \ipa{*h₂énh₂ts}).} f. duck.
	

	
		\key{ancja.} {\footnotesize (From PG \ipa{*ankijǭ}).} f. ankle.
	

	
		\key{anda.} {\footnotesize (From PG \ipa{*anadô}).} m. zeal.
	

	
		\key{andeti.} {\footnotesize (From PIE \ipa{*h₂enǵʰ-}).} v. to obligate, to bind.
	

	
		\key{andios.} {\footnotesize (From PG \ipa{*andijaz}).} m. end.
	

	
		\key{andjóti.} {\footnotesize (From PG \ipa{*andijōną}).} v. to end, to bring to an end, to finish.
	

	
		\key{Andórra.} {\footnotesize (From Sp. \ipa{Andorra}. Compare Fr. \ipa{Andorre}).} f. Andorra.
	

	
		\key{Andréa.} {\footnotesize (From AG \gr{Ἀνδρέᾱς}).} m.
			\textbf{1.} Andrew, André, Andrés. f.
			\textbf{2.} Andrea.
	

	
		\key{Anferai Fedomes.}  f. Low Countries, Netherlands.
	

	
		\key{anferos.} {\footnotesize (From PIE \ipa{*h₁n̥dʰér-o-s}).} adj. low.
	

	
		\key{anfis.} {\footnotesize (From PIE \ipa{*h₂éngʷʰis}).} m. snake.
	

	
		\key{anfos.} {\footnotesize (From PIE \ipa{*h₂éndʰos}).} n. bloom.
	

	
		\key{anfros.} {\footnotesize (From PIE \ipa{*n̥bʰrós}).} m.
			\textbf{1.} rain-cloud.
			\textbf{2.} rain.
	

	
		\key{anfus.} {\footnotesize (From PIE \ipa{*h₂énǵʰus}).} adj. narrow, tight.
	

	
		\key{anga.} {\footnotesize (From PG \ipa{*angô}).} m.
			\textbf{1.} a bend, crook.
			\textbf{2.} hook, barb, angle.
			\textbf{3.} spear, dart.
			\textbf{4.} arrow.
	

	
		\key{angastis.} {\footnotesize (From PG \ipa{*angustiz}).} f.
			\textbf{1.} narrowness, tightness.
			\textbf{2.} fear, anxiety.
	

	
		\key{angja.} {\footnotesize (From PG \ipa{*angijō}).} f. a low lying meadow in a valley or near a river.
	

	
		\key{angla.} {\footnotesize (From PG \ipa{*angulō}).} f. a hook, fishhook.
	

	
		\key{Anglia.}  f. England.
	

	
		\key{angliska.} {\footnotesize (From Dan. \ipa{Anglia}).} f. English language.
	

	
		\key{angros¹.} {\footnotesize (From PG \ipa{*angraz}).} m. lowland, meadow.
	

	
		\key{angros².} {\footnotesize (From PG \ipa{*angazaz}).} m. worry, grief, anxiety.
	

	
		\key{Ankara.}  f. Ankara, capital of Turkey.
	

	
		\key{anklos.} {\footnotesize (From PG \ipa{*ankulaz}).} m. hip.
	

	
		\key{ankos.} {\footnotesize (From PIE \ipa{*h₂énkos}).} n. curve, bend.
	

	
		\key{ankva.} {\footnotesize (From PG \ipa{*ankwô}).} m. ointment, butter.
	

	
		\key{ans.} {\footnotesize (Unknown).} prep. (+gen) without.
	

	
		\key{ansaketi.} {\footnotesize (From PG \ipa{*andasakaną}).} v. to speak against, to refute.
	

	
		\key{ansja.} {\footnotesize (From PG \ipa{*ansijō}).} f. eyelet, eye of a needle.
	

	
		\key{ansos.} {\footnotesize (From PG \ipa{*ansaz}).} m. beam, wooden beam, support beam.
	

	
		\key{anstíeti.} {\footnotesize (From PG \ipa{*anstijaną}).} v. to ask, to request.
	

	
		\key{anstis.} {\footnotesize (From PG \ipa{*anstiz}).} f.
			\textbf{1.} favor, affection, grace.
			\textbf{2.} joy.
	

	
		\key{ansuros.} {\footnotesize (From PG \ipa{*ansuz}).} adj. powerful.
	

	
		\key{Antarctica.} {\footnotesize (From Lat. \ipa{Antarctica}).} f.{\footnotesize \textit{ continent.}} Antarctica.
	

	
		\key{antjos.} {\footnotesize (From PIE \ipa{*h₂entíos}).} m.
			\textbf{1.} front.
			\textbf{2.} forehead.
	

	
		\key{Antonios.} {\footnotesize (From Lat. \ipa{Antonius}).} m. Anthony.
	

	
		\key{anvlitiom.} {\footnotesize (From PG \ipa{*andawlitją}).} n. appearance, form.
	

	
		\key{anvurdjom.} {\footnotesize (From PG \ipa{*andawurdiją}).} n. answer, reply.
	

	
		\key{apaldros.} {\footnotesize (From PG \ipa{*apaldraz}).} m. apple tree.
	

	
		\key{aplos.} {\footnotesize (From PG \ipa{*aplaz}).} m. potato.
	

	
		\key{apo.} {\footnotesize (From PIE \ipa{*h₂epó}).} prep. (+abl) off, away, from.
	

	
		\key{apoluketis.} {\footnotesize (From Dan. \ipa{apolúketor}).} f.
			\textbf{1.} appearance, the way in which something seems.
			\textbf{2.} apparition.
	

	
		\key{apolúketor.} {\footnotesize (From Dan. \ipa{apo- + lúketor}).} v. to appear.
	

	
		\key{aprilis.} {\footnotesize (From Lat. \ipa{Aprīlis}).} m.{\footnotesize \textit{ month.}} April.
	

	
		\key{arbaidis.} {\footnotesize (From PG \ipa{*arbaidiz}).} f.
			\textbf{1.} labour, work, toil.
			\textbf{2.} hardship, suffering.
	

	
		\key{arbja.} {\footnotesize (From PG \ipa{*arbijô}).} m. heir.
	

	
		\key{arbjom.} {\footnotesize (From PG \ipa{*arbiją}).} n. heritage, inheritance.
	

	
		\key{arbos.} {\footnotesize (From PG \ipa{*arbaz}).} m. working animal, ox.
	

	
		\key{ardantom.} {\footnotesize (From PIE \ipa{*h₂r̥ǵn̥tóm}).} n.{\footnotesize \textit{ chemistry.}} silver.
	

	
		\key{ardis.} {\footnotesize (From PG \ipa{*ardiz}).} m.
			\textbf{1.} characteristic, inborn quality.
			\textbf{2.} nature, character.
			\textbf{3.} type; kind; manner; sort.
	

	
		\key{argíeti.} {\footnotesize (From PG \ipa{*argijaną}).} v.
			\textbf{1.} to spoil, to ruin.
			\textbf{2.} to become a coward.
	

	
		\key{argón.}  m.{\footnotesize \textit{ chemistry.}} argon.
	

	
		\key{arja.}  f. unit of area equivalent to one hundred square meters.
	

	
		\key{armáti.} {\footnotesize (From PG \ipa{*armāną}).} v. to pity.
	

	
		\key{Armenia.} {\footnotesize (From Lat. \ipa{Armenia}, from AG \gr{Ἀρμενία}).} f. Armenia.
	

	
		\key{armodiom.} {\footnotesize (From PG \ipa{*armōdiją}).} n. poverty.
	

	
		\key{armos.} {\footnotesize (From PG \ipa{*armaz}).} m. arm.
	

	
		\key{arnis.} {\footnotesize (From PG \ipa{*azaniz}).} f. harvest, harvest time.
	

	
		\key{arnóti.} {\footnotesize (From PG \ipa{*azanōną}).} v. to harvest, to reap.
	

	
		\key{arsenicom.}  n.{\footnotesize \textit{ chemistry.}} arsenic.
	

	
		\key{arsos.} {\footnotesize (From PG \ipa{*arsaz}).} m.{\footnotesize \textit{ vulgar.}} arse, ass, anus, buttocks.
	

	
		\key{arta.} {\footnotesize (From PG \ipa{*artǭ}).} f. wagtail, a type of bird.
	

	
		\key{artíeti.} {\footnotesize (From PG \ipa{*artijaną}).} v. to tease, to taunt, to grill, to provoke.
	

	
		\key{Artobolos.} {\footnotesize (From Dan. \ipa{artos + bolos}).} m. Danetian personal name, meaning \textit{strong bear}.
	

	
		\key{Artoner.} {\footnotesize (From Dan. \ipa{artos + ner}).} m. Danetian personal name, meaning \textit{powerful bear}.
	

	
		\key{artos.} {\footnotesize (From PIE \ipa{*h₂ŕ̥tḱos}).} m. bear.
	

	
		\key{artrom.} {\footnotesize (From PIE \ipa{*h₂érh₃trom}).} n. plough, ard.
	

	\key{arts,}
		\textit{artes.} {\footnotesize (From PG \ipa{*aruts}).} m. ore.
	

	
		\key{arva¹.} {\footnotesize (From PG \ipa{*arfô}).} m. chickweed.
	

	
		\key{arva².} {\footnotesize (From PG \ipa{*arhwō}).} f. arrow.
	

	
		\key{arvíeti.} {\footnotesize (From PG \ipa{*arbijaną}).} v. to inherit.
	

	\key{arvíts,}
		\textit{arvítes.} {\footnotesize (From PG \ipa{*arwīts}).} f. pea.
	

	
		\key{arvos.} {\footnotesize (From PG \ipa{*arwaz}).} n. scar.
	

	
		\key{Asia.} {\footnotesize (From AG \gr{Ἀσῐ́ᾱ}).} f.{\footnotesize \textit{ continent.}} Asia.
	

	
		\key{asja.} {\footnotesize (From PG \ipa{*asjǭ}).} f.{\footnotesize \textit{ geology.}} soapstone.
	

	
		\key{aska.} {\footnotesize (From PG \ipa{*askǭ}).} f. ash, ashes.
	

	
		\key{askos.} {\footnotesize (From PG \ipa{*askaz}).} m.{\footnotesize \textit{ botany.}} ash tree.
	

	\key{asmón,}
		\textit{samnes, smanmos.} {\footnotesize (From PIE \ipa{*h₂éḱmō}).} m. stone.
	

	
		\key{asna.} {\footnotesize (From PG \ipa{*asnō}).} f.
			\textbf{1.} harvest.
			\textbf{2.} earning, wage, payment.
	

	
		\key{asnjos.} {\footnotesize (From PG \ipa{*asnijaz}).} m. servant, laborer.
	

	
		\key{aspa.} {\footnotesize (From PG \ipa{*aspō}).} f. aspen.
	

	
		\key{asphaltos.} {\footnotesize (From AG \gr{ἄσφᾰλτος}).} m.
			\textbf{1.} asphalt, bitumen; pitch, tar.
			\textbf{2.} asphalt concrete; material used to pave roads.
	

	
		\key{assót.} {\footnotesize (From Fr. \ipa{assaut}).} m. assault.
	

	
		\key{astatom.}  n.{\footnotesize \textit{ chemistry.}} astatine.
	

	
		\key{aster.} {\footnotesize (Variant of Dan. \ipa{ster}).} m. star.
	

	
		\key{astos.} {\footnotesize (From PG \ipa{*astaz}).} m. bush.
	

	
		\key{Athénai.} {\footnotesize (From Lat. \ipa{Athēnae}, from AG \gr{Ἀθῆναι}. $\blacklozenge$~This term is in plural).} f. Athens.
	

	
		\key{atifestríeti.} {\footnotesize (Related to Dan. \ipa{fesor}).} v. to shape.
	

	
		\key{atirama.} {\footnotesize (From Dan. \ipa{ati + rama}).} f. firmament (biblical).
	

	
		\key{atistati.} {\footnotesize (From Dan. \ipa{ati- + stati}).} v. to stop, to cease.
	

	
		\key{atjeria.} {\footnotesize (From Dan. \ipa{ati- + jeros + -ia}).} f. (in plural) eternity.
	

	
		\key{audia.} {\footnotesize (From PG \ipa{*awadī}).} f. eider.
	

	
		\key{audos.} {\footnotesize (From PG \ipa{*audaz}).} m. wealth, riches.
	

	
		\key{augíeti.} {\footnotesize (From PG \ipa{*augijaną}).} v. to show, to display.
	

	
		\key{augustos.} {\footnotesize (From Lat. \ipa{Augustus}).} m.{\footnotesize \textit{ month.}} August.
	

	
		\key{auria.} {\footnotesize (From PG \ipa{*aurī}).} f. a river bank, or sandbank at the river's mouth.
	

	
		\key{ause.} {\footnotesize (From Dan. \ipa{ausós}).} adv. tomorrow.
	

	\key{ausós,}
		\textit{uses.} {\footnotesize (From PIE \ipa{*h₂éwsōs}).} f. dawn.
	

	
		\key{Austina Germania.}  f. East Germany.
	

	
		\key{austinos.} {\footnotesize (From Dan. \ipa{austom}).} adj. eastern, east, easterly.
	

	
		\key{austom.} {\footnotesize (From PG \ipa{*austrą}).} n. east.
	

	
		\key{Australia.} {\footnotesize (From En. \ipa{Australia}).} f.{\footnotesize \textit{ continent.}} Australia.
	

	
		\key{Austria.} {\footnotesize (From Lat. \ipa{Austria}).} f. Austria.
	

	
		\key{automobilis.} {\footnotesize (From Fr. \ipa{automobile}, influenced by Lat. \ipa{mōbilis}).} f. automotive.
	

	
		\key{aventúra.} {\footnotesize (From Fr. \ipa{aventure}).} f.
			\textbf{1.} adventure.
			\textbf{2.} venture.
			\textbf{3.} (romantic) affair.
	

	
		\key{averón.} {\footnotesize (From PG \ipa{*aferô}).} m. offspring, descendant, successor.
	

	
		\key{avidiom.} {\footnotesize (From PG \ipa{*awidiją}).} n. a herd or flock of sheep.
	

	\key{avis,}
		\textit{uvies, vimos.} {\footnotesize (From PIE \ipa{*h₂éwis}).} f. bird.
	

	
		\key{avisjeti.} {\footnotesize (From PG \ipa{*abisjaną}).} v. to cut, shear.
	

	
		\key{avja.} {\footnotesize (From PG \ipa{*awjō}).} f.
			\textbf{1.} floodplain, meadow.
			\textbf{2.} island.
	

	
		\key{avlóti.} {\footnotesize (From PG \ipa{*aflōną}).} v. to acquire, to contrive.
	

	
		\key{avniom.} {\footnotesize (From PG \ipa{*abniją}).} n. material, stuff.
	

	
		\key{avón.} {\footnotesize (From PG \ipa{*abô}).} m. husband.
	

	
		\key{avundis.} {\footnotesize (From PG \ipa{*abundiz}).} f. envy.
	

	
		\key{avurdíeti.} {\footnotesize (From PG \ipa{*andawurdijaną}).} v. to answer, to respond.
	

	
		\key{Azerbaigján.} {\footnotesize (From Rus. \cyr{Азербайджа́н}).} m. Azerbaijan.
	

	
		\key{badgjet.} {\footnotesize (From En. \ipa{budget}).} m. budget.
	

	
		\key{badjom.} {\footnotesize (From PG \ipa{*badją}).} n. bed.
	

	
		\key{bados.} {\footnotesize (From PIE \ipa{*bʰeh₂ǵos}).} f. beech.
	

	
		\key{badóti.} {\footnotesize (From PG \ipa{*badōną}).} v. to press, to distress, to frighten.
	

	
		\key{badus.} {\footnotesize (From PIE \ipa{*bʰeh₂ǵʰús}).} m. arm.
	

	
		\key{badva.} {\footnotesize (From PG \ipa{*badwō}).} f. battle, fight.
	

	
		\key{bagasj.} {\footnotesize (From Fr. \ipa{bagage}).} m. baggage, luggage.
	

	
		\key{bagét.} {\footnotesize (From Fr. \ipa{baguette}).} f. baguette.
	

	
		\key{bainom.} {\footnotesize (From PG \ipa{*bainą}).} n. leg.
	

	
		\key{bairos.} {\footnotesize (From PG \ipa{*bairaz}).} m. boar.
	

	
		\key{baita.} {\footnotesize (From PG \ipa{*baitō}).} f.
			\textbf{1.} food.
			\textbf{2.} bait.
	

	
		\key{baitilos.} {\footnotesize (From PG \ipa{*baitilaz}).} m. chisel.
	

	
		\key{baitos.} {\footnotesize (From PG \ipa{*baitaz}).} m. boat, ship.
	

	
		\key{bakarios.} {\footnotesize (From PG \ipa{*bakārijaz}).} m. baker.
	

	
		\key{bakis.} {\footnotesize (From PG \ipa{*bakiz}).} f. beach.
	

	
		\key{Baku.} {\footnotesize (From Rus. \cyr{Баку́}).} n. Baku, capital of Azerbaijan.
	

	
		\key{balcón.} {\footnotesize (From Fr. \ipa{balcon}).} m. balcony.
	

	
		\key{baldia.} {\footnotesize (From PG \ipa{*balþį̄}).} f. boldness, strength, courage.
	

	
		\key{balgis.} {\footnotesize (From PG \ipa{*balgiz}).} m. bag.
	

	
		\key{baljo.} {\footnotesize (From PIE \ipa{*bʰāl-}).} v. to fail, to err, to miss.
	

	
		\key{balnia.} {\footnotesize (From Dan. \ipa{baljo}).} f. failure, error.
	

	
		\key{baltiskos.}  adj. pertaining to the Baltic region.
	

	
		\key{baltjos.} {\footnotesize (From PG \ipa{*baltijaz}).} m. belt.
	

	
		\key{balvom.} {\footnotesize (From PG \ipa{*balwą}).} n. evil, misfortune.
	

	
		\key{bandia.} {\footnotesize (From PG \ipa{*bandī}).} f. fetter, bond.
	

	
		\key{bankis.} {\footnotesize (From PG \ipa{*bankiz}).} f. bench.
	

	
		\key{bantis.} {\footnotesize (From PIE \ipa{*gʷʰn̥téys}).} f. striking, killing.
	

	
		\key{bantos.} {\footnotesize (From PIE \ipa{*gʷm̥tós}).} adj. passable.
	

	
		\key{baptisman.} {\footnotesize (From AG \gr{βάπτισμα}).} n. baptism.
	

	
		\key{barda.} {\footnotesize (From PIE \ipa{*bʰardʰéh₂}).} f. beard.
	

	
		\key{bardátos.} {\footnotesize (From PIE \ipa{*bʰardʰéh₂tos}).} adj. bearded.
	

	
		\key{bariom.}  n.{\footnotesize \textit{ chemistry.}} barium.
	

	
		\key{bataljón.} {\footnotesize (From Fr. \ipa{bataillon}).} m. batallion.
	

	
		\key{be.}  n. letter \textit{b}.
	

	
		\key{befrus.} {\footnotesize (From PIE \ipa{*bʰébʰrus}).} m. beaver.
	

	
		\key{Belarossia.}  f. Belarus.
	

	
		\key{Belgia.} {\footnotesize (From Lat. \ipa{Belgia}).} f. Belgium.
	

	
		\key{Belgrádom.}  n. Belgrad, capital of Serbia.
	

	
		\key{belia.} {\footnotesize (From PIE \ipa{*bel-}).} f.
			\textbf{1.} strength.
			\textbf{2.} (physics) force.
	

	
		\key{bena.} {\footnotesize (From PIE \ipa{*gʷḗn}).} f. woman.
	

	
		\key{benfus.} {\footnotesize (From PIE \ipa{*bʰénǵʰus}).} adj. dense, thick.
	

	
		\key{Benjamín.} {\footnotesize (From AG \gr{Βενιᾱμῑ́ν}).} m. Benjamin.
	

	\key{bentis,}
		\textit{banties.} {\footnotesize (From PIE \ipa{*gʷémtis}).} f. step, act of walking.
	

	\key{berdonts,}
		\textit{berdantes.} {\footnotesize (From PIE \ipa{*bʰérǵʰonts}).} adj. high, tall.
	

	
		\key{berdos.} {\footnotesize (From PIE \ipa{*bʰerHǵós}).} m.
			\textbf{1.} birch.
			\textbf{2.} bark (of a tree).
	

	
		\key{bereti.} {\footnotesize (From PIE \ipa{*bʰéreti}).} v. to carry.
	

	
		\key{berkeliom.}  n.{\footnotesize \textit{ chemistry.}} berkelium.
	

	
		\key{Berlín.}  m. Berlin.
	

	
		\key{beros.} {\footnotesize (From PIE \ipa{*gʷʰéros}).} n.
			\textbf{1.} heat.
			\textbf{2.} warm weather.
	

	\key{bertis,}
		\textit{barties.} {\footnotesize (From PIE \ipa{*bʰértis}).} f. the act of carrying, bearing.
	

	\key{beru,}
		\textit{breves, brumos.} {\footnotesize (From PIE \ipa{*gʷéru}).} n. spear.
	

	
		\key{berylliom.}  n.{\footnotesize \textit{ chemistry.}} beryllium.
	

	
		\key{beta.}  f. Greek letter beta.
	

	
		\key{bibliom.} {\footnotesize (From AG \gr{βῐβλῐ́ον}).} n. book.
	

	\key{bideti¹,}
		\textit{befideti.} {\footnotesize (From PIE \ipa{*bʰeyd-}).} v.
			\textbf{1.} to split.
			\textbf{2.} to separate, to divide.
	

	\key{bideti²,}
		\textit{befideti.} {\footnotesize (From PIE \ipa{*bʰeydʰ-}).} v. to trust.
	

	
		\key{bidinos.} {\footnotesize (From Dan. \ipa{bideti²}).} adj. believable, credible.
	

	
		\key{biftos.} {\footnotesize (From PIE \ipa{*bʰidʰtós}).} adj. trusted, trustworthy.
	

	
		\key{biologia.} {\footnotesize (From Lat. \ipa{biologia}).} f. biology.
	

	
		\key{bismútom.}  n.{\footnotesize \textit{ chemistry.}} bismuth.
	

	
		\key{bistos.} {\footnotesize (From PIE \ipa{*bʰidtós}).} adj. split, cloven.
	

	
		\key{biva.} {\footnotesize (From PIE \ipa{*gʷih₃wós}).} f. life.
	

	
		\key{biveter.} {\footnotesize (From Dan. \ipa{biveti}).} m. resident, inhabitant.
	

	
		\key{biveti.} {\footnotesize (From Dan. \ipa{biva}).} v. to live.
	

	
		\key{bivetis.} {\footnotesize (From Dan. \ipa{biveti}).} f. residence, habitat.
	

	
		\key{bivos.} {\footnotesize (From PIE \ipa{*gʷih₃wós}).} adj. alive, living.
	

	
		\key{blesnos.} {\footnotesize (From PIE \ipa{*bʰleǵ-}).} m. bait, lure.
	

	
		\key{blevos.} {\footnotesize (From PG \ipa{*blēwaz}).} adj. blue.
	

	
		\key{bloman.} {\footnotesize (From Dan. \ipa{bloti}).} n. flower.
	

	
		\key{blomenikom.} {\footnotesize (From Dan. \ipa{bloman}).} n. little flower.
	

	
		\key{blomia.} {\footnotesize (From Dan. \ipa{bloti}).} f. freshness.
	

	
		\key{bloti.} {\footnotesize (From PIE \ipa{*bʰleh₃-}).} v.
			\textbf{1.} to grow (of plants).
			\textbf{2.} to sprout, to germinate.
			\textbf{3.} to bloom, to flower.
	

	
		\key{bogeti.} {\footnotesize (From PIE \ipa{*gʷog-}).} v.
			\textbf{1.} to shake.
			\textbf{2.} to agitate.
	

	
		\key{boida.} {\footnotesize (From PIE \ipa{*bʰoydʰ-}).} f.
			\textbf{1.} trust.
			\textbf{2.} faith.
	

	
		\key{bolos.} {\footnotesize (From PIE \ipa{*bel-}).} adj. strong, mighty.
	

	
		\key{bomba.} {\footnotesize (From It. \ipa{bomba}).} f. bomb.
	

	\key{bor,}
		\textit{beres, barmos.} {\footnotesize (From PIE \ipa{*bʰṓr}).} m. thief.
	

	
		\key{bora.} {\footnotesize (From PIE \ipa{*gʷorH-eh₂}).} f. mountain; hill.
	

	
		\key{boriom.}  n.{\footnotesize \textit{ chemistry.}} bohrium.
	

	
		\key{borom.}  n.{\footnotesize \textit{ chemistry.}} boron.
	

	
		\key{Bosnia i Hertsegovína.}  f. Bosnia and Herzegovina.
	

	
		\key{bosos.} {\footnotesize (From PIE \ipa{*bʰosós}).} adj. bare, barefoot.
	

	
		\key{botíca.} {\footnotesize (From Fr. \ipa{boutique}, from Old Occ. \ipa{botica}).} f. small shop; boutique.
	

	
		\key{botón.} {\footnotesize (From Fr. \ipa{bouton}, from OF \ipa{boton}).} m.
			\textbf{1.} button.
			\textbf{2.} (flower) bud.
			\textbf{3.} pimple, spot, zit.
	

	\key{bovs,}
		\textit{beves, bumos.} {\footnotesize (From PIE \ipa{*gʷṓws}).} m. cattle.
	

	
		\key{Braján.} {\footnotesize (From En. \ipa{Brian}).} m. Brian, Bryan.
	

	
		\key{brantos.} {\footnotesize (Related to ON \ipa{brant}).} adj. steep.
	

	
		\key{Brasil.}  m. Brazil.
	

	
		\key{brasmos.} {\footnotesize (From PIE \ipa{*semh₂-}).} m. summer.
	

	
		\key{brater.} {\footnotesize (From PIE \ipa{*bʰréh₂tēr}).} m.
			\textbf{1.} brother.
			\textbf{2.} Gemini (in the plural, constellation).
	

	
		\key{Bratisláva.} {\footnotesize (From Rus. \ipa{Братисла́ва}).} f. Bratislava, capital of Slovakia.
	

	\key{bravón,}
		\textit{brunes, bravanmos.} {\footnotesize (From PIE \ipa{*gʷréh₂wō}).} m. quern, millstone.
	

	
		\key{bredíeti.} {\footnotesize (From PIE \ipa{*bʰredʰ-}).} v. to roam, to wander freely.
	

	\key{brevan,}
		\textit{bravens, brunamos.} {\footnotesize (From PIE \ipa{*bʰréh₁wr̥}).} n. well, spring, source, fountain.
	

	
		\key{brigáda.} {\footnotesize (From Fr. \ipa{brigade}).} f. brigade.
	

	
		\key{Britannia.}  f. Britain.
	

	
		\key{Britanniskom Imperiom.} {\footnotesize (From Dan. \ipa{britanniskom + imperiom}).} n. British Empire.
	

	
		\key{britanniskos.} {\footnotesize (From Dan. \ipa{Britannia}).} adj. British.
	

	
		\key{briva.} {\footnotesize (From PIE \ipa{*gʷriHwéh₂}).} f.
			\textbf{1.} mane, nape.
			\textbf{2.} neck, thoat.
			\textbf{3.} estuary, mouth of an inlet.
	

	
		\key{bromom.}  n.{\footnotesize \textit{ chemistry.}} bromine.
	

	
		\key{brunos.} {\footnotesize (From PIE \ipa{*bʰruHnós}).} adj. brown.
	

	\key{brus¹,}
		\textit{bruves, brumos.} {\footnotesize (From PIE \ipa{*gʷr̥h₂ús}).} adj.
			\textbf{1.} heavy.
			\textbf{2.} difficult, hard.
			\textbf{3.} severe, grave.
	

	\key{brus²,}
		\textit{bruves, brumos.} {\footnotesize (From PIE \ipa{*h₃bʰrúHs}).} f. eyebrow.
	

	
		\key{Brussellai.} {\footnotesize (From Fr. \ipa{Bruxelles}, with a latinized ending).} f. Brussels, capital of Belgium.
	

	
		\key{Bucuresjti.} {\footnotesize (From Rom. \ipa{București}).} n. Bucharest, capital of Romania.
	

	\key{bufmén,}
		\textit{bufnes, budanmos.} {\footnotesize (From PIE \ipa{*bʰudʰmḗn}).} m. bottom.
	

	
		\key{buftos.} {\footnotesize (From PIE \ipa{*bʰudʰtós}).} adj. aware.
	

	
		\key{buga.} {\footnotesize (From PIE \ipa{*bʰugéh₂}).} f. flight, escape.
	

	
		\key{bujeti.}  v.
			\textbf{1.} (+\ipa{en} + accusative; active) to turn something into.
			\textbf{2.} (+\ipa{en} + accusative; passive) to become, to turn into.
	

	
		\key{Bulgaria.}  f. Bulgaria.
	

	
		\key{buljón.} {\footnotesize (From Fr. \ipa{bouillon}).} m. broth (water in which food has been boiled).
	

	
		\key{cadmiom.}  n.{\footnotesize \textit{ chemistry.}} cadmium.
	

	
		\key{caisiom.}  n.{\footnotesize \textit{ chemistry.}} caesium, cesium.
	

	
		\key{calciom.}  n.{\footnotesize \textit{ chemistry.}} calcium.
	

	
		\key{calculos.} {\footnotesize (From Lat. \ipa{calculus}).} m. calculus.
	

	
		\key{californiom.}  n.{\footnotesize \textit{ chemistry.}} californium.
	

	
		\key{Cambria.} {\footnotesize (From Lat. \ipa{Cambria}).} f. Wales.
	

	
		\key{Canada.}  f. Canada.
	

	
		\key{capitalismos.}  m. capitalism.
	

	
		\key{ce.}  n. letter \textit{c}.
	

	
		\key{ceriom.}  n.{\footnotesize \textit{ chemistry.}} cerium.
	

	
		\key{cetos.}  m.
			\textbf{1.} sea monster.
			\textbf{2.} whale; cetacean.
	

	
		\key{chemia.} {\footnotesize (From Ger. \ipa{Chemie}).} f. chemistry.
	

	
		\key{chi.}  n. Greek letter chi.
	

	
		\key{chlorom.}  n.{\footnotesize \textit{ chemistry.}} chlorine.
	

	
		\key{christiános.} {\footnotesize (From Lat. \ipa{Chrīstiānus}, from AG \gr{Χρῑστιᾱνός}).} adj. Christian.
	

	
		\key{Christophoros.} {\footnotesize (From AG \gr{Χρῑστόφορος}).} m. Christopher.
	

	
		\key{Christos.} {\footnotesize (From Lat. \ipa{Christus}, from AG \gr{Χριστός}).} m. Christ, Messiah.
	

	
		\key{chromiom.}  n.{\footnotesize \textit{ chemistry.}} chromium.
	

	
		\key{Cjechia.} {\footnotesize (From Pol. \ipa{Czech}).} m. Czech Republic, Czechia.
	

	
		\key{classiska mechanica.}  f. classical mechanics.
	

	
		\key{climan.} {\footnotesize (From Lat. \ipa{clima}, from AG \gr{κλίμα}).} n.
			\textbf{1.} climate.
			\textbf{2.}{\footnotesize \textit{ mathematics.}} slope, inclination.
	

	
		\key{cobaltiom.}  n.{\footnotesize \textit{ chemistry.}} cobalt.
	

	
		\key{Columbia.}  f. Colombia.
	

	
		\key{commandier.} {\footnotesize (From Fr. \ipa{commandeur}).} m. commander.
	

	
		\key{communismos.}  m. communism.
	

	
		\key{communísta.} {\footnotesize (From Dan. \ipa{communismos}).} adj. communist.
	

	
		\key{companjia.} {\footnotesize (From Fr. \ipa{compagnie}).} f.
			\textbf{1.} a team; a group of people who work together professionally.
			\textbf{2.} military company.
			\textbf{3.} (business) company.
	

	
		\key{computa.} {\footnotesize (From Fr. \ipa{compte}, influenced by Lat. \ipa{computare}).} f.
			\textbf{1.} account (bank or user account).
			\textbf{2.} count (the action of counting).
	

	
		\key{conditión.}  f. condition.
	

	
		\key{Condrados.} {\footnotesize (Related to Ger. \ipa{Konrad}).} m. Conrad, Konrad.
	

	
		\key{conscientia.} {\footnotesize (From Lat. \ipa{cōnscientia}).} f. conscience.
	

	
		\key{Constantinopolis.} {\footnotesize (From AG \gr{Κωνσταντινούπολις}).} f. Constantinople; Istanbul.
	

	
		\key{constitutión.} {\footnotesize (From Lat. \ipa{cōnstitūtiō}).} f. constitution.
	

	
		\key{coperniciom.}  n.{\footnotesize \textit{ chemistry.}} copernicium.
	

	
		\key{Coréa.}  f. Korea.
	

	
		\key{Croatia.} {\footnotesize (From Lat. \ipa{Croātia}).} f. Croatia.
	

	
		\key{Cuba.} {\footnotesize (From Sp. \ipa{Cuba}).} f. Cuba.
	

	
		\key{curier.} {\footnotesize (From Fr. \ipa{courrier}).} m.
			\textbf{1.} mail.
			\textbf{2.} post office.
			\textbf{3.} mail postal service.
	

	
		\key{curiom.}  n.{\footnotesize \textit{ chemistry.}} curium.
	

	
		\key{Cypros.} {\footnotesize (From Latin \ipa{Cyprus}, from AG \gr{Κύπρος}).} f. Cyprus.
	

	
		\key{da.}  conj. and.
	

	
		\key{daitis.} {\footnotesize (From PIE \ipa{*déh₂itis}).} f. time.
	

	
		\key{dakru.} {\footnotesize (From PIE \ipa{*dáḱru}).} n. tear (in the eyes).
	

	
		\key{dalgos.} {\footnotesize (From PIE \ipa{*dl̥h₁gʰós}).} adj.
			\textbf{1.} long.
			\textbf{2.} tall.
			\textbf{3.} far, far away.
	

	
		\key{dambros.} {\footnotesize (From PIE \ipa{*ǵm̥-rós}).} m. son in law.
	

	
		\key{damis.} {\footnotesize (From PIE \ipa{*ǵm̥-is}).} f. marriage, matrimony.
	

	
		\key{Daneta.} {\footnotesize (From Dan. \ipa{danu}).} f. Danetian river goddess, freed from the clutches of the water serpent.
	

	
		\key{danetiska.} {\footnotesize (From Dan. \ipa{danetiskos}).} f. Danetian language.
	

	
		\key{danetiskos.} {\footnotesize (From Dan. \ipa{Daneta}).} adj. Danetian.
	

	
		\key{Dania.}  f. Denmark.
	

	
		\key{Daniel.} {\footnotesize (From AG \gr{Δᾱνῑήλ}).} m. Daniel.
	

	
		\key{daniska.} {\footnotesize (From Dan. \ipa{Dania}).} f. Danish language.
	

	
		\key{danu.} {\footnotesize (From PIE \ipa{*déh₂nu}).} n. river.
	

	
		\key{danva.} {\footnotesize (From PIE \ipa{*dn̥ǵʰwéh₂s}).} f.
			\textbf{1.} tongue.
			\textbf{2.} language.
	

	
		\key{Daramstat.} {\footnotesize (From Ger. \ipa{Darmstadt}).} m. Darmstadt.
	

	
		\key{daramstatiom.}  n.{\footnotesize \textit{ chemistry.}} darmstadtium.
	

	
		\key{dat.} {\footnotesize (From Dan. \ipa{datos}).} conj. if.
	

	
		\key{datos.} {\footnotesize (From PIE \ipa{*dh₃tós}).} adj. given.
	

	
		\key{David.} {\footnotesize (From AG \gr{Δαυῑ́δ}).} m. David.
	

	
		\key{de.}  n. letter \textit{d}.
	

	
		\key{december.} {\footnotesize (From Lat. \ipa{December}).} m.{\footnotesize \textit{ month.}} December.
	

	
		\key{dégar.} {\footnotesize (Possibly cognate with AG \gr{γᾰ́ρ}).} conj. for, since.
	

	
		\key{dege.} {\footnotesize (From PIE \ipa{*de + *gʰe}).} adv.
			\textbf{1.} even.{\footnotesize \ipa{Dege dorú bivonta}. \textit{Even trees are alive}.}
			\textbf{2.} indeed, of course, in fact.
	

	
		\key{delta.}  f. Greek letter delta.
	

	
		\key{delus.} {\footnotesize (From PIE \ipa{*ǵelu-}).} m. leech.
	

	
		\key{Demetrios.} {\footnotesize (From AG \gr{Δημήτριος}).} m. Demetrius, Dimitri, Dmitri.
	

	\key{denaman,}
		\textit{danamens.} {\footnotesize (From PIE \ipa{*ǵénh₁mn̥}).} n.
			\textbf{1.} offspring, seed.
			\textbf{2.} (genealogical) generation.
	

	
		\key{denati.} {\footnotesize (From PIE \ipa{*ǵenh₁-}).} v.
			\textbf{1.} to produce, to create.
			\textbf{2.} to give birth.
			\textbf{3.} to be born (when used in the mediopassive voice).
	

	\key{denatis,}
		\textit{danaties.} {\footnotesize (From PIE \ipa{*ǵénh₁tis}).} f.
			\textbf{1.} birth.
			\textbf{2.} production.
	

	\key{denator,}
		\textit{danatres.} {\footnotesize (From PIE \ipa{*ǵénh₁tōr}).} m.
			\textbf{1.} one who begets.
			\textbf{2.} one who produces.
			\textbf{3.} parent.
	

	
		\key{denos.} {\footnotesize (From PIE \ipa{*ǵénh₁os}).} n. race, lineage.
	

	
		\key{denspotis.} {\footnotesize (From PIE \ipa{*déms pótis}).} m. master of the house.
	

	\key{denus,}
		\textit{danves, danumos.} {\footnotesize (From PIE \ipa{*ǵénus}).} f.
			\textbf{1.} cheek.
			\textbf{2.} jaw.
			\textbf{3.} chin.
	

	
		\key{dera.} {\footnotesize (From PIE \ipa{*ǵērh₂-s}).} f. old age.
	

	
		\key{deronts.} {\footnotesize (From PIE \ipa{*ǵérh₂onts}).} adj. old.
	

	
		\key{desam.} {\footnotesize (From PIE \ipa{*déḱm̥}).} num. ten.
	

	
		\key{desamos.}  adj. tenth.
	

	
		\key{desos.} {\footnotesize (From PIE \ipa{*déḱos}).} n. that which is proper.
	

	
		\key{dictionariom.} {\footnotesize (From Lat. \ipa{dictiōnārium}).} n. dictionary.
	

	
		\key{din.} {\footnotesize (From PIE {*dyew-}).} m. day.
	

	
		\key{din Jovies.}  m. Thursday.
	

	
		\key{din Lunas.}  m. Monday.
	

	
		\key{din Martes.}  m. Tuesday.
	

	
		\key{din Mercuriosjo.}  m. Wednesday.
	

	
		\key{din Saturnosjo.}  m. Saturday.
	

	
		\key{din Soles.}  m. Sunday.
	

	
		\key{din Venesos.}  m. Friday.
	

	
		\key{dinosauros.} {\footnotesize (From Fr. \ipa{dinosaure}, with a latinized ending).} m. dinosaur.
	

	
		\key{districtos.} {\footnotesize (From Lat. \ipa{districtus}).} m. district.
	

	
		\key{división.}  f. division.
	

	
		\key{divjos.} {\footnotesize (From PIE \ipa{*diwyós}).} adj. heavenly.
	

	
		\key{divos.} {\footnotesize (From PIE \ipa{*deywós}).} m. god.
	

	
		\key{djolet.} {\footnotesize (From a Turkic language; compare Uyghur \ipa{dölet} and Uzbek \ipa{davlat}).} m. (political) state.
	

	
		\key{djoletiskos.} {\footnotesize (From Dan. \ipa{djolet}).} adj. pertaining to a political state or sovereign polity.
	

	\key{djus,}
		\textit{dives, djumos.} {\footnotesize (From PIE \ipa{*dyḗws}).} m. sky, heaven.
	

	
		\key{Djus Pater.}  m. Sky Father, supreme Danetian deity.
	

	
		\key{do.} {\footnotesize (From PIE \ipa{*do}).} prep. (+gen) up to, till, until.
	

	
		\key{doctor.} {\footnotesize (From Lat. \ipa{doctor}).} m. doctor.
	

	\key{dom,}
		\textit{dems.} {\footnotesize (From PIE \ipa{*dṓm}).} m. house, home.
	

	
		\key{Donald.} {\footnotesize (From En. \ipa{Donald}).} m. Donald.
	

	
		\key{Donald Tramp.}  m. Donald Trump.
	

	
		\key{donfíeti.} {\footnotesize (From Dan. \ipa{donvos}).} v. to count.
	

	
		\key{donia.} {\footnotesize (From Dan. \ipa{denatis}).} f. motherland.
	

	
		\key{donom.} {\footnotesize (From PIE \ipa{*déh₃nom}).} n. gift.
	

	
		\key{donos.} {\footnotesize (From PIE \ipa{*ǵónh₁os}).} m. offspring.
	

	\key{donts,}
		\textit{dantes.} {\footnotesize (From PIE \ipa{*h₃dónts}).} m. tooth.
	

	\key{donu,}
		\textit{danves, danumos.} {\footnotesize (From PIE \ipa{*ǵónu}).} n. knee.
	

	
		\key{donvos.} {\footnotesize (From PIE \ipa{*ǵómbʰos}).} m. number.
	

	
		\key{doros.} {\footnotesize (From PIE \ipa{*ǵorh₂-o-s}).} adj. free (having freedom).
	

	\key{doru,}
		\textit{dreves, drumos.} {\footnotesize (From PIE \ipa{*dóru}).} n. tree.
	

	\key{dotis,}
		\textit{daties.} {\footnotesize (From PIE \ipa{*déh₃tis}).} f. act of giving.
	

	\key{dotor,}
		\textit{datres.} {\footnotesize (From PIE \ipa{*déh₃tōr}).} m. giver.
	

	
		\key{drageti.} {\footnotesize (From PG \ipa{*draganą}).} v.
			\textbf{1.} to drag.
			\textbf{2.} (in the passive) to drag oneself, to crawl.
	

	
		\key{drajus.} {\footnotesize (From PIE \ipa{*ǵr̥h₂-yéw-s}).} f. old woman.
	

	
		\key{dranom.} {\footnotesize (From PIE \ipa{*ǵr̥h₂nóm}).} n. grain.
	

	
		\key{dublevé.}  n. letter \textit{w}.
	

	
		\key{Dublín.} {\footnotesize (From Fr. \ipa{Dublin}).} m. Dublin, capital of Ireland.
	

	
		\key{dubniom.}  n.{\footnotesize \textit{ chemistry.}} dubnium.
	

	
		\key{dukti.} {\footnotesize (From PIE \ipa{*dewk-ti}).} v.
			\textbf{1.} to lead.
			\textbf{2.} to pull, to draw.
	

	
		\key{dulgos.} {\footnotesize (From PG \ipa{*dulgaz}).} m. debt.
	

	
		\key{dus-.} {\footnotesize (From PIE \ipa{*dus-}).} pref. bad.
	

	
		\key{dusdjus.} {\footnotesize (From Dan. \ipa{dus- + -djus}).} m. bad sky, stormy sky, stormy weather.
	

	
		\key{duseti.} {\footnotesize (From PIE \ipa{*ǵéwseti}).} v. to taste, to try, to choose.
	

	
		\key{dusmenés.} {\footnotesize (From PIE \ipa{*dusmenḗs}).} adj. ill-disposed, wicked, evil.
	

	
		\key{dusmenistos.} {\footnotesize (From Dan. \ipa{dusmenés}).} adj. most wicked, most evil.
	

	
		\key{dustus.} {\footnotesize (From PIE \ipa{*ǵéwstus}).} m. taste.
	

	
		\key{dveleti.} {\footnotesize (From PIE \ipa{*ǵwélH-e-ti}).} v. to burn, to shine.
	

	
		\key{dvibrus.}  m.
			\textbf{1.} balance, scale.
			\textbf{2.} Libra (constellation).
	

	
		\key{dvis.} {\footnotesize (From PIE \ipa{*dwís}).} adv.
			\textbf{1.} twice, doubly.
			\textbf{2.} in two.
	

	
		\key{dvo.} {\footnotesize (From PIE \ipa{*dwóh₁}).} num. two.
	

	
		\key{dvoidesam.} {\footnotesize (From PIE \ipa{*dwóh₁ + *déḱm̥}).} num. twelve.
	

	
		\key{dvoidesamos.}  adj. twelfth.
	

	
		\key{dysprosiom.}  n.{\footnotesize \textit{ chemistry.}} dysprosium.
	

	
		\key{e.}  n. letter \textit{e}.
	

	
		\key{ecjíeti.} {\footnotesize (From PG \ipa{*aikijaną}).} v. to irritate, to annoy, to bother.
	

	
		\key{Edém.} {\footnotesize (From Gr. \gr{Εδέμ}).} m. Eden, Garden of Eden.
	

	
		\key{Edvardos.} {\footnotesize (Ultimately from PG \ipa{*Audawarduz}).} m. Edward, Eduardo.
	

	
		\key{ef.}  n. letter \textit{f}.
	

	\key{eks,}
		\textit{ekes.} {\footnotesize (From PG \ipa{*aiks}).} f.
			\textbf{1.} oak tree.
			\textbf{2.} oak wood.
	

	
		\key{ekter.} {\footnotesize (From PG \ipa{*aihtēr}).} m. owner, possessor.
	

	
		\key{ektis.} {\footnotesize (From PG \ipa{*aihtiz}).} f. property, possessions.
	

	
		\key{ektróti.} {\footnotesize (From PG \ipa{*aihtrōną}).} v. to beg.
	

	
		\key{ekverna.} {\footnotesize (From PG \ipa{*aikwernô}).} m. squirrel.
	

	
		\key{el.}  n. letter \textit{l}.
	

	
		\key{electión.} {\footnotesize (From Lat. \ipa{ēlēctiō}).} f.
			\textbf{1.} a choice, selection.
			\textbf{2.} (politics) election.
	

	
		\key{electroniskos.} {\footnotesize (From Fr. \ipa{électronique}, with a Danetian ending).} adj. electronic.
	

	
		\key{electroniskos curier.} {\footnotesize (From Dan. \ipa{electroniskos + curier}).} m. email, e-mail, electronic mail.
	

	
		\key{eleméntom.} {\footnotesize (From Lat. \ipa{elementum}).} n.
			\textbf{1.} element.
			\textbf{2.} chemical element.
			\textbf{3.} basic principle, rudiment.
	

	
		\key{Elia.} {\footnotesize (From AG \gr{Ἠλίᾱς}).} m. Elijah.
	

	
		\key{eligeti.} {\footnotesize (From Lat. \ipa{ēligō}).} v.
			\textbf{1.} to choose something; to make a decision.
			\textbf{2.} (politics) to elect.
	

	
		\key{ellípsis.} {\footnotesize (From Lat. \ipa{ellīpsis}, from AG \gr{ἔλλειψῐς}).} f.{\footnotesize \textit{ mathematics.}} ellipse.
	

	
		\key{em.}  n. letter \textit{m}.
	

	
		\key{en¹.}  prep.
			\textbf{1.} (+loc) in.
			\textbf{2.} (+acc) into.
	

	
		\key{en².}  n. letter \textit{n}.
	

	
		\key{encyclopaidia.} {\footnotesize (From New Latin \ipa{encyclopaedīa}).} f. encyclopedia.
	

	
		\key{energia.} {\footnotesize (From Lat. \ipa{energia}, from AG \gr{ἐνέργειᾰ}).} f. energy (in physics, and in the colloquial sense).
	

	
		\key{eno.} {\footnotesize (From PIE \ipa{*h₁énos}).} pron. third person pronoun.
	

	
		\key{ensembla.} {\footnotesize (From Fr. \ipa{ensemble}).} f.
			\textbf{1.}{\footnotesize \textit{ mathematics.}} a set.
			\textbf{2.}{\footnotesize \textit{ music.}} an ensemble.
	

	
		\key{enstangeti.} {\footnotesize (From Dan. \ipa{en- + stanga + -eti}).} v. to erect, to raise, to hoist.
	

	
		\key{enteros.} {\footnotesize (From PIE \ipa{*h₁énteros}).} adj.
			\textbf{1.} inside, within.
			\textbf{2.} inner, located inside.
	

	
		\key{enu.} {\footnotesize (From PIE \ipa{*h₁(n̥)new}).} prep. (+gen) without.
	

	
		\key{enurejos.} {\footnotesize (From PIE \ipa{*h₁(n̥)new + *reh₁is}).} adj. poor, wantsome.
	

	
		\key{envigjom.} {\footnotesize (From PG \ipa{*ainawīgiją}).} n. single combat.
	

	
		\key{Ephraím.} {\footnotesize (From AG \gr{Ἐφραΐμ}).} m. Ephraim.
	

	
		\key{epsilón.}  f. Greek letter epsilon.
	

	
		\key{er.}  n. letter \textit{r}.
	

	
		\key{erbiom.}  n.{\footnotesize \textit{ chemistry.}} erbium.
	

	
		\key{erdagos.} {\footnotesize (From PG \ipa{*airidagaz}).} m.
			\textbf{1.} days of yore; bygone days.
			\textbf{2.} the past.
	

	
		\key{Erik.} {\footnotesize (Ultimately from PG \ipa{*Aizarīkijaz}).} m. Eric, Erik.
	

	
		\key{Ernéstos.} {\footnotesize (From Ger. \ipa{Ernst}, influenced by Sp. \ipa{Ernesto}).} m. Ernesto, Ernst, Ernest.
	

	
		\key{Ernéstos Raderford.}  m. Ernest Rutherford.
	

	
		\key{erstidis.} {\footnotesize (From PG \ipa{*airi + *tīdiz}).} f. season.
	

	
		\key{es.}  n. letter \textit{s}.
	

	\key{esan,}
		\textit{sens, sanmos.} {\footnotesize (From PIE \ipa{*h₁ésh₂r̥}).} n. (flowing) blood.
	

	
		\key{esclerier.} {\footnotesize (From Fr. \ipa{éclaireur}, with ending analogous to Dan. \ipa{ingjenier}).} m. (military) scout.
	

	
		\key{espe.} {\footnotesize (From PIE \ipa{*h₁éskʷe}).} adv. yet, still.
	

	
		\key{esti¹.} {\footnotesize (From PIE \ipa{*h₁ésti}).} v. to be.
	

	
		\key{esti².} {\footnotesize (From PIE \ipa{*h₁édti}).} v. to eat.
	

	
		\key{estom.} {\footnotesize (From PG \ipa{*aistô}).} n. testicle.
	

	
		\key{Estonia.}  f. Estonia.
	

	
		\key{estos.} {\footnotesize (From PG \ipa{*aistaz}).} m. kiln.
	

	
		\key{esvos.} {\footnotesize (From PIE \ipa{*h₁éḱwos}).} m. horse, stallion.
	

	
		\key{Európa.} {\footnotesize (From AG \gr{Εὐρώπη}).} f. Europe.
	

	
		\key{europiom.}  n.{\footnotesize \textit{ chemistry.}} europium.
	

	
		\key{Europiskom Somjugos.} {\footnotesize (From Dan. \ipa{europiskom + somjugos}).} n. European Union.
	

	
		\key{europiskos.} {\footnotesize (From Dan. \ipa{Európa}).} adj. European.
	

	\key{famón,}
		\textit{famnes, famanmos.} {\footnotesize (From PIE \ipa{*ǵʰm̥nés}).} m. earthling.
	

	
		\key{fans.} {\footnotesize (From PIE \ipa{*ǵʰh₂éns}).} f. goose.
	

	
		\key{fascismos.}  m. fascism.
	

	
		\key{fastos.} {\footnotesize (From PIE \ipa{*ǵʰasto-}).} m. stick.
	

	
		\key{febeti.} {\footnotesize (From PIE \ipa{*bʰegʷ-}).} v.
			\textbf{1.} to run.
			\textbf{2.} to flee.
	

	
		\key{februarios.} {\footnotesize (From Lat. \ipa{Februārius}).} m.{\footnotesize \textit{ month.}} February.
	

	\key{fedóm,}
		\textit{fames.} {\footnotesize (From PIE \ipa{*dʰéǵʰōm}).} f.
			\textbf{1.} earth.
			\textbf{2.} land.
			\textbf{3.} country.
	

	
		\key{Fedóm Mater.}  f. Earth Mother.
	

	\key{fejóm,}
		\textit{fimes.} {\footnotesize (From PIE \ipa{*ǵʰéyōm}).} f. winter.
	

	
		\key{fekti.} {\footnotesize (From PIE \ipa{*dʰeh₁-}).} v.
			\textbf{1.} to make, to do.
			\textbf{2.} to put, to place.
	

	
		\key{feljeti.} {\footnotesize (From PIE \ipa{*ǵʰélh₃-ye-ti}).} v. to flourish.
	

	\key{feman,}
		\textit{famens.} {\footnotesize (From PIE \ipa{*dʰéh₁mn̥}).} n.
			\textbf{1.} abode, dwelling.
			\textbf{2.} state, order.
			\textbf{3.} manner, mode, form.
	

	
		\key{fembeti.} {\footnotesize (From PIE \ipa{*gʷʰemb-}).} v. to spring, hop, jump.
	

	
		\key{fer.} {\footnotesize (From PIE \ipa{*ǵʰḗr}).} m. hedgehog.
	

	
		\key{fermiom.}  n.{\footnotesize \textit{ chemistry.}} fermium.
	

	
		\key{fesor.} {\footnotesize (From PIE \ipa{*ǵʰésōr}).} f. hand.
	

	\key{fetis,}
		\textit{faties.} {\footnotesize (From PIE \ipa{*dʰéh₁tis}).} f.
			\textbf{1.} act of putting, placement.
			\textbf{2.} act of doing.
			\textbf{3.} heroic deed, feat.
	

	
		\key{Finnia.} {\footnotesize (From Lat. \ipa{Finnia}).} f. Finland.
	

	
		\key{finniska.} {\footnotesize (From Dan. \ipa{Finnia}).} f. Finnish language.
	

	
		\key{fitis.} {\footnotesize (From PIE \ipa{*dʰgʷʰéytis}).} f. decay, destruction.
	

	
		\key{fjena.} {\footnotesize (From PIE \ipa{*dʰeh₁(y)-}).} f. young woman; virgin, maiden.
	

	
		\key{fjesi.} {\footnotesize (From PIE \ipa{*dʰǵʰyési}).} adv. yesterday.
	

	
		\key{fjesteros.} {\footnotesize (From PIE \ipa{*dʰǵʰyésteros}).} adj. pertaining to yesterday.
	

	
		\key{flaros.} {\footnotesize (From PIE \ipa{*ǵʰl̥h₃-ró-s}).} v. yellow.
	

	
		\key{fleroviom.}  n.{\footnotesize \textit{ chemistry.}} flerovium.
	

	
		\key{Florida.} {\footnotesize (From En. \ipa{Florida}).} f. Florida.
	

	
		\key{flugeti.} {\footnotesize (From PG \ipa{*fleugō}).} v. to fly.
	

	
		\key{fluorom.}  n.{\footnotesize \textit{ chemistry.}} fluorine.
	

	
		\key{fodón.} {\footnotesize (From PG \ipa{*fōdô}).} m. food, nourishment.
	

	
		\key{foideratión.} {\footnotesize (From Lat. \ipa{foederatio}).} f. federation.
	

	
		\key{fomos.} {\footnotesize (From PIE \ipa{*dʰóh₁mos}).} m.
			\textbf{1.} that which is put, placed.
			\textbf{2.} judgement.
	

	
		\key{fona.} {\footnotesize (From PIE \ipa{*dʰoHnéh₂}).} f. bread.
	

	
		\key{fortos.} {\footnotesize (From PIE \ipa{*ǵʰórtos}).} m. enclosure.
	

	
		\key{frags.} {\footnotesize (From PIE \ipa{*dʰrā́ks}).} m. dregs, sediment (of wine, oil, fat).
	

	
		\key{franciom.}  n.{\footnotesize \textit{ chemistry.}} francium.
	

	
		\key{Frederik.} {\footnotesize (From PG \ipa{*Friþurīks}).} m. Frederic, Frederick, Frederik, Friedrich, Federico, Frederico.
	

	
		\key{fructus.} {\footnotesize (From Lat. \ipa{frūctus}).} m. fruit.
	

	
		\key{frugos.} {\footnotesize (From PIE \ipa{*dʰrówgʰos}).} m. a deception, falsehood, lie.
	

	
		\key{fubros.} {\footnotesize (From PIE \ipa{*dʰubʰrós}).} adj. deep.
	

	\key{fugater,}
		\textit{fugtres.} {\footnotesize (From PIE \ipa{*dʰugh₂tḗr}).} f.
			\textbf{1.} daughter.
			\textbf{2.} Virgo (constellation).
			\textbf{3.} girl.
	

	
		\key{fumos.} {\footnotesize (From PIE \ipa{*dʰuh₂mós}).} m. smoke.
	

	
		\key{fundamental.} {\footnotesize (From Lat. \ipa{fundāmentālis}).} adj. fundamental.
	

	
		\key{fundamentalista.} {\footnotesize (From Dan. \ipa{fundamental}).} adj. most fundamental.
	

	
		\key{fundamentalísta.} {\footnotesize (From Dan. \ipa{fundamental}).} adj. fundamentalist.
	

	
		\key{furista.} {\footnotesize (From Dan. \ipa{furistos}).} f. princess.
	

	
		\key{furistos.} {\footnotesize (From PG \ipa{*furistaz}).} m. prince.
	

	
		\key{furops.} {\footnotesize (From PIE \ipa{*ǵʰwer-o-h₃kʷs}).} m. wild beast.
	

	
		\key{fusilier.} {\footnotesize (From Fr. \ipa{fusilier}).} m. (military) rifleman.
	

	
		\key{fusilieriskos.} {\footnotesize (From Dan. \ipa{fusilier}).} adj. pertaining to riflemen.
	

	
		\key{futi.} {\footnotesize (From PIE \ipa{*dʰéw-e-ti}).} v. to run.
	

	
		\key{futje.} {\footnotesize (From Dan. \ipa{futjos}).} adv. fast, rapidly.
	

	
		\key{futjos.} {\footnotesize (From Dan. \ipa{futi}).} adj. fast.
	

	
		\key{futos.} {\footnotesize (From PIE \ipa{*ǵʰutós}).} adj. poured.
	

	
		\key{gadoliniom.}  n.{\footnotesize \textit{ chemistry.}} gadolinium.
	

	
		\key{galái.} {\footnotesize (From a Turkic language. Compare Azerbaijani \ipa{qalay}).} m.{\footnotesize \textit{ chemistry.}} tin (metal).
	

	
		\key{Gallia.}  f. France.
	

	
		\key{galliom.}  n.{\footnotesize \textit{ chemistry.}} gallium.
	

	
		\key{galliska.} {\footnotesize (From Dan. \ipa{Gallia}).} f. French language.
	

	
		\key{galliskos.} {\footnotesize (From Dan. \ipa{Gallia}).} adj. French.
	

	
		\key{gamma.}  f. Greek letter gamma.
	

	
		\key{gar.} {\footnotesize (Possibly cognate with AG \gr{γᾰ́ρ}).} conj. so that, for the purpose that.
	

	
		\key{garaidíeti.} {\footnotesize (From Dan. \ipa{garaidos}).} v.
			\textbf{1.} to make ready.
			\textbf{2.} to get ready (when used in the mediopassive voice).
	

	
		\key{garaidos.} {\footnotesize (From PG \ipa{*garaidaz}).} adj.
			\textbf{1.} ready, prepared.
			\textbf{2.} prepared, arranged.
	

	
		\key{ge.}  n. letter \textit{g}.
	

	
		\key{general.} {\footnotesize (From Lat. \ipa{generālis}).} adj.
			\textbf{1.} general.
			\textbf{2.} generic.
	

	
		\key{geographia.} {\footnotesize (From Lat. \ipa{geōgraphia}, from AG \gr{γεωγραφία}).} f. geography.
	

	
		\key{geologia.} {\footnotesize (From NL \ipa{geōlogia}).} f. geology.
	

	
		\key{Georgios.} {\footnotesize (From AG \gr{Γεώργῐος}).} m. George, Jorge.
	

	\key{gerán,}
		\textit{granes, granmos.} {\footnotesize (From PIE \ipa{*gerh₂ḗn}).} m. crane, heron.
	

	
		\key{Germania.}  f. Germany.
	

	
		\key{germaniom.}  n.{\footnotesize \textit{ chemistry.}} germanium.
	

	
		\key{germaniska.} {\footnotesize (From Dan. \ipa{Germania}).} f. German language.
	

	
		\key{germaniskos.} {\footnotesize (From Dan. \ipa{Germania}).} adj. German.
	

	
		\key{gladjeti.} {\footnotesize (From PG \ipa{*gladaz}).} v. to please, to make happy, to satisfy.
	

	
		\key{Glen.} {\footnotesize (From En. \ipa{Glen}).} m. Glen, Glenn.
	

	
		\key{Glen Theodoros Siborg.}  m. Glenn Theodore Seaborg.
	

	\key{glis,}
		\textit{gladies, glimos.} {\footnotesize (From PIE \ipa{*gl̥h₁éys}).} m. weasel.
	

	
		\key{gorbos.}  m.
			\textbf{1.} crab.
			\textbf{2.} Cancer (constellation).
	

	
		\key{graftos.} {\footnotesize (From PIE \ipa{*gr̥bʰtós}).} adj.
			\textbf{1.} carved.
			\textbf{2.} written.
	

	
		\key{grammatica.} {\footnotesize (From Lat. \ipa{grammatica}).} f. grammar.
	

	
		\key{grammom.} {\footnotesize (From Fr. \ipa{gramme}, from AG \gr{γράμμα}).} n. gram (unit of measurement).
	

	
		\key{Gregorios.} {\footnotesize (From AG \gr{Γρηγόριος}).} m. Gregory.
	

	
		\key{gronis.} {\footnotesize (From PG \ipa{*grōniz}).} adj. green.
	

	
		\key{gronts.} {\footnotesize (From Dan. \ipa{groti}).} m. vegetation; herbs, pasture; plants.
	

	
		\key{groti.} {\footnotesize (From PG \ipa{*grōaną}).} v. to grow.
	

	
		\key{gruppos.} {\footnotesize (From Italian \ipa{gruppo}).} m. group.
	

	
		\key{guarda.} {\footnotesize (From OF \ipa{guarde}).} f. guard.
	

	
		\key{gursjun.} {\footnotesize (From a Turkic language. Compare Turkmen \ipa{gurşun} and Turkish \ipa{kurşun}).} m.{\footnotesize \textit{ chemistry.}} lead (metal).
	

	
		\key{ha.}  n. letter \textit{h}.
	

	
		\key{Hadriános.} {\footnotesize (From Lat. \ipa{Hadriānus}).} m. Adrian, Hadrian.
	

	
		\key{Hafnia.} {\footnotesize (From Lat. \ipa{Hafnia}).} f. Copenhagen.
	

	
		\key{hafniom.}  n.{\footnotesize \textit{ chemistry.}} hafnium.
	

	
		\key{hassiom.}  n.{\footnotesize \textit{ chemistry.}} hassium.
	

	
		\key{haunis.} {\footnotesize (From PG \ipa{*hauniz}).} adj.
			\textbf{1.} despised, poor, miserable.
			\textbf{2.} sad, depressed.
	

	
		\key{heliom.}  n.{\footnotesize \textit{ chemistry.}} helium.
	

	
		\key{Helsingia.} {\footnotesize (From Lat. \ipa{Helsingia}).} f. Helsinki, capital of Finland.
	

	
		\key{Helvetia.}  f. Switzerland.
	

	
		\key{Hendrik.} {\footnotesize (Ultimately from PG \ipa{*Haimarīks}).} m. Henry, Henrik, Henri, Heinrich, Enrique, Enrico.
	

	
		\key{hertogos.} {\footnotesize (Ultimately from PG \ipa{*harjatogō}. Compare German \ipa{Herzog} and Dutch \ipa{Hertog}).} m. duke.
	

	
		\key{heta.}  f. Greek letter eta.
	

	
		\key{Hibernia.}  f. Ireland.
	

	
		\key{Hispania.}  f. Spain.
	

	
		\key{historia.} {\footnotesize (From Lat. \ipa{historia}, from AG \gr{ἱστορία}).} f.
			\textbf{1.} history.
			\textbf{2.} story, account.
	

	
		\key{Hitania.} {\footnotesize (From a Turkic language. Compare Uzbek \cyr{Хитой}, Turkmen \ipa{Hytaý}, and Tajik \cyr{Хитой}).} f. China.
	

	
		\key{hitaniska.} {\footnotesize (From Dan. \ipa{Hitania}).} f. Chinese language.
	

	
		\key{Hitaniska Ludina Respublica.}  f. People's Republic of China.
	

	
		\key{hitaniskos.} {\footnotesize (From Dan. \ipa{Hitania}).} adj. Chinese.
	

	
		\key{Holmia.} {\footnotesize (From Lat. \ipa{Holmia}).} f. Stockholm.
	

	
		\key{holmiom.}  n.{\footnotesize \textit{ chemistry.}} holmium.
	

	
		\key{honor.} {\footnotesize (From Lat. \ipa{honor}).} m. honor.
	

	
		\key{hostis.} {\footnotesize (From PIE \ipa{*gʰóstis}).} m. guest.
	

	
		\key{hreman.} {\footnotesize (From PIE \ipa{*gʰreh₁}).} n. grass.
	

	
		\key{humánom estin.} {\footnotesize (From Dan. \ipa{humánom + estin}).} n. human being.
	

	
		\key{humános.} {\footnotesize (From Lat. \ipa{hūmānus}).} m. human.
	

	
		\key{hydrogenos.}  n.{\footnotesize \textit{ chemistry.}} hydrogen.
	

	
		\key{hyperbola.} {\footnotesize (From AG \gr{ὑπερβολή}).} f.
			\textbf{1.}{\footnotesize \textit{ mathematics.}} hyperbola.
			\textbf{2.} exaggeration, hyperbole.
	

	
		\key{hypothesis.} {\footnotesize (From Lat. \ipa{hypothesis}, from AG \gr{ὑπόθεσις}).} f. hypothesis.
	

	
		\key{i¹.}  n. letter \textit{i}.
	

	
		\key{i².} {\footnotesize (From PIE \ipa{*ei}).} conj. and.
	

	
		\key{icón.}  f.
			\textbf{1.} image.
			\textbf{2.} icon.
	

	
		\key{ideologia.}  f. ideology.
	

	
		\key{igra.} {\footnotesize (From PIE \ipa{*h₂ig-réh₂}).} f. game.
	

	
		\key{igrek.}  n. letter \textit{y}.
	

	
		\key{igreti.} {\footnotesize (From Dan. \ipa{igra}).} v.
			\textbf{1.} to play a game (use the accusative).
			\textbf{2.} to play an instrument (use \ipa{ana} + the musical instrument).
	

	
		\key{-ilos.} {\footnotesize (Perhaps from PIE \ipa{*-lós}).} elem. diminutive suffix.
	

	
		\key{imperator.} {\footnotesize (From Lat. \ipa{imperātor}).} m. emperor.
	

	
		\key{imperiom.} {\footnotesize (From Lat. \ipa{imperium}).} n. empire.
	

	
		\key{India.}  f. India.
	

	
		\key{indiom.}  n.{\footnotesize \textit{ chemistry.}} indium.
	

	
		\key{ingjenier.} {\footnotesize (From Fr. \ipa{ingénieur}).} m. engineer.
	

	
		\key{ingjenieria.} {\footnotesize (From Fr. \ipa{ingénierie}).} f. engineering.
	

	
		\key{intelligentia.}  f.
			\textbf{1.} intelligence.
			\textbf{2.} intelligentsia, intellectual elite.
	

	
		\key{ipelve.} {\footnotesize (Related to Dan. \ipa{pelus}).} conj. also, in addition, moreover.
	

	
		\key{Irania.}  f. Iran.
	

	
		\key{iridiom.}  n.{\footnotesize \textit{ chemistry.}} iridium.
	

	
		\key{Isak.} {\footnotesize (From AG \gr{Ἰσαᾱ́κ}).} m. Isaac.
	

	
		\key{Isak Njutón.}  m. Isaac Newton.
	

	
		\key{Islandia.} {\footnotesize (From Lat. \ipa{Islandia}).} f. Iceland.
	

	
		\key{istrom.} {\footnotesize (From Dan. \ipa{isus}).} n. bow (weapon).
	

	
		\key{istrom.}  m.
			\textbf{1.} bow (weapon).
			\textbf{2.} Sagittarius (constellation).
	

	
		\key{isus.} {\footnotesize (From PIE \ipa{*(H)ísus}).} m. arrow.
	

	
		\key{Italia.}  f. Italy.
	

	
		\key{italiska.} {\footnotesize (From Dan. \ipa{Italia}).} f. Italian language.
	

	
		\key{itan.} {\footnotesize (From PIE \ipa{*h₁éytr̥}).} n. way, course.
	

	
		\key{iti.} {\footnotesize (From PIE \ipa{*h₁éyti}, with some inflectional forms suppleted by \ipa{*gʷem-}).} v. to go.
	

	
		\key{itos.} {\footnotesize (From PIE \ipa{*h₁itós}).} adj. passable.
	

	
		\key{ive.} {\footnotesize (From PIE \ipa{*(h₁)i-we}).} conj. as, like, in the same manner as.
	

	
		\key{iveman.} {\footnotesize (From Dan. \ipa{ive}).} n. similarity.
	

	
		\key{ix.}  n. letter \textit{x}.
	

	
		\key{ja.}  n. letter \textit{j}.
	

	
		\key{Jacobos.} {\footnotesize (From Lat. \ipa{Jacōbus}).} m. Jacob, James, Jacques.
	

	
		\key{januarios.} {\footnotesize (From Lat. \ipa{Jānuārius}).} m.{\footnotesize \textit{ month.}} January.
	

	
		\key{januarios fimes.}  m. un janvier d'hiver.
	

	
		\key{Japonia.}  f. Japan.
	

	
		\key{japoniska.} {\footnotesize (From Dan. \ipa{Japonia}).} f. Japanese language.
	

	
		\key{japoniskos.} {\footnotesize (From Dan. \ipa{Japonia}).} adj. Japanese.
	

	
		\key{Jasón.} {\footnotesize (From AG \gr{Ἰάσων}).} m. Jason.
	

	
		\key{jehania.} {\footnotesize (From a Turkic language, and ultimately from Middle Persian \ipa{gēhān}; compare Bashkir \cyr{йыһан}).} f. world, universe, cosmos.
	

	
		\key{jemos.} {\footnotesize (From PIE \ipa{*yémHos}).} m. twin.
	

	
		\key{Jereván.} {\footnotesize (From Rus. \cyr{Ерева́н}).} m. Yerevan, capital of Armenia.
	

	
		\key{jeros.} {\footnotesize (From PIE \ipa{*yóh₁r̥}).} m. year.
	

	
		\key{Jesus.} {\footnotesize (From AG \gr{Ἰησοῦς}).} m. Jesus, Joshua.
	

	
		\key{Jesus Christos.}  m. Jesus Christ.
	

	
		\key{jevos.} {\footnotesize (From PIE \ipa{*yéwos}).} n.
			\textbf{1.} cereal, grain.
			\textbf{2.} barley, spelt.
	

	
		\key{Joánna.} {\footnotesize (From AG \gr{Ἰωάννης}).} m. John.
	

	
		\key{Joánna Karlos Frederik Gauss.}  m. Johann Carl Friedrich Gauss.
	

	
		\key{jodi.} {\footnotesize (From Dan. \ipa{jo- + -di}. The suffix \ipa{-di} comes from PIE \ipa{*dʰi}).} proadv. (relative) where.
	

	
		\key{jodom.}  n.{\footnotesize \textit{ chemistry.}} iodine.
	

	
		\key{joimos.}  proadv.
			\textbf{1.} (relative) to where.
			\textbf{2.} (relative) why, the reason for which. conj.
			\textbf{3.} because.
	

	
		\key{joives.} {\footnotesize (From Dan. \ipa{jo- + -ives}. The suffix \ipa{-ives} comes from \ipa{juves}, the genitive of \ipa{oju}).} proadv.
			\textbf{1.} (relative) when. conj.
			\textbf{2.} when.
	

	
		\key{Jonathán.} {\footnotesize (From AG \gr{Ἰωνάθαν}).} m. Jonathan.
	

	
		\key{Jonia.} {\footnotesize (From AG \gr{Ἰωνῐ́ᾱ}).} f. Greece.
	

	
		\key{joniska.} {\footnotesize (From Dan. \ipa{Jonia}).} f. Greek language.
	

	\key{jopan,}
		\textit{jepens.} {\footnotesize (From PIE \ipa{*yókʷr̥}).} n. liver.
	

	
		\key{Josephos.} {\footnotesize (From Lat. \ipa{Iōsēphus}).} m. Joseph, José.
	

	
		\key{josmad.}  proadv. (relative) from where.
	

	
		\key{josme.}  proadv. (relative) how.
	

	
		\key{jota.}  f. Greek letter iota.
	

	
		\key{Jotfredos.} {\footnotesize (Ultimately from PG \ipa{*Gautafriþuz}).} m. Jeffrey, Geoffrey, Jauffre, Gaufredo, Gottfried.
	

	
		\key{joti.}  proadv.
			\textbf{1.} (relative) how much. conj.
			\textbf{2.} than (for making comparisons).
	

	
		\key{Jovis.}  m. Jove, Jupiter (the god).
	

	
		\key{jugom.} {\footnotesize (From PIE \ipa{*yugóm}).} n. yoke.
	

	
		\key{jugos.} {\footnotesize (From PIE \ipa{*yéwgos}).} n.
			\textbf{1.} team.
			\textbf{2.} pair of yoked animals.
	

	
		\key{juktos.} {\footnotesize (From PIE \ipa{*yugtós}).} adj. yoked, joined.
	

	
		\key{julios.} {\footnotesize (From Lat. \ipa{Jūlius}).} m.{\footnotesize \textit{ month.}} July.
	

	
		\key{junios.} {\footnotesize (From Lat. \ipa{Jūnius}).} m.{\footnotesize \textit{ month.}} June.
	

	
		\key{junisos.} {\footnotesize (From Dan. \ipa{junsos}).} adj. younger.
	

	
		\key{junisos sergjant.}  m. (military) junior sergeant; corporal.
	

	
		\key{junistos.} {\footnotesize (From Dan. \ipa{junsos}).} adj. youngest.
	

	
		\key{junsos.} {\footnotesize (From PIE \ipa{*h₂yuHn̥ḱós}).} adj. young.
	

	
		\key{junta.} {\footnotesize (From PIE \ipa{*h₂yuHn̥téh₂}).} f. youth.
	

	
		\key{Justínos.} {\footnotesize (From Lat. \ipa{Iūstīnus}).} m. Justin.
	

	
		\key{ka.}  n. letter \textit{k}.
	

	
		\key{kagjom.} {\footnotesize (From PIE \ipa{*kagʰyóm}).} n. enclosure, hedge.
	

	
		\key{kaikos.} {\footnotesize (From PIE \ipa{*káykos}).} adj. blind.
	

	
		\key{kaliom.}  n.{\footnotesize \textit{ chemistry.}} potassium.
	

	
		\key{kamán.} {\footnotesize (From Persian \arab{کمان}).} m. bow (weapon).
	

	
		\key{kania.}  f. song.
	

	
		\key{kappa.}  f. Greek letter kappa.
	

	
		\key{kapros.} {\footnotesize (From PIE \ipa{*kápros}).} m.
			\textbf{1.} goat.
			\textbf{2.} Capricornus (constellation).
	

	
		\key{kaptos.} {\footnotesize (From PIE \ipa{*kaptós}).} adj. seized, held.
	

	
		\key{kaput.} {\footnotesize (From PIE \ipa{*káput}).} n. head.
	

	
		\key{kaputjos.} {\footnotesize (From Dan. \ipa{kaput}).} adj. main, principal.
	

	
		\key{Karlos.} {\footnotesize (From PG \ipa{*karlaz}).} m. Charles, Carlos.
	

	
		\key{Karsna Bora.}  f. Montenegro, Monte Negro.
	

	
		\key{karsnos.} {\footnotesize (From PIE \ipa{*kr̥snós}).} adj.
			\textbf{1.} black.
			\textbf{2.} dark, dusky.
	

	
		\key{Kazachstán.} {\footnotesize (From Rus. \cyr{Казахста́н}).} m. Kazakhstan.
	

	\key{kentis,}
		\textit{kanties.} {\footnotesize (From PIE \ipa{*ken-}).} f. beginning, start.
	

	
		\key{Kiev.} {\footnotesize (From Rus. \cyr{Ки́ев}).} m. Kyiv, Kiev, capital of Ukraine.
	

	
		\key{Kisjinjov.} {\footnotesize (From Rus. \ipa{Кишинёв}).} m. Chisinau, capital of Moldova.
	

	
		\key{klava.} {\footnotesize (From PG \ipa{*klawō}).} f. claw.
	

	
		\key{koksa.} {\footnotesize (From PIE \ipa{*kóḱs-}).} f. (anatomy) joint.
	

	\key{kotus,}
		\textit{katves.} {\footnotesize (From PIE \ipa{*kéh₃tus}).} m. fight.
	

	
		\key{kreva.} {\footnotesize (From PIE \ipa{*kréwh₂s}).} f. (raw) blood.
	

	
		\key{kros.} {\footnotesize (From PIE \ipa{*h₂ḱrós}).} adj. sharp.
	

	
		\key{krosna.} {\footnotesize (From PIE \ipa{*ḱerh₂-}).} f. horn.
	

	
		\key{kruros.} {\footnotesize (From PIE \ipa{*kruh₂rós}).} adj. bloody.
	

	
		\key{kryptón.}  m.{\footnotesize \textit{ chemistry.}} krypton.
	

	
		\key{kudeti.} {\footnotesize (From PIE \ipa{*kéwdʰh₁-e-ti}).} v. to forge.
	

	
		\key{kulom.} {\footnotesize (From PG \ipa{*kulą}).} n.
			\textbf{1.} coal.
			\textbf{2.}{\footnotesize \textit{ chemistry.}} carbon.
	

	
		\key{lambda.}  f. Greek letter lambda.
	

	
		\key{lana.} {\footnotesize (From PIE \ipa{*h₂wĺ̥h₁neh₂}).} f. wool.
	

	
		\key{lanfros.} {\footnotesize (From PIE \ipa{*h₁ln̥gʷʰrós}).} adj. lightweight.
	

	
		\key{lanthanom.}  n.{\footnotesize \textit{ chemistry.}} lanthanum.
	

	
		\key{latiniska.}  f. Latin language.
	

	
		\key{latos.} {\footnotesize (From PIE \ipa{*tl̥h₂tós}).} adj. lifted.
	

	
		\key{laurentiom.}  n.{\footnotesize \textit{ chemistry.}} lawrencium.
	

	
		\key{Laurentios.} {\footnotesize (From Lat. \ipa{Laurentius}).} m. Lawrence, Laurence, Lorenzo, Laurent.
	

	
		\key{lekseter.} {\footnotesize (From Dan. \ipa{lekseti}).} m. defender, defensor, protector.
	

	
		\key{lekseti.} {\footnotesize (From PIE \ipa{*h₂lékseti}).} v. to protect, defend.
	

	
		\key{leksetis.} {\footnotesize (From Dan. \ipa{lekseti}).} f. defense, protection.
	

	
		\key{Leningrad.} {\footnotesize (From Rus. \cyr{Ленингра́д}).} m. Leningrad.
	

	
		\key{león.}  m.
			\textbf{1.} lion.
			\textbf{2.} Leo (constellation).
	

	
		\key{Leonardos Óiler.}  m. Leonhard Euler.
	

	
		\key{Lettonia.} {\footnotesize (From Lat. \ipa{Lettonia}).} f. Latvia.
	

	
		\key{Lichtenstain.} {\footnotesize (From Ger. \ipa{Liechtenstein}).} m. Liechtenstein.
	

	
		\key{linepti.}  v. to leave.
	

	
		\key{linguistica.} {\footnotesize (From Fr. \ipa{linguistique}, with a latinized ending).} f. linguistics.
	

	
		\key{linja.} {\footnotesize (From Fr. \ipa{ligne}).} f. line.
	

	
		\key{lithiom.}  n.{\footnotesize \textit{ chemistry.}} lithium.
	

	
		\key{Lithuania.} {\footnotesize (From Lat. \ipa{Lithuania}).} f. Lithuania.
	

	
		\key{livermoriom.}  n.{\footnotesize \textit{ chemistry.}} livermorium.
	

	
		\key{loisa.} {\footnotesize (From PIE \ipa{*lóyseh₂}).} f.
			\textbf{1.} following, track.
			\textbf{2.} (agriculture) furrow.
	

	\key{lokus,}
		\textit{alkves.} {\footnotesize (From PIE \ipa{*lókus}).} m.
			\textbf{1.} pond, pool.
			\textbf{2.} lake.
	

	
		\key{Londiniom.} {\footnotesize (From Lat. \ipa{Londīnium}).} n. London.
	

	
		\key{longos.} {\footnotesize (From PIE \ipa{*dlongʰos}).} adj. long.
	

	
		\key{Lubiána.} {\footnotesize (From It. \ipa{Lubiana}).} f. Ljubljana, capital of Slovenia.
	

	
		\key{lucta.} {\footnotesize (From Lat. \ipa{lucta}).} f. fight.
	

	
		\key{Ludina Opisvomiska Orda.}  f. People's Liberation Army.
	

	
		\key{ludinos.} {\footnotesize (From Dan. \ipa{ludis}).} adj.
			\textbf{1.} of the people.
			\textbf{2.} popular.
	

	
		\key{ludis.} {\footnotesize (From PIE \ipa{*h₁léwdʰis}).} m. people.
	

	
		\key{lufjeti.} {\footnotesize (From PIE \ipa{*lewbʰ-}).} v. to love.
	

	
		\key{luketer.} {\footnotesize (From Dan. \ipa{lúketor}).} v. shiner, illuminator, source of light.
	

	
		\key{lúketor.} {\footnotesize (From PIE \ipa{*lewk-}).} v. to shine, to illuminate.
	

	
		\key{lukos.} {\footnotesize (From PIE \ipa{*lewk-}).} m. light.
	

	
		\key{Luna.} {\footnotesize (From PIE \ipa{*lówksneh₂}, with influence from Lat. \ipa{lūna}).} f. the Moon.
	

	
		\key{lus.} {\footnotesize (From PIE \ipa{*lewH-}).} m. louse.
	

	
		\key{lutetiom.}  n.{\footnotesize \textit{ chemistry.}} lutetium.
	

	
		\key{luvos.} {\footnotesize (From Dan. \ipa{lufjeti}).} m. love.
	

	
		\key{Luxembúrgom.} {\footnotesize (From Lat. \ipa{Luxemburgum}).} n. Luxembourg.
	

	
		\key{Madrid.}  m. Madrid.
	

	
		\key{magnesiom.}  n.{\footnotesize \textit{ chemistry.}} magnesium.
	

	
		\key{maitneriom.}  n.{\footnotesize \textit{ chemistry.}} meitnerium.
	

	
		\key{majos.} {\footnotesize (From Lat. \ipa{Maius}).} m.{\footnotesize \textit{ month.}} May.
	

	
		\key{makros.} {\footnotesize (From PIE \ipa{*mh₂ḱrós}).} adj. lean, meager.
	

	
		\key{Malta.}  f. Malta.
	

	
		\key{manganiom.}  n.{\footnotesize \textit{ chemistry.}} manganese.
	

	
		\key{mánjetor.} {\footnotesize (From PIE \ipa{*mn̥-yé-tor}).} v.
			\textbf{1.} to be thinking, to ponder.
			\textbf{2.} to think about something (use the preposition about).
	

	
		\key{manus.}  m. person, human.
	

	
		\key{Marcos.} {\footnotesize (From Lat. \ipa{Marcus}).} m. Mark.
	

	
		\key{mardus.} {\footnotesize (From PIE \ipa{*mr̥ǵʰús}).} adj.
			\textbf{1.} short.
			\textbf{2.} brief.
	

	
		\key{mardus.} {\footnotesize (From PIE \ipa{*mr̥ǵʰús}).} adj.
			\textbf{1.} short.
			\textbf{2.} brief.
	

	
		\key{márjetor.} {\footnotesize (From PIE \ipa{*mr̥-yé-tor}).} v. to die.
	

	
		\key{martios.} {\footnotesize (From Lat. \ipa{Mārtius}).} m.{\footnotesize \textit{ month.}} March.
	

	
		\key{martos.} {\footnotesize (From PIE \ipa{*mr̥tós}).} adj.
			\textbf{1.} dead.
			\textbf{2.} mortal.
	

	
		\key{Marts.}  m.
			\textbf{1.} Mars (the god).
			\textbf{2.} Mars (the planet).
	

	
		\key{mater.} {\footnotesize (From PIE \ipa{*méh₂tēr}).} f. mother.
	

	
		\key{mathematica.} {\footnotesize (From Lat. \ipa{mathēmatica}).} f. mathematics.
	

	
		\key{Matthaios.} {\footnotesize (From AG \gr{Ματθαῖος}).} m. Matthew.
	

	
		\key{Mauritios.} {\footnotesize (From Lat. \ipa{Mauritius}).} m. Maurice, Mauricio, Mauritius.
	

	
		\key{medanistos.} {\footnotesize (From Dan. \ipa{medanos}).} adj. biggest.
	

	
		\key{medanjés.} {\footnotesize (From Dan. \ipa{medanos}).} adj. bigger.
	

	
		\key{medanos.} {\footnotesize (From PIE \ipa{*méǵh₂s}).} adj. big, great.
	

	
		\key{medi.} {\footnotesize (From PIE \ipa{*médʰyos}).} prep.
			\textbf{1.} (+ins) between, among, in the middle of (stationary, answers the question "where?").
			\textbf{2.} (+acc) between, among (with movement, answers the question "where to?").
	

	
		\key{medjos.} {\footnotesize (From PIE \ipa{*médʰyos}).} adj.
			\textbf{1.} middle. m.
			\textbf{2.} sex, gender.
	

	\key{melit,}
		\textit{malites.} {\footnotesize (From PIE \ipa{*mélit}).} n. honey.
	

	
		\key{melus.}  pron. little; few; not much.
	

	
		\key{melvjés.} {\footnotesize (From Dan. \ipa{melus}).} adj.
			\textbf{1.} smaller; minor.
			\textbf{2.} younger.
	

	
		\key{memorabilis.} {\footnotesize (From Lat. \ipa{memorabilis}).} adj. memorable.
	

	
		\key{mendeleviom.}  n.{\footnotesize \textit{ chemistry.}} mendelevium.
	

	
		\key{menos.} {\footnotesize (From PIE \ipa{*ménos}).} n. mind.
	

	
		\key{mens¹.} {\footnotesize (From PIE \ipa{*mḗh₁n̥s}).} m.
			\textbf{1.} moon.
			\textbf{2.} month.
	

	
		\key{mens².} {\footnotesize (From PIE \ipa{*mḗms}).} n. flesh, meat.
	

	\key{mentis,}
		\textit{manties.} {\footnotesize (From PIE \ipa{*méntis}).} f.
			\textbf{1.} thought.
			\textbf{2.} memory.
	

	
		\key{Mercurios.}  m.
			\textbf{1.} Mercury (the god).
			\textbf{2.} Mercury  (the planet).
	

	
		\key{merti.} {\footnotesize (From PIE \ipa{*mér-ti}).} v. to die.
	

	
		\key{mertinos.} {\footnotesize (From Dan. \ipa{mertis + -inos}).} adj. deadly.
	

	\key{mertis,}
		\textit{marties.} {\footnotesize (From PIE \ipa{*mértis}).} f. death.
	

	
		\key{metrom.} {\footnotesize (From Fr. \ipa{mètre}, from AG \gr{μέτρον}).} n. meter (unit of measurement).
	

	
		\key{Mexicom.}  n. Mexico.
	

	
		\key{Miami.} {\footnotesize (From En. \ipa{Miami}).} n. Miami.
	

	
		\key{Michael.} {\footnotesize (From AG \gr{Μῐχᾱήλ}).} m. Michael.
	

	
		\key{mihla.} {\footnotesize (From PIE \ipa{*h₃migʰleh₂}).} f. mist, fog.
	

	
		\key{militar.} {\footnotesize (From Lat. \ipa{mīlitāris}).} adj. military.
	

	
		\key{militia.}  f. militia.
	

	
		\key{Miniskos.} {\footnotesize (From Rus. \cyr{Минск}, with a Danetian ending).} m. Minsk, capital of Belarus.
	

	
		\key{ministos.} {\footnotesize (From Dan. \ipa{minos}).} adj. smallest.
	

	
		\key{minjés.} {\footnotesize (From Dan. \ipa{minos}).} adj. smaller.
	

	
		\key{minos.}  adj. small.
	

	
		\key{misdos.} {\footnotesize (From PIE \ipa{*misdʰós}).} m.
			\textbf{1.} reward, prize.
			\textbf{2.} wages, payment.
	

	
		\key{mjan.} {\footnotesize (Onomatopoeia).} intj. meow.
	

	
		\key{mjanka.} {\footnotesize (From the diminutive of Dan. \ipa{mjan}).} f.
			\textbf{1.} gentle meow.
			\textbf{2.} little cat, kitten.
	

	
		\key{Moldavia.} {\footnotesize (From Rus. \cyr{Молда́вия}).} f. Moldova, Moldavia, Moldovia.
	

	
		\key{moldus.} {\footnotesize (From PIE \ipa{*(h₂)moldus}).} adj. soft, weak.
	

	
		\key{molybdainom.}  n.{\footnotesize \textit{ chemistry.}} molybdenum.
	

	
		\key{Monaco.}  n. Monaco.
	

	\key{mori,}
		\textit{maries.} {\footnotesize (From PIE \ipa{*móri}).} n. sea.
	

	
		\key{Mori Mantióm.}  n. primordial sea of thoughts.
	

	\key{morvis,}
		\textit{marvies.} {\footnotesize (From PIE \ipa{*morwi-}).} f. ant.
	

	
		\key{Moscovia.}  f. Moscow.
	

	
		\key{moscoviom.}  n.{\footnotesize \textit{ chemistry.}} moscovium.
	

	
		\key{mosgos.} {\footnotesize (From PIE \ipa{*mosgʰós}).} m.
			\textbf{1.} marrow.
			\textbf{2.} brain.
	

	\key{mus,}
		\textit{muses.} {\footnotesize (From PIE \ipa{*múh₂s}).} m. mouse.
	

	
		\key{musica.} {\footnotesize (From Fr. \ipa{musique}).} f. music.
	

	
		\key{mutrom.} {\footnotesize (From PIE \ipa{*múHtrom}).} n.
			\textbf{1.} excrement, urine.
			\textbf{2.} filth, sediment.
	

	
		\key{my.}  n. Greek letter mu.
	

	
		\key{nas.} {\footnotesize (From PIE \ipa{*néh₂s}).} f. nose.
	

	
		\key{natos¹.} {\footnotesize (From PIE \ipa{*ǵn̥h₁tós}).} adj. begotten, produced.
	

	
		\key{natos².} {\footnotesize (From PIE \ipa{*ǵn̥h₃tós}).} adj. known, recognizable.
	

	
		\key{natriom.}  n.{\footnotesize \textit{ chemistry.}} sodium.
	

	
		\key{natúra.} {\footnotesize (From Lat. \ipa{nātūra}).} f. nature.
	

	
		\key{natural.} {\footnotesize (From Lat. \ipa{nātūrālis}).} adj. natural.
	

	
		\key{natural donvos.}  m.{\footnotesize \textit{ mathematics.}} natural number.
	

	
		\key{nauman.} {\footnotesize (From PIE \ipa{*neh₂ut-}).} n. need, necessity.
	

	\key{naus,}
		\textit{naves, naumos.} {\footnotesize (From PIE \ipa{*néh₂us}).} f. boat, ship.
	

	
		\key{nebidinos.} {\footnotesize (From Dan. \ipa{bidinos}).} adj. unbelievable, incredible.
	

	\key{nebufmén,}
		\textit{nebufnes, nebudanmos.} {\footnotesize (From Dan. \ipa{ne + bufmén}).} m. abyss, bottomless pit.
	

	
		\key{nefos.} {\footnotesize (From PIE \ipa{*nébʰos}).} n. cloud.
	

	
		\key{nefros.} {\footnotesize (From PIE \ipa{*negʷʰrós}).} m. kidney.
	

	
		\key{nege.} {\footnotesize (From PIE \ipa{*ne + *gʰe}).} adv. not even.
	

	
		\key{nemartos.} {\footnotesize (From Dan. \ipa{ne + martos}).} adj. immortal, unperishable.
	

	
		\key{nemeti.} {\footnotesize (From PG \ipa{*nemaną}).} v. to take.
	

	
		\key{nemos.} {\footnotesize (From PIE \ipa{*némos}).} n.
			\textbf{1.} sacrifice.
			\textbf{2.} worship.
	

	
		\key{neodymiom.}  n.{\footnotesize \textit{ chemistry.}} neodymium.
	

	
		\key{neón.}  m.{\footnotesize \textit{ chemistry.}} neon.
	

	
		\key{neptuniom.}  n.{\footnotesize \textit{ chemistry.}} neptunium.
	

	\key{ner,}
		\textit{andres, narmos.} {\footnotesize (From PIE \ipa{*h₂nḗr}).} m. power, force, vital energy.
	

	
		\key{nespe.} {\footnotesize (From Dan. \ipa{ne + espe}).} adv. not yet.
	

	
		\key{Nestor.} {\footnotesize (From AG \gr{Νέστωρ}).} m.{\footnotesize \textit{ personal name.}} Nestor. {\footnotesize \ipa{Nestor, tu luvos menjas bivas}. \textit{Nestor, you are the love of my life}.}
	

	
		\key{nesus.} {\footnotesize (From PIE \ipa{*néḱus}).} adj.
			\textbf{1.} dead.
			\textbf{2.} mortal.
	

	
		\key{Neva Jersis.} {\footnotesize (From Dan. \ipa{neva} and En. \ipa{Jersey}).} f. New Jersey.
	

	
		\key{Neva York.} {\footnotesize (From Dan. \ipa{neva + York}).} f. New York.
	

	
		\key{nevamos.}  adj. ninth.
	

	
		\key{nevan.} {\footnotesize (From PIE \ipa{*h₁néwn̥}).} num. nine.
	

	
		\key{nevandesam.} {\footnotesize (From PIE \ipa{*h₁néwn̥ + *déḱm̥}).} num. nineteen.
	

	
		\key{nevandesamos.}  adj. nineteenth.
	

	
		\key{nevandsont.} {\footnotesize (From PIE \ipa{*h₁néwn̥dḱomt}).} num. ninety.
	

	
		\key{nevandsontos.} {\footnotesize (From PIE \ipa{*h₁néwn̥dḱomt}).} adj. ninetieth.
	

	
		\key{nevidmos.} {\footnotesize (From Dan. \ipa{voide}).} adj. unseen.
	

	
		\key{nevos.} {\footnotesize (From PIE \ipa{*néwos}).} adj. new.
	

	
		\key{niccolom.}  n.{\footnotesize \textit{ chemistry.}} nickel.
	

	
		\key{Nicola.} {\footnotesize (Ultimately from AG \gr{Νῑκόλᾱος}).} m. Nicholas, Nikola, Nikolai.
	

	
		\key{Nicosia.} {\footnotesize (From Fr. \ipa{Nicosie}, from AG \gr{Λευκωσία}).} f. Nicosia, capital of Cyprus.
	

	
		\key{nihoniom.}  n.{\footnotesize \textit{ chemistry.}} nihonium.
	

	
		\key{Nils.} {\footnotesize (From Danish \ipa{Niels}).} m. Niels.
	

	
		\key{Nils Bor.}  m. Niels Bohr.
	

	
		\key{niobiom.}  n.{\footnotesize \textit{ chemistry.}} niobium.
	

	
		\key{nipodi.} {\footnotesize (From Dan. \ipa{nipo- + -di}. The suffix \ipa{-di} comes from PIE \ipa{*dʰi}).} proadv. (negatory) nowhere.
	

	
		\key{nipoimos.}  proadv.
			\textbf{1.} (negatory) to nowhere.
			\textbf{2.} (negatory) for no reason.
	

	
		\key{nipoives.} {\footnotesize (From Dan. \ipa{nipo- + -ives}. The suffix \ipa{-ives} comes from \ipa{juves}, the genitive of \ipa{oju}).} proadv. (negatory) never.
	

	
		\key{niposmad.}  proadv. (negatory) from nowhere.
	

	
		\key{niposme.}  proadv. (negatory) in no way, in no manner.
	

	
		\key{nipoti.}  proadv. (negatory) in no amount, by no amount.
	

	
		\key{nisdos.} {\footnotesize (From PIE \ipa{*nisdós}).} m. nest.
	

	
		\key{nitrogenos.}  n.{\footnotesize \textit{ chemistry.}} nitrogen.
	

	
		\key{nobeliom.}  n.{\footnotesize \textit{ chemistry.}} nobelium.
	

	
		\key{nofilos.} {\footnotesize (From PIE \ipa{*h₃nobʰilos}).} m. navel.
	

	\key{nogs,}
		\textit{neges.} {\footnotesize (From PIE \ipa{*h₃nṓgʰs}).} m. nail (of the finger or toe).
	

	\key{noman,}
		\textit{nemens.} {\footnotesize (From PIE \ipa{*h₁nómn̥}).} n. name.
	

	\key{nopts,}
		\textit{neptes.} {\footnotesize (From PIE \ipa{*nókʷts}).} f. night.
	

	
		\key{Nordina America.} {\footnotesize (From Dan. \ipa{nordina + America}).} f.{\footnotesize \textit{ continent.}} North America.
	

	
		\key{Nordina Coréa.}  f. North Korea.
	

	
		\key{Nordina Macedonia.}  f. North Macedonia.
	

	
		\key{nordinos.} {\footnotesize (From Dan. \ipa{nordom}).} adj. northern, north, northerly.
	

	
		\key{nordom.} {\footnotesize (From PG \ipa{*nurþrą}).} n. north.
	

	
		\key{Norvegia.} {\footnotesize (From Lat. \ipa{Norvegia}).} f. Norway.
	

	
		\key{norvegiska.} {\footnotesize (From Dan. \ipa{Norvegia}).} f. Norwegian language.
	

	
		\key{november.} {\footnotesize (From Lat. \ipa{November}).} m.{\footnotesize \textit{ month.}} November.
	

	
		\key{nu.} {\footnotesize (From PIE \ipa{*nu}).} conj. but.
	

	
		\key{numan.} {\footnotesize (From PIE \ipa{*pnéwmn̥}).} n.
			\textbf{1.} breath.
			\textbf{2.} spirit, soul.
	

	
		\key{numenjeti.} {\footnotesize (From Dan. \ipa{numan}).} v. to breathe.
	

	
		\key{Nur-Sultán.}  m. Nur-Sultan, capital of Kazakhstan.
	

	
		\key{nuti.} {\footnotesize (Unknown).} prep. (+abl) without.
	

	
		\key{nutipretiopés.} {\footnotesize (From Dan. \ipa{nuti + pretiops}).} adj. faceless, formless, shapeless.
	

	
		\key{ny.}  n. Greek letter nu.
	

	
		\key{o.}  n. letter \textit{o}.
	

	
		\key{october.} {\footnotesize (From Lat. \ipa{Octōber}).} m.{\footnotesize \textit{ month.}} October.
	

	
		\key{odés.} {\footnotesize (From PIE \ipa{*h₃edʰḗs}).} m. adze, axe, hatchet.
	

	
		\key{oganessón.}  m.{\footnotesize \textit{ chemistry.}} oganesson.
	

	
		\key{oiconomia.} {\footnotesize (From Lat. \ipa{oeconomia}, from AG \gr{οἰκονομία}).} f. economy.
	

	
		\key{oinika.} {\footnotesize (From Dan. \ipa{oinos}).} f. unit.
	

	
		\key{oinodesam.} {\footnotesize (From PIE \ipa{*óynos + *déḱm̥}).} num. eleven.
	

	
		\key{oinodesamos.}  adj. eleventh.
	

	
		\key{oinos.} {\footnotesize (From PIE \ipa{*óynos}).} num. one.
	

	
		\key{oitos.} {\footnotesize (From PIE \ipa{*h₁óytos}).} m. oath, covenant.
	

	
		\key{ojom.} {\footnotesize (From PIE \ipa{*h₂ōwyóm}).} n. egg.
	

	\key{oju,}
		\textit{juves, jumos.} {\footnotesize (From PIE \ipa{*h₂óyu}).} n. long life, lifetime.
	

	\key{okris,}
		\textit{ekris.} {\footnotesize (From PIE \ipa{*h₂óḱris}).} f.
			\textbf{1.} top, protrusion.
			\textbf{2.} sharp edge, corner.
	

	
		\key{oktamos.}  adj. eighth.
	

	
		\key{okto.} {\footnotesize (From PIE \ipa{*oḱtṓw}).} num. eight.
	

	
		\key{oktodesam.} {\footnotesize (From PIE \ipa{*oḱtṓw + *déḱm̥}).} num. eighteen.
	

	
		\key{oktodesamos.}  adj. eighteenth.
	

	
		\key{oktodsont.} {\footnotesize (From PIE \ipa{*oḱtódḱomt}).} num. eighty.
	

	
		\key{oktodsontos.} {\footnotesize (From PIE \ipa{*oḱtódḱomt}).} adj. eightieth.
	

	
		\key{Olisipón.} {\footnotesize (From Lat. \ipa{Olisīpō}).} f. Lisbon, capital of Portugal.
	

	
		\key{olos.} {\footnotesize (From PIE \ipa{*h₁ólos}).} m. awl.
	

	\key{omban,}
		\textit{ambens.} {\footnotesize (From PIE \ipa{*h₃éngʷn̥}).} n. fat, butter.
	

	
		\key{ombol.} {\footnotesize (From PIE \ipa{*h₂óngʷōl}).} m. charcoal.
	

	
		\key{omega.}  f. Greek letter omega.
	

	
		\key{omicrón.}  f. Greek letter omicron.
	

	\key{oms,}
		\textit{ames.} {\footnotesize (From PIE \ipa{*h₂ṓms}).} m. shoulder.
	

	
		\key{onkos.} {\footnotesize (From PIE \ipa{*h₂ónkos}).} m. bend, turn.
	

	
		\key{onos.} {\footnotesize (From PIE \ipa{*h₃énh₂os}).} n. load, burden.
	

	
		\key{onostos.} {\footnotesize (From PIE \ipa{*h₃enh₂ostos}).} adj. loaded; burdened.
	

	
		\key{opi.} {\footnotesize (From PIE \ipa{*h₁opi}).} prep.
			\textbf{1.} (+gen) near, at, by.
			\textbf{2.} (+gen) before (in the presence of).
	

	
		\key{opigladjeti.} {\footnotesize (From Dan. \ipa{gladjeti}).} v. to comfort, to console.
	

	
		\key{opisagjeti.} {\footnotesize (From Dan. \ipa{opi- + sagjeti}).} v. to celebrate.
	

	
		\key{opistanati.}  v. to remain.
	

	
		\key{opisvomia.} {\footnotesize (From Dan. \ipa{opi + svomia}).} f. liberation.
	

	
		\key{opisvomíeti.} {\footnotesize (From Dan. \ipa{opisvomia}).} v. to liberate, to free, to set free.
	

	
		\key{opíti.} {\footnotesize (From Dan. \ipa{opi} and \ipa{iti}).} v. to come.
	

	\key{ops,}
		\textit{opes.} {\footnotesize (From PIE \ipa{*h₃ókʷs}).} n. eye.
	

	
		\key{orbita.} {\footnotesize (From Lat. \ipa{orbita}).} f. orbit.
	

	
		\key{orda.} {\footnotesize (From OT \ipa{ordu}, from PT \ipa{*or-}. Compare Rus. \cyr{орда}).} f. army.
	

	
		\key{ordinator.} {\footnotesize (From Fr. \ipa{ordinateur}, influenced by Lat. \ipa{ōrdinātor}).} m. computer.
	

	
		\key{ordis.} {\footnotesize (From PIE \ipa{*h₄órǵʰis}).} m. testicle.
	

	
		\key{orfos.} {\footnotesize (From PIE \ipa{*h₃órbʰos}).} m.
			\textbf{1.} orphan.
			\textbf{2.} servant, slave, worker.
	

	\key{orón,}
		\textit{arnes, ranmos.} {\footnotesize (From PIE \ipa{*h₃érō}).} m. eagle.
	

	
		\key{orthographia.} {\footnotesize (From Lat. \ipa{orthographia}, from AG \gr{ὀρθογρᾰφῐ́ᾱ}).} f. orthography.
	

	
		\key{os.} {\footnotesize (From PIE \ipa{*h₁óh₃s}).} n. mouth.
	

	
		\key{osdos.} {\footnotesize (From PIE \ipa{*h₃ósdos}).} m.
			\textbf{1.} branch.
			\textbf{2.} (religious) denomination.
	

	
		\key{Oslo.}  n. Oslo, capital of Norway.
	

	
		\key{osmiom.}  n.{\footnotesize \textit{ chemistry.}} osmium.
	

	\key{ost,}
		\textit{astes.} {\footnotesize (From PIE \ipa{*h₂óst}).} n. bone.
	

	\key{ovis,}
		\textit{avies.} {\footnotesize (From PIE \ipa{*h₂ówis}).} f.
			\textbf{1.} sheep.
			\textbf{2.} Aries (constellation).
	

	
		\key{oxygenos.}  n.{\footnotesize \textit{ chemistry.}} oxygen.
	

	
		\key{palatial.} {\footnotesize (From Fr. \ipa{palatial}, with a latinized ending).} adj. palatial, pertaining to palaces.
	

	
		\key{palatiom.} {\footnotesize (From Lat. \ipa{palātium}).} n. palace.
	

	
		\key{Palatiom Versaljóm.} {\footnotesize (From Dan. \ipa{palaciom + Versaljai}).} n. Palace of Versailles.
	

	
		\key{palladiom.}  n.{\footnotesize \textit{ chemistry.}} palladium.
	

	
		\key{paltania.} {\footnotesize (From Dan. \ipa{paltanos}).} f. width, breadth.
	

	
		\key{paltanos.} {\footnotesize (From PIE \ipa{*pl̥th₂enós}).} adj. wide, broad.
	

	
		\key{panstis.} {\footnotesize (From PIE \ipa{*pn̥kʷstis}).} f. fist.
	

	
		\key{parabola.} {\footnotesize (From AG \gr{παραβολή}).} f.
			\textbf{1.}{\footnotesize \textit{ mathematics.}} parabola.
			\textbf{2.} parable, proverb.
	

	
		\key{paráda.} {\footnotesize (From Fr. \ipa{parade}).} f. parade.
	

	
		\key{París.}  f. Paris.
	

	
		\key{parmis.} {\footnotesize (From PIE \ipa{*kʷŕ̥mis}).} m. worm.
	

	
		\key{particula.} {\footnotesize (From Lat. \ipa{particula}).} f.{\footnotesize \textit{ physics.}} particle.
	

	
		\key{partisaniskos.}  adj. pertaining to guerrillas.
	

	
		\key{pater.} {\footnotesize (From PIE \ipa{*ph₂tḗr}).} m. father.
	

	
		\key{patrinos.} {\footnotesize (From Dan. \ipa{pater + inos}).} adj. patriotic.
	

	
		\key{Paulos.} {\footnotesize (From Lat. \ipa{Paulus}).} m. Paul.
	

	
		\key{Paulos Hadriános Mauritios Dirac.}  m. Paul Adrien Maurice Dirac.
	

	
		\key{pauros.} {\footnotesize (From PIE \ipa{*peh₂w-}).} pron. few, little of.
	

	
		\key{pavan.} {\footnotesize (From PIE \ipa{*péh₂wr̥}).} n. bonfire.
	

	
		\key{pavinos.} {\footnotesize (From Dan. \ipa{pavan}).} adj. fiery.
	

	
		\key{pe.}  n. letter \textit{p}.
	

	
		\key{pekaputjeti.} {\footnotesize (From Dan. \ipa{pe- + kaput + -jeti}).} v. to lead, to head, to be at head.
	

	
		\key{pelman.} {\footnotesize (From Dan. \ipa{pelus + -man}).} n. fruit.
	

	
		\key{pelnom.} {\footnotesize (From PIE \ipa{*pel}).} n. skin.
	

	
		\key{pelus.} {\footnotesize (From PIE \ipa{*pélh₁-u-}).} pron. much; a lot; plenty.
	

	
		\key{pelvjés.} {\footnotesize (From Dan. \ipa{pelus}).} adj.
			\textbf{1.} bigger; major.
			\textbf{2.} older.
	

	
		\key{pemanjeti.} {\footnotesize (From Dan. \ipa{pe- + mánjetor}).} v.
			\textbf{1.} to remind (in the active voice. Use dative case for the person being reminded).
			\textbf{2.} to remember (in the passive voice).
	

	
		\key{pempe.} {\footnotesize (From PIE \ipa{*pénkʷe}).} num. five.
	

	
		\key{pempedesam.} {\footnotesize (From PIE \ipa{*pénkʷe + *déḱm̥}).} num. fifteen.
	

	
		\key{pempedesamos.}  adj. fifteenth.
	

	
		\key{pempedsont.} {\footnotesize (From PIE \ipa{*kʷétwr̥dḱomt}).} num. fifty.
	

	
		\key{pempedsontos.} {\footnotesize (From PIE \ipa{*kʷétwr̥dḱomt}).} adj. fiftieth.
	

	
		\key{pemptos.}  adj. fifth.
	

	
		\key{peplos.} {\footnotesize (From PIE \ipa{*kʷékʷlos}).} m.
			\textbf{1.} wheel.
			\textbf{2.} circle.
	

	
		\key{pereti.} {\footnotesize (From PIE \ipa{*per-}).} v. to pierce, to run through.
	

	
		\key{peri.}  prep.
			\textbf{1.} (+gen) around.
			\textbf{2.} (+gen) about, concerning.
	

	
		\key{peris.} {\footnotesize (From PIE \ipa{*per-}).} f. fear.
	

	
		\key{Perkunos.} {\footnotesize (From PIE \ipa{*perkʷunos}).} m. Danetian god of thunder.
	

	\key{perkus,}
		\textit{parkves.} {\footnotesize (From PIE \ipa{*pérkus}).} m. oak.
	

	
		\key{Persia.}  f. Persia.
	

	
		\key{persna.} {\footnotesize (From PIE \ipa{*tpḗrsneh₂}).} f. heel, upper thigh.
	

	\key{persus,}
		\textit{parsves.} {\footnotesize (From PIE \ipa{*pérḱus}).} f.
			\textbf{1.} chest.
			\textbf{2.} rib.
	

	\key{pertus,}
		\textit{partves.} {\footnotesize (From PIE \ipa{*pértus}).} m. crossing.
	

	
		\key{peruti.} {\footnotesize (From PIE \ipa{*péruti}).} adv. last year.
	

	
		\key{Peruvia.}  f. Peru.
	

	\key{pervan,}
		\textit{parvens, prumos.} {\footnotesize (From PIE \ipa{*pérwr̥}).} n. ford, crossing.
	

	
		\key{pesnom.} {\footnotesize (From PIE \ipa{*pes-}).} n.{\footnotesize \textit{ vulgar.}} penis.
	

	\key{pesu,}
		\textit{pasves.} {\footnotesize (From PIE \ipa{*péḱu}).} n. livestock.
	

	\key{petan,}
		\textit{patens.} {\footnotesize (From PIE \ipa{*péth₂r̥}).} n. wing.
	

	
		\key{Petros.} {\footnotesize (From AG \gr{Πέτρος}).} m. Peter, Pedro.
	

	
		\key{petvardsont.} {\footnotesize (From PIE \ipa{*kʷétwr̥dḱomt}).} num. forty.
	

	
		\key{petvardsontos.} {\footnotesize (From PIE \ipa{*kʷétwr̥dḱomt}).} adj. fortieth.
	

	
		\key{petvártos.}  adj. fourth.
	

	
		\key{petvordesam.} {\footnotesize (From PIE \ipa{*kʷetwóres + *déḱm̥}).} num. fourteen.
	

	
		\key{petvordesamos.}  adj. fourteenth.
	

	
		\key{petvóres.} {\footnotesize (From PIE \ipa{*kʷetwóres}).} num. four.
	

	
		\key{phi.}  n. Greek letter phi.
	

	
		\key{Philippos.} {\footnotesize (From AG \gr{Φῐ́λῐππος}).} m. Philip, Felipe.
	

	
		\key{philosophia.} {\footnotesize (From Lat. \ipa{philosophia}, from AG \gr{φιλοσοφία}).} f. philosophy.
	

	
		\key{phosphoros.}  m.{\footnotesize \textit{ chemistry.}} phosphorus.
	

	
		\key{physica.} {\footnotesize (From Lat. \ipa{physica}, from AG \gr{φυσική}).} f. physics.
	

	
		\key{pi.}  n.
			\textbf{1.} Greek letter pi.
			\textbf{2.}{\footnotesize \textit{ mathematics.}} ratio between the circumference of a circle and its diameter.
	

	
		\key{piboti.} {\footnotesize (From PIE \ipa{*píph₃eti}, with dissimilation of the second \ipa{p}).} v. to drink.
	

	
		\key{pisda.} {\footnotesize (From PIE \ipa{*písdeh₂}).} f.{\footnotesize \textit{ vulgar.}} vulva.
	

	
		\key{pisjeti.} {\footnotesize (From PIE \ipa{*peyḱ-}).} v. to write.
	

	
		\key{pistis.} {\footnotesize (From PIE \ipa{*peysḱ-}).} m.
			\textbf{1.} fish.
			\textbf{2.} Pisces (in the plural, constellation).
	

	
		\key{planéta.} {\footnotesize (From Lat. \ipa{planeta}, from AG \gr{πλανήτης}).} m. planet.
	

	
		\key{planos.} {\footnotesize (From PIE \ipa{*pl̥h₁nós}).} adj. full.
	

	
		\key{platína.}  f.{\footnotesize \textit{ chemistry.}} platinum.
	

	
		\key{pleníeti.} {\footnotesize (From Dan. \ipa{planos}).} v. to fill.
	

	
		\key{pletia.} {\footnotesize (From Dan. \ipa{pletus}).} adj.
			\textbf{1.} square, plaza.
			\textbf{2.}{\footnotesize \textit{ mathematics.}} area of a shape or figure.
	

	
		\key{pletos.} {\footnotesize (From PIE \ipa{*pléth₂os}).} n. breadth.
	

	
		\key{pletus.} {\footnotesize (From PIE \ipa{*pléth₂us}).} adj. flat, broad.
	

	\key{plotus,}
		\textit{platves.} {\footnotesize (From PIE \ipa{*pléh₃tus}).} m. flood.
	

	
		\key{pludeti.} {\footnotesize (From PIE \ipa{*plewd-}).} v. to swim.
	

	
		\key{pluksna.} {\footnotesize (From PIE \ipa{*plúnksneh₂}).} f. feather.
	

	\key{plumón,}
		\textit{palvamnes, plumanmos.} {\footnotesize (From PIE \ipa{*pléwmō}).} m. lung.
	

	
		\key{plusis.} {\footnotesize (From PIE \ipa{*plúsis}).} f. flea.
	

	
		\key{pluti.} {\footnotesize (From PIE \ipa{*pléw-e-ti}).} v. to rain.
	

	
		\key{plutoniom.}  n.{\footnotesize \textit{ chemistry.}} plutonium.
	

	
		\key{plutos.} {\footnotesize (From PIE \ipa{*plutós}).} adj.
			\textbf{1.} washed.
			\textbf{2.} floated.
	

	
		\key{podegos.} {\footnotesize (From Dan. \ipa{po + dege}).} pron. each.
	

	
		\key{Podgoritsa.} {\footnotesize (From Serbo-Croatian \ipa{Pòdgorica} / \cyr{По̀дгорица}).} f. Podgorica, capital of Montenegro.
	

	
		\key{podi.} {\footnotesize (From Dan. \ipa{po- + -di}. The suffix \ipa{-di} comes from PIE \ipa{*dʰi}).} proadv. (interrogative) where.
	

	
		\key{podi.}  proadv. (interrogative) how much.
	

	\key{pods,}
		\textit{pedes.} {\footnotesize (From PIE \ipa{*pṓds}).} m. foot.
	

	
		\key{poiman.} {\footnotesize (From PIE \ipa{*poh₂-i-mn̥}).} n.
			\textbf{1.} salvation.
			\textbf{2.} cover, lid.
	

	
		\key{poimos.}  proadv.
			\textbf{1.} (interrogative) to where, whither.
			\textbf{2.} (interrogative) to what end, why.
	

	
		\key{poina.} {\footnotesize (From PIE \ipa{*kʷoynéh₂}).} f.
			\textbf{1.} payment.
			\textbf{2.} punishment, vengeance.
	

	
		\key{poives.} {\footnotesize (From Dan. \ipa{po- + -ives}. The suffix \ipa{-ives} comes from \ipa{juves}, the genitive of \ipa{oju}).} proadv. (interrogative) when.
	

	
		\key{polar.} {\footnotesize (From Lat. \ipa{polaris}).} adj. polar.
	

	
		\key{Polonia.}  f. Poland.
	

	
		\key{poloniom.}  n.{\footnotesize \textit{ chemistry.}} polonium.
	

	\key{ponts,}
		\textit{pantes.} {\footnotesize (From PIE \ipa{*póntoh₁s}).} m. path, way.
	

	
		\key{pornom.} {\footnotesize (From PIE \ipa{*pornóm}).} n. fern, leaf.
	

	
		\key{porsos.} {\footnotesize (From PIE \ipa{*pórḱos}).} m. pig, piglet, hog, swine.
	

	
		\key{Porto-Rico.}  n. Puerto Rico.
	

	
		\key{Portugal.}  f. Portugal.
	

	
		\key{portugaliska.} {\footnotesize (From Dan. \ipa{Portugalia}).} f. Portuguese language.
	

	
		\key{pos.} {\footnotesize (From PIE \ipa{*pós}).} prep.
			\textbf{1.} (+loc) behind, at the back side of (stationary, answers the question "where?").
			\textbf{2.} (+loc) after (temporal).
			\textbf{3.} (+acc) behind (with movement, answers the question "where to?").
	

	
		\key{posjolik.} {\footnotesize (From a Turkic language; compare Uzbek \ipa{podsholik}).} m. kingdom.
	

	
		\key{posmad.}  proadv. (interrogative) from where, whence.
	

	
		\key{posme.}  proadv. (interrogative) how, in what manner.
	

	
		\key{potíeti.} {\footnotesize (From Dan. \ipa{potis}).} v.
			\textbf{1.} (with accusative) to govern, to administer.
			\textbf{2.} (with \ipa{úper}) to rule over, to govern, to dominate.
	

	
		\key{potis.} {\footnotesize (From PIE \ipa{*pótis}).} m.
			\textbf{1.} master, ruler.
			\textbf{2.} husband.
			\textbf{3.} lord.
	

	
		\key{potodi.} {\footnotesize (From Dan. \ipa{poto- + -di}. The suffix \ipa{-di} comes from PIE \ipa{*dʰi}).} proadv. (existential) somewhere.
	

	
		\key{potoimos.}  proadv.
			\textbf{1.} (existential) to somewhere.
			\textbf{2.} (existential) for some reason.
	

	
		\key{potoives.} {\footnotesize (From Dan. \ipa{poto- + -ives}. The suffix \ipa{-ives} comes from \ipa{juves}, the genitive of \ipa{oju}).} proadv. (existential) some time, sometimes.
	

	
		\key{potosmad.}  proadv. (existential) from somewhere.
	

	
		\key{potosme.}  proadv. (existential) in some way, in some manner.
	

	
		\key{pototi.}  proadv. (existential) by some amount, in some amount.
	

	
		\key{Praga.} {\footnotesize (From Lat. \ipa{Praga}).} f. Prague.
	

	
		\key{praisident.} {\footnotesize (From Lat. \ipa{praesidēns}).} m. president.
	

	
		\key{praman.}  n. beginning, start.
	

	
		\key{pramos.}  adj. first.
	

	
		\key{praseodymiom.}  n.{\footnotesize \textit{ chemistry.}} praseodymium.
	

	
		\key{preseti.} {\footnotesize (From PIE \ipa{*preḱ-}).} v.
			\textbf{1.} to ask, to question; (Use the dative case for the person receiving the question.).
			\textbf{2.} to ask for, to petition, to request; (Use the ablative case for the person receiving the request.).
	

	
		\key{pretiops.} {\footnotesize (From PIE \ipa{*préti + *h₃ókʷs}).} n. face.
	

	
		\key{prios.} {\footnotesize (From PIE \ipa{*priHós}).} adj.
			\textbf{1.} dear, beloved.
			\textbf{2.} happy, free.
	

	
		\key{pro.} {\footnotesize (From PIE \ipa{*pró}).} prep.
			\textbf{1.} (+loc) in front of (spatial; stationary, answers the question "where?").
			\textbf{2.} (+loc) before (temporal).
			\textbf{3.} (+acc) in front of (spatial; with movement, answers the question "where to?").
			\textbf{4.} (+acc) according to.
			\textbf{5.} (+dat) on behalf of, for the sake of, for the benefit of.
			\textbf{6.} (+dat) so that, for the purpose of.
	

	
		\key{probereti.} {\footnotesize (From Dan. \ipa{pro + bereti}).} v. to carry forth, to bring forth.
	

	
		\key{proíti.} {\footnotesize (From Dan. \ipa{pro- + iti}).} v.
			\textbf{1.} to occur, to happen.
			\textbf{2.} (of days, weeks, years, etc.) to pass.
	

	
		\key{promenáda.} {\footnotesize (From Fr. \ipa{promenade}).} f. walk, stroll, walk for enjoyment.
	

	
		\key{promethiom.}  n.{\footnotesize \textit{ chemistry.}} promethium.
	

	
		\key{protactiniom.}  n.{\footnotesize \textit{ chemistry.}} protactinium.
	

	
		\key{proti.} {\footnotesize (From PIE \ipa{*pr-ó-ti}).} prep.
			\textbf{1.} (+gen) against.
			\textbf{2.} (+dat) towards, in the direction of, facing towards.
	

	
		\key{protiaerios.}  m. anti-aircraft.
	

	
		\key{protislabjeti.} {\footnotesize (From Dan. \ipa{slabyeti}).} v. to understand, to comprehend.
	

	
		\key{psi.}  n. Greek letter psi.
	

	
		\key{psycha.} {\footnotesize (From AG \ipa{ψῡχή}).} f. soul.
	

	
		\key{psychologia.} {\footnotesize (From Lat. \ipa{psychologia}).} f. psychology.
	

	
		\key{pulnis.} {\footnotesize (From PIE \ipa{*pulh₂-}).} f. hair.
	

	
		\key{puros.} {\footnotesize (From PIE \ipa{*puHrós}).} m. wheat.
	

	
		\key{qu.}  n. letter \textit{q}.
	

	
		\key{quantiska mechanica.}  f. quantum mechanics.
	

	
		\key{raderfordiom.}  n.{\footnotesize \textit{ chemistry.}} rutherfordium.
	

	
		\key{radiom.}  n.{\footnotesize \textit{ chemistry.}} radium.
	

	
		\key{radón.}  m.{\footnotesize \textit{ chemistry.}} radon.
	

	
		\key{rads.} {\footnotesize (From PIE \ipa{*wréh₂ds}).} f. root.
	

	
		\key{Raikstag.} {\footnotesize (From Ger. \ipa{Reichstag}).} m. Reichstag.
	

	
		\key{Raimondos.} {\footnotesize (Equivalent to En. and Fr. \ipa{Raymond}).} m. Raymond, Ramón.
	

	
		\key{Raján.} {\footnotesize (From En. \ipa{Ryan}).} m. Ryan.
	

	
		\key{rama.} {\footnotesize (From OHG \ipa{rama}).} f. frame (of a building), framework.
	

	
		\key{rebos.} {\footnotesize (From PIE \ipa{*h₁régʷos}).} n. darkness.
	

	
		\key{redissa.} {\footnotesize (From Dan. \ipa{reds}).} f. queen.
	

	
		\key{reds.} {\footnotesize (From PIE \ipa{*h₃rḗǵs}).} m. king, ruler.
	

	
		\key{Reicjavik.}  From Icelandic \ipa{Reykjavík}. Reykjavik, capital of Iceland.
	

	
		\key{reja.} {\footnotesize (From PIE \ipa{*reh₁is}).} f. wealth, goods, riches.
	

	
		\key{remos.} {\footnotesize (From PIE \ipa{*h₁réh₁}).} adj. alone, lonely.
	

	
		\key{Renvaldos.} {\footnotesize (Ultimately from PG \ipa{*Raginawaldaz}).} m. Ronald, Reginald, Reynold, Reinhold.
	

	
		\key{respublica.} {\footnotesize (From Lat. \ipa{rēspūblica}).} f. republic.
	

	
		\key{restos.} {\footnotesize (From PIE \ipa{*h₃reǵtós}).} adj. straightened, right, upright.
	

	
		\key{retos.} {\footnotesize (From PIE \ipa{*h₁réh₁}).} adj. empty.
	

	
		\key{rheniom.}  n.{\footnotesize \textit{ chemistry.}} rhenium.
	

	
		\key{Rhenos.}  m. Rhine.
	

	
		\key{rho.}  n. Greek letter rho.
	

	
		\key{rhodiom.}  n.{\footnotesize \textit{ chemistry.}} rhodium.
	

	
		\key{Richardos.} {\footnotesize (From Lat. \ipa{Richardus}).} m. Richard, Ricardo.
	

	
		\key{Richardos Philippos Fainman.}  m. Richard Phillips Feynman.
	

	
		\key{Riga.} {\footnotesize (From Rus. \cyr{Ри́га}).} f. Riga, capital of Latvia.
	

	
		\key{Robértos.} {\footnotesize (From Lat. \ipa{Robertus}).} m. Robert.
	

	
		\key{rodetis.} {\footnotesize (From Dan. \ipa{rodia}).} f. government.
	

	
		\key{rodia.} {\footnotesize (From PIE \ipa{*h₃reǵ-}).} f.
			\textbf{1.} law.
			\textbf{2.} (law) right.
	

	
		\key{rodis.} {\footnotesize (From Dan. \ipa{reds}).} f.
			\textbf{1.} rule (regla).
			\textbf{2.} rule (reino).
	

	
		\key{rodukter.}  m. judge.
	

	
		\key{roinos.} {\footnotesize (From PIE \ipa{*roypnós}).} m. rope.
	

	
		\key{rointgeniom.}  n.{\footnotesize \textit{ chemistry.}} roentgenium.
	

	
		\key{Roma.}  f. Rome.
	

	
		\key{Romania.}  f. Romania.
	

	
		\key{Romiskom Imperiom.} {\footnotesize (From Dan. \ipa{romiskom + imperiom}).} n. Roman Empire.
	

	
		\key{Rossia.} {\footnotesize (From BG \gr{Ῥωσία}).} f. Russia.
	

	
		\key{rossiska.} {\footnotesize (From Dan. \ipa{Rossia}).} f. Russian language.
	

	
		\key{Rossiska Foideratión.}  f. Russian Federation.
	

	
		\key{rossiskos.} {\footnotesize (From Dan. \ipa{Rossia}).} adj. Russian.
	

	
		\key{rubidiom.}  n.{\footnotesize \textit{ chemistry.}} rubidium.
	

	
		\key{Rufra Orda.}  f. Red Army.
	

	
		\key{Rufra Pletia.}  f. Red Square.
	

	
		\key{rufros.} {\footnotesize (From PIE \ipa{*h₁rudʰrós}).} adj. red.
	

	
		\key{rugis.} {\footnotesize (From PIE \ipa{*rúgʰis}).} m. rye.
	

	
		\key{rutheniom.}  n.{\footnotesize \textit{ chemistry.}} ruthenium.
	

	
		\key{sadeti.} {\footnotesize (From PIE \ipa{*ḱh₂d-}).} v. to fall, to fall down.
	

	
		\key{sagjeman.} {\footnotesize (From Dan. \ipa{sagjeti}).} n.
			\textbf{1.} flag, banner.
			\textbf{2.} sign.
	

	
		\key{sagjeti.} {\footnotesize (From PIE \ipa{*séh₂gyeti}).} v.
			\textbf{1.} to give a sign.
			\textbf{2.} to mark.
			\textbf{3.} to mean, to signify.
	

	
		\key{sagjetis.} {\footnotesize (From Dan. \ipa{sagjeti}).} f.
			\textbf{1.} meaning, the meaning of something.
			\textbf{2.} a sign or a mark.
	

	
		\key{sals.} {\footnotesize (From PIE \ipa{*séh₂ls}).} m. salt.
	

	
		\key{samariom.}  n.{\footnotesize \textit{ chemistry.}} samarium.
	

	
		\key{Samuel.} {\footnotesize (From AG \gr{Σαμουήλ}).} m. Samuel.
	

	
		\key{Sancta Bíblia.} {\footnotesize (From Lat. \ipa{Sancta Biblia}. $\blacklozenge$~This term is in plural).} n. holy books; Holy Bible.
	

	
		\key{Sancta Petropolis.} {\footnotesize (From Lat. \ipa{sancta + Petropolis}).} f. Saint Petersburg.
	

	
		\key{sanctos.} {\footnotesize (From Lat. \ipa{sānctus}).} adj. saint, holy, sacred.
	

	
		\key{Sanctos Joánna.}  m.
			\textbf{1.} Saint John, John the Apostle.
			\textbf{2.} city of San Juan, capital of Puerto Rico.
	

	
		\key{Sanctos Marínos.} {\footnotesize (From It. \ipa{San Marino}).} m. San Marino.
	

	
		\key{Sarajevo.}  n. Sarajevo, capital of Bosnia and Herzegovina.
	

	
		\key{sarsos.} {\footnotesize (From PIE \ipa{*ḱr̥sós}).} m. vehicle.
	

	
		\key{satenjos.} {\footnotesize (From PIE \ipa{*ḱm̥tóm}).} adj. hundredth.
	

	
		\key{satjeti.} {\footnotesize (From PG \ipa{satjaną}).} v. to set, to place down.
	

	
		\key{satón.} {\footnotesize (From PIE \ipa{*ḱm̥tóm}).} num. hundred; one hundred.
	

	
		\key{Satúrnos.}  m.
			\textbf{1.} Saturn (the god).
			\textbf{2.} Saturn (the planet).
	

	
		\key{scandiom.}  n.{\footnotesize \textit{ chemistry.}} scandium.
	

	
		\key{schola.} {\footnotesize (From Lat. \ipa{schola}, from AG \gr{σχολή}).} f. school.
	

	
		\key{scientia.} {\footnotesize (From Lat. \ipa{scientia}).} f. science.
	

	
		\key{scorpios.}  m.
			\textbf{1.} scorpion.
			\textbf{2.} Scorpius (constellation).
	

	
		\key{Scotia.}  f. Scotland.
	

	
		\key{Scupoi.} {\footnotesize (From Lat. \ipa{Scūpī}).} m. Skopje, capital of North Macedonia.
	

	
		\key{Sebastopolis.} {\footnotesize (From AG \gr{Σεβαστόπολις}).} f. Sevastopol.
	

	
		\key{sebastopoliskos.} {\footnotesize (From Dan. \ipa{Sebastopolis}).} adj. pertaining to Sevastopol.
	

	
		\key{secundom.} {\footnotesize (From Lat. \ipa{secundum}).} n. second (unit of measurement).
	

	
		\key{sédjetor.} {\footnotesize (From PIE \ipa{*sed-}).} v. to sit, to sit down.
	

	
		\key{sedos.} {\footnotesize (From PIE \ipa{*séǵʰos}).} n. power, strength.
	

	
		\key{seleniom.}  n.{\footnotesize \textit{ chemistry.}} selenium.
	

	
		\key{selos.} {\footnotesize (From PIE \ipa{*sélos}).} n. marsh.
	

	
		\key{selva.} {\footnotesize (From OF \ipa{selve}).} f. forest, thick forest.
	

	\key{seman,}
		\textit{samens.} {\footnotesize (From PIE \ipa{*séh₁mn̥}).} n.
			\textbf{1.} seed.
			\textbf{2.} semen.
	

	
		\key{semi.} {\footnotesize (From PIE \ipa{*sēmi}).} adv. half, partially.
	

	
		\key{semia.} {\footnotesize (From PIE \ipa{*sēmi}).} f. half.
	

	
		\key{sénjetor.} {\footnotesize (From Dan. \ipa{senos}).} v. to get older, to grow old.
	

	
		\key{senos.} {\footnotesize (From PIE \ipa{*sénos}).} adj. old.
	

	
		\key{sep.}  conj. so, thus, then.
	

	
		\key{septam.} {\footnotesize (From PIE \ipa{*septḿ̥}).} num. seven.
	

	
		\key{septamos.}  adj. seventh.
	

	
		\key{septandesam.} {\footnotesize (From PIE \ipa{*septḿ̥ + *déḱm̥}).} num. seventeen.
	

	
		\key{septandesamos.}  adj. seventeenth.
	

	
		\key{septandsont.} {\footnotesize (From PIE \ipa{*septḿ̥dḱomt}).} num. seventy.
	

	
		\key{septandsontos.} {\footnotesize (From PIE \ipa{*septḿ̥dḱomt}).} adj. seventieth.
	

	
		\key{september.} {\footnotesize (From Lat. \ipa{September}).} m.{\footnotesize \textit{ month.}} September.
	

	
		\key{septoives.} {\footnotesize (From Dan. \ipa{sep + toives}).} conj. therefore, thus, hence.
	

	
		\key{Serbia.}  f. Serbia.
	

	\key{serd,}
		\textit{sardes.} {\footnotesize (From PIE \ipa{*ḱḗr}).} n. heart.
	

	
		\key{sergjant.} {\footnotesize (From Fr. \ipa{sergent}).} m. sergeant.
	

	\key{sersón,}
		\textit{krasnes, krasanmos.} {\footnotesize (From PIE \ipa{*ḱérh₂sō}).} m. head, skull.
	

	
		\key{sestos.} {\footnotesize (From PIE \ipa{*sedtós}).} adj. seated.
	

	
		\key{setoi.} {\footnotesize (From PIE {*ḱe-}).} proadv. here.
	

	
		\key{siborgiom.}  n.{\footnotesize \textit{ chemistry.}} seaborgium.
	

	
		\key{siesj.} {\footnotesize (From Fr. \ipa{siège}).} m. siege.
	

	
		\key{sigma.}  f. Greek letter sigma.
	

	
		\key{siliciom.}  n.{\footnotesize \textit{ chemistry.}} silicon.
	

	
		\key{simap.} {\footnotesize (From a Turkic language; compare Uzbek \ipa{simob}, Uyghur \ipa{simab}, and Turkmen \ipa{simap}).} m.{\footnotesize \textit{ chemistry.}} mercury (metal).
	

	
		\key{sivos.} {\footnotesize (Unknown).} adj. good.
	

	
		\key{sjar.} {\footnotesize (From Per. \arab{شهر}).} m. city.
	

	
		\key{sjofer.} {\footnotesize (From Fr. \ipa{chauffeur}).} m. chauffeur, driver.
	

	
		\key{skula.} {\footnotesize (From PG \ipa{*skulō}).} f.
			\textbf{1.} troop; group; crowd.
			\textbf{2.} multitude.
	

	
		\key{slabjeti.} {\footnotesize (From PIE \ipa{*sl̥h₂gʷ-}).} v. to seize, to grasp, to take.
	

	
		\key{slevos.} {\footnotesize (From PIE \ipa{*ḱléwos}).} n. fame.
	

	
		\key{slita.} {\footnotesize (From PIE \ipa{*ḱléyteh₂}).} f. slope, incline.
	

	
		\key{slitos.} {\footnotesize (From PIE \ipa{*ḱlitós}).} adj. covered, sheltered.
	

	
		\key{slitra.} {\footnotesize (From PIE \ipa{*ḱléytreh₂}).} f. shelter, covering.
	

	
		\key{slivos.} {\footnotesize (From PIE \ipa{*ḱleywós}).} adj. sloped, curved.
	

	
		\key{Slovakia.} {\footnotesize (From Rus. \ipa{Слова́кия}).} f. Slovakia.
	

	
		\key{Slovenía.}  f. Slovenia.
	

	
		\key{slunis.} {\footnotesize (From PIE \ipa{*ḱlównis}).} f. buttocks, kip.
	

	
		\key{slutos.} {\footnotesize (From PIE \ipa{*ḱlutós}).} adj.
			\textbf{1.} heard, listened.
			\textbf{2.} known, famous.
	

	
		\key{smerdeti.} {\footnotesize (From PIE \ipa{*(s)merd-}).} v. to bite.
	

	
		\key{smernos.} {\footnotesize (From PIE \ipa{*smerd-nós}).} m. terrible, dreadful, horrible.
	

	\key{smerus,}
		\textit{smarves.} {\footnotesize (From PIE \ipa{*smérus}).} m. grease.
	

	\key{snetan,}
		\textit{snatens.} {\footnotesize (From PIE \ipa{*(s)néHtr̥}).} n. adder (type of snake).
	

	\key{snevan,}
		\textit{snavens.} {\footnotesize (From PIE \ipa{*snéh₁wr̥}).} n. sinew, tendon.
	

	
		\key{snoivos.} {\footnotesize (From PIE \ipa{*snóygʷʰos}).} m. snow.
	

	
		\key{so-.} {\footnotesize (From PIE \ipa{*ḱo}).} pref. here.
	

	
		\key{socialismos.}  m. socialism.
	

	
		\key{socialísta.} {\footnotesize (From Dan. \ipa{socialismos}).} adj. socialist.
	

	
		\key{socialistiskos.} {\footnotesize (From Fr. \ipa{socialiste}).} adj. socialist.
	

	
		\key{sodi.} {\footnotesize (From Dan. \ipa{so- + -di}. The suffix \ipa{-di} comes from PIE \ipa{*dʰi}).} proadv. (proximal) this place, here.
	

	
		\key{sodínes.} {\footnotesize (From PIE {*ḱo-} and Dan. \ipa{din}).} adv. today.
	

	
		\key{soia.} {\footnotesize (From PIE \ipa{*(s)ḱeh₃ih₂}).} f. shadow, shade.
	

	
		\key{soimos.}  proadv.
			\textbf{1.} (proximal) to this place, to here.
			\textbf{2.} (proximal) for this reason.
	

	
		\key{soitos.} {\footnotesize (From PIE \ipa{*soytós}).} m. magic.
	

	
		\key{soives.} {\footnotesize (From Dan. \ipa{so- + -ives}. The suffix \ipa{-ives} comes from \ipa{juves}, the genitive of \ipa{oju}).} proadv. (proximal) at this time, now.
	

	
		\key{Sol.}  m.
			\textbf{1.} Sun (the god).
			\textbf{2.} the Sun.
	

	
		\key{soldátos.}  m. soldier.
	

	
		\key{solnos.} {\footnotesize (From PIE \ipa{*solh₂-}).} pron. whole (in singular), all (in plural).
	

	
		\key{som.} {\footnotesize (From PIE \ipa{*ḱóm}).} prep. (+ins) with.
	

	
		\key{somblesnos.} {\footnotesize (From Dan. \ipa{som + blesnos}).} m. seduction, temptation.
	

	
		\key{Somjugestoi Djoletes Americas.}  m. United States of America.
	

	
		\key{somjugestos.} {\footnotesize (From Dan. \ipa{somjugos}).} adj. united.
	

	
		\key{somjugeti.} {\footnotesize (From Dan. \ipa{somjugos}).} v.
			\textbf{1.} to unite (transitive).
			\textbf{2.} to gather.
	

	
		\key{somjugos.} {\footnotesize (From Dan. \ipa{som- + jugos}).} n. union.
	

	
		\key{Somjugos Sovietiskóm Socialistiskóm Respublicóm.}  n. Union of Soviet Socialist Republics.
	

	
		\key{sondenés.} {\footnotesize (From Dan. \ipa{som + denos}).} adj. related.
	

	
		\key{sonféretor.} {\footnotesize (From Dan. \ipa{bereti}).} v. to behave.
	

	
		\key{sonfis.} {\footnotesize (From PIE \ipa{*som-dʰh₁-is}, from \ipa{*dʰeh₁-}).} m. judge.
	

	
		\key{sontíeti.} {\footnotesize (From PIE \ipa{*sont-éye-}).} v. to send.
	

	
		\key{sontis.} {\footnotesize (From Dan. \ipa{esti}).} f. essence.
	

	
		\key{sonvistiés.} {\footnotesize (From Dan. \ipa{vistis}).} adj.
			\textbf{1.} messenger.
			\textbf{2.} angel.
	

	
		\key{Sophia.} {\footnotesize (From AG \gr{σοφία}).} f.
			\textbf{1.} Sophia, Sofia, Sophie.
			\textbf{2.} Sofia, capital of Bularia.
	

	
		\key{sopos¹.} {\footnotesize (From PIE \ipa{*sokʷos}).} m. juice.
	

	
		\key{sopos².} {\footnotesize (From PIE \ipa{*ḱoph₂ós}).} m. hoof.
	

	\key{sosan,}
		\textit{sasens.} {\footnotesize (From PIE \ipa{*sóḱr̥}).} n. dung, excrement.
	

	
		\key{sosmad.}  proadv. (proximal) from this place, from here.
	

	
		\key{sosme.}  proadv. (proximal) in this way, in this manner.
	

	
		\key{sosna.} {\footnotesize (From PIE \ipa{*(s)h₁es-}).} f. autumn.
	

	
		\key{soti.}  proadv. (proximal) this much.
	

	\key{sovan,}
		\textit{savens.} {\footnotesize (From PIE \ipa{*sóh₂wl̥}).} n. sun.
	

	
		\key{Sovietiskom Somjugos.}  n. Soviet Union.
	

	
		\key{sovietiskos.} {\footnotesize (From Ru. \cyr{сове́т}).} adj. Soviet (pertaining to the Soviet Union).
	

	
		\key{spalos.} {\footnotesize (From PIE \ipa{*(s)kʷálos}).} m.
			\textbf{1.} large fish.
			\textbf{2.} sheatfish.
	

	\key{speks,}
		\textit{spakes.} {\footnotesize (From PIE \ipa{*spéḱs}).} m.
			\textbf{1.} watcher.
			\textbf{2.} spy.
	

	
		\key{sper.} {\footnotesize (From PIE \ipa{*spḗr}).} m. sparrow.
	

	
		\key{spesa.} {\footnotesize (From PIE \ipa{*spéh₁s}).} f. prosperity, success.
	

	
		\key{sphaira.} {\footnotesize (From AG \gr{σφαῖρᾰ}).} f. sphere.
	

	
		\key{stadis.} {\footnotesize (From PG \ipa{*stadiz}).} m. place, location.
	

	
		\key{staman.} {\footnotesize (From PIE \ipa{*stéh₂mn̥}).} n. that which stands, stature.
	

	
		\key{stanga.} {\footnotesize (From PG \ipa{*stangō}).} f. bar, rod, stake, pole, beam.
	

	
		\key{stati.} {\footnotesize (From Dan. \ipa{statos}).} v. to stand.
	

	
		\key{statis.} {\footnotesize (From PIE \ipa{*stéh₂tis}).} f.
			\textbf{1.} standing, position.
			\textbf{2.} place, location.
	

	
		\key{statos.} {\footnotesize (From PIE \ipa{*sth₂tós}).} adj. placed, standing.
	

	
		\key{stelos.} {\footnotesize (From PIE \ipa{*stelh₃ós}).} adj. wide.
	

	
		\key{Steloslevés.} {\footnotesize (From Dan. \ipa{stelom + slevos}).} m. Danetian personal name, meaning \textit{wide fame}.
	

	\key{sten,}
		\textit{stanes.} {\footnotesize (From PIE \ipa{*pstḗn}).} m. breast, teat.
	

	
		\key{Stephanos.} {\footnotesize (From AG \gr{Στέφανος}).} m. Steven, Stephen, Stephanus.
	

	
		\key{ster.} {\footnotesize (From PIE \ipa{*h₂stḗr}).} m. star.
	

	
		\key{stibiom.} {\footnotesize (From Lat. \ipa{stibium}).} n.{\footnotesize \textit{ chemistry.}} antimony.
	

	
		\key{stoigos.} {\footnotesize (From PIE \ipa{*stóygʰos}).} m. path, way.
	

	
		\key{stratos.} {\footnotesize (From PIE \ipa{*str̥h₃tós}).} m. valley.
	

	\key{streps,}
		\textit{strapes.} {\footnotesize (From PIE \ipa{*(s)rḗps}).} f. turnip, rape, cabbage.
	

	
		\key{stridos.} {\footnotesize (From PIE \ipa{*strīdaz}).} m. dispute.
	

	
		\key{strontiom.}  n.{\footnotesize \textit{ chemistry.}} strontium.
	

	
		\key{stultos.} {\footnotesize (From PG \ipa{*stultaz}).} adj. proud.
	

	
		\key{sugeti.} {\footnotesize (From PG \ipa{*sugō}).} v. to sow; to scatter or disperse seeds.
	

	
		\key{Sundina America.} {\footnotesize (From Dan. \ipa{sundina + America}).} f.{\footnotesize \textit{ continent.}} South America.
	

	
		\key{Sundina Coréa.}  f. South Korea.
	

	
		\key{sundinos.} {\footnotesize (From Dan. \ipa{sundom}).} adj. southern, south, southerly.
	

	
		\key{sundom.} {\footnotesize (From PG \ipa{*sunþrą}).} n. south.
	

	
		\key{sunus.} {\footnotesize (From PIE \ipa{*suHnús}).} m. son.
	

	
		\key{suros.} {\footnotesize (From PIE \ipa{*súHros}).} adj. sour, acidic.
	

	
		\key{susnos.} {\footnotesize (From PIE \ipa{*h₂sews-}).} adj. dry.
	

	
		\key{svadus.} {\footnotesize (From PIE \ipa{*swéh₂dus}).} adj. sweet.
	

	
		\key{sveblos.} {\footnotesize (From PG \ipa{*sweblaz}).} m.{\footnotesize \textit{ chemistry.}} sulfur.
	

	
		\key{Svedia.} {\footnotesize (Related to Ger. \ipa{Schweden}).} f. Sweden.
	

	
		\key{svediska.} {\footnotesize (From Dan. \ipa{Svedia}).} f. Swedish language.
	

	
		\key{svekdesam.} {\footnotesize (From PIE \ipa{*swéḱs + *déḱm̥}).} num. sixteen.
	

	
		\key{svekdesamos.}  adj. sixteenth.
	

	
		\key{svekdsont.} {\footnotesize (From PIE \ipa{*swéḱsdḱomt}).} num. sixty.
	

	
		\key{svekdsontos.} {\footnotesize (From PIE \ipa{*swéḱsdḱomt}).} adj. sixtieth.
	

	
		\key{sveks.} {\footnotesize (From PIE \ipa{*swéḱs}).} num. six.
	

	
		\key{svektos.}  adj. sixth.
	

	
		\key{svenati.} {\footnotesize (From PIE \ipa{*swénh₂-ti}).} v. to make a sound.
	

	
		\key{svenis.} {\footnotesize (From Danetian \ipa{sve}).} f. image, likeness.
	

	
		\key{svenjíeti.} {\footnotesize (From PIE \ipa{*ḱwen-}).} v. to make holy, to hallow, to make sacred, to sanctify.
	

	
		\key{svenjos.} {\footnotesize (From PIE \ipa{*ḱwen-}).} adj. holy, sacred.
	

	
		\key{svennos.} {\footnotesize (From PIE \ipa{*supnós}).} m. sleep, slumber.
	

	
		\key{svenos.} {\footnotesize (From Dan. \ipa{svenati}).} m. sound.
	

	
		\key{svenseti.} {\footnotesize (From PIE \ipa{*ḱwen-}).} v. to make holy, to hallow, to make sacred, to sanctify.
	

	\key{svesor,}
		\textit{sustres, susarmos.} {\footnotesize (From PIE \ipa{*swésōr}).} f. sister.
	

	
		\key{svomia.} {\footnotesize (From PIE \ipa{*swé}).} f. freedom, liberty.
	

	
		\key{svomjos.} {\footnotesize (From Dan. \ipa{svomia}).} adj. free.
	

	\key{svon,}
		\textit{sunes, svanmos.} {\footnotesize (From PIE \ipa{*ḱwṓ}).} m. dog.
	

	
		\key{symbolom.} {\footnotesize (From Lat. \ipa{symbolum}, from AG \gr{σύμβολον}).} n. symbol.
	

	
		\key{synthesis.} {\footnotesize (From Lat. \ipa{synthesis}, from AG \gr{σύνθεσις}).} f. synthesis.
	

	
		\key{Talín.} {\footnotesize (From Rus. \cyr{Та́ллин}).} m. Tallinn, capital of Estonia.
	

	
		\key{tantaliom.}  n.{\footnotesize \textit{ chemistry.}} tantalum.
	

	
		\key{tantos.} {\footnotesize (From PIE \ipa{*tn̥tós}).} adj. stretched, extended.
	

	
		\key{tarnos.} {\footnotesize (From PIE \ipa{*tr̥nós}).} adj. sharp, stiff.
	

	
		\key{tau.}  n.
			\textbf{1.} Greek letter tau.
			\textbf{2.}{\footnotesize \textit{ mathematics.}} ratio between the circumference of a circle and its radius. Equivalent to two times pi.
	

	
		\key{tauros.} {\footnotesize (From PIE \ipa{*táwros}).} m.
			\textbf{1.} bull.
			\textbf{2.} Taurus (constellation).
	

	
		\key{te.}  n. letter \textit{t}.
	

	
		\key{technetiom.}  n.{\footnotesize \textit{ chemistry.}} technetium.
	

	
		\key{tegus.} {\footnotesize (From PIE \ipa{*tégus}).} adj. thick.
	

	\key{tektón,}
		\textit{taksnes, taksanmos.} {\footnotesize (From PIE \ipa{*tétḱō}).} m. carpenter.
	

	
		\key{telluriom.}  n.{\footnotesize \textit{ chemistry.}} tellurium.
	

	
		\key{temir.} {\footnotesize (From a Turkic language; compare Uzbek \ipa{temir}, Turkmen \ipa{demir}, Uyghur \ipa{tömür}, and Kazakh \cyr{темір}).} m.{\footnotesize \textit{ chemistry.}} iron.
	

	
		\key{tendeman.} {\footnotesize (From Dan. \ipa{tendeti}).} n. space.
	

	
		\key{tendeti.} {\footnotesize (From PIE \ipa{*tend-}).} v. to stretch, to extend.
	

	
		\key{tennessom.}  n.{\footnotesize \textit{ chemistry.}} tennessine.
	

	
		\key{tenos.} {\footnotesize (From PIE \ipa{*ténos}).} n. extension.
	

	\key{tentis,}
		\textit{tanties.} {\footnotesize (From PIE \ipa{*téntis}).} f. extension.
	

	
		\key{tenus.} {\footnotesize (From PIE \ipa{*ténh₂us}).} adj. thin.
	

	
		\key{tepos.} {\footnotesize (From PIE \ipa{*tépos}).} n. warmth, heat.
	

	
		\key{terbiom.}  n.{\footnotesize \textit{ chemistry.}} terbium.
	

	\key{terman,}
		\textit{tarmens.} {\footnotesize (From PIE \ipa{*térmn̥}).} n. boundary, end.
	

	
		\key{terribilis.} {\footnotesize (From Lat. \ipa{terribilis}).} adj. terrible.
	

	
		\key{thalliom.}  n.{\footnotesize \textit{ chemistry.}} thallium.
	

	
		\key{Theodoros.} {\footnotesize (From AG \gr{Θεόδωρος}).} m. Theodore.
	

	
		\key{theoreman.} {\footnotesize (From Lat. \ipa{theōrēma}, from AG \gr{θεώρημᾰ}).} n. theorem.
	

	
		\key{theoria.} {\footnotesize (From Lat. \ipa{theōria}, from AG \gr{θεωρῐ́ᾱ}).} f. theory.
	

	
		\key{thermodynamica.}  f. thermodynamics.
	

	
		\key{thermometrom.} {\footnotesize (From Fr. \ipa{thermomètre}, with a latinized ending).} n. thermometer.
	

	
		\key{theta.}  f. Greek letter theta.
	

	
		\key{Thoma.} {\footnotesize (From AG \gr{Θωμᾶς}).} m. Thomas.
	

	
		\key{thoriom.}  n.{\footnotesize \textit{ chemistry.}} thorium.
	

	
		\key{thuliom.}  n.{\footnotesize \textit{ chemistry.}} thulium.
	

	
		\key{Tiflis.} {\footnotesize (From Rus. \cyr{Тифли́с}).} f. Tbilisi, capital of Georgia.
	

	
		\key{Timotheos.} {\footnotesize (From AG \gr{Τῑμόθεος}).} m. Timothy.
	

	
		\key{Tirána.}  f. Tirana, capital of Albania.
	

	
		\key{titaniom.}  n.{\footnotesize \textit{ chemistry.}} titanium.
	

	
		\key{titta.} {\footnotesize (From PG \ipa{*titt-}).} f.
			\textbf{1.} breast, teat, tit.
			\textbf{2.} nipple.
	

	
		\key{todi.} {\footnotesize (From Dan. \ipa{to- + -di}. The suffix \ipa{-di} comes from PIE \ipa{*dʰi}).} proadv. (distal) that place, there.
	

	
		\key{toimos.}  proadv.
			\textbf{1.} (distal) to that place.
			\textbf{2.} (distal) for that reason.
	

	
		\key{toives.} {\footnotesize (From Dan. \ipa{to- + -ives}. The suffix \ipa{-ives} comes from \ipa{juves}, the genitive of \ipa{oju}).} proadv.
			\textbf{1.} (distal) at that time, then. conj.
			\textbf{2.} then.
	

	
		\key{tonos.} {\footnotesize (From PIE \ipa{*tónos}).} m. chord, fiber, string.
	

	
		\key{topographia.} {\footnotesize (From AG \gr{τοπογραφία}).} f. topography.
	

	
		\key{topologia.} {\footnotesize (From Fr. \ipa{topologie}, with a latinized ending).} f.{\footnotesize \textit{ mathematics.}} topology.
	

	
		\key{tosmad.}  proadv. (distal) from that place, from there.
	

	
		\key{tosme.}  proadv. (distal) in that manner, in that way.
	

	
		\key{toti.}  proadv. (distal) that much.
	

	
		\key{tractátos.} {\footnotesize (From Lat. \ipa{tractātus}).} m.
			\textbf{1.} treaty.
			\textbf{2.} treatise, tract.
			\textbf{3.} management, treatment.
			\textbf{4.} touching, handling, working, treating.
	

	
		\key{tremos.} {\footnotesize (From PIE \ipa{*trem-}).} m.
			\textbf{1.} trembling, tremor, shudder.
			\textbf{2.} fear, terror.
	

	
		\key{tres.} {\footnotesize (From PIE \ipa{*tréyes}).} num. three.
	

	
		\key{tresón.} {\footnotesize (From PIE \ipa{*tres-}).} adj.
			\textbf{1.} shy, timid.
			\textbf{2.} (of doves) timorous, shy, pavid.
	

	
		\key{tridesam.} {\footnotesize (From PIE \ipa{*tréyes + *déḱm̥}).} num. thirteen.
	

	
		\key{tridesamos.}  adj. thirteenth.
	

	
		\key{tridsont.} {\footnotesize (From PIE \ipa{*tridḱómt}).} num. thirty.
	

	
		\key{tridsontos.} {\footnotesize (From PIE \ipa{*tridḱómt}).} adj. thirtieth.
	

	
		\key{Trinidad i Tobago.}  f. Trinidad and Tobago.
	

	
		\key{tritos.}  adj. third.
	

	
		\key{tros.} {\footnotesize (From PIE \ipa{*trh₂os}).} prep. (+acc) through.
	

	
		\key{trosbereti.} {\footnotesize (From Dan. \ipa{tros + bereti}).} v. to finish, to complete.
	

	
		\key{trostanati.} {\footnotesize (From Dan. \ipa{tros- + stanati}).} v. to persevere.
	

	
		\key{tualét.} {\footnotesize (From Fr. \ipa{toilette}).} f. toilet.
	

	
		\key{Turcia.}  f. Turkey.
	

	
		\key{tusontis.} {\footnotesize (From PIE \ipa{*tuHsont-}).} num. thousand; one thousand.
	

	
		\key{tusontjos.} {\footnotesize (From PIE \ipa{*tuHsont-}).} adj. thousandth.
	

	
		\key{tuta.} {\footnotesize (From PIE \ipa{*tewtéh₂}).} f. people, tribe.
	

	
		\key{u.}  n. letter \textit{u}.
	

	
		\key{Ucrania.}  f. Ukraine.
	

	
		\key{ud.} {\footnotesize (From PIE \ipa{*úd}).} prep. (+abl) out from.
	

	
		\key{udan.} {\footnotesize (From PIE \ipa{*h₁ówdʰr̥}).} n. udder.
	

	
		\key{udbastia.}  f. exodus.
	

	
		\key{uderos.} {\footnotesize (From PIE \ipa{*úderos}).} m. abdomen, stomach, belly.
	

	
		\key{udlimanos.} {\footnotesize (From Dan. \ipa{linepti}).} adj. omitted; left out.
	

	
		\key{udros.} {\footnotesize (From PIE \ipa{*udrós}).} adj. of water, aquatic.
	

	
		\key{úper.} {\footnotesize (From PIE \ipa{*upér}).} prep.
			\textbf{1.} (+loc) above, over, on top of (stationary, answers the question "where?").
			\textbf{2.} (+acc) over, beyond, above (with movement, answers the question "where to?").
	

	
		\key{uperbereti.} {\footnotesize (From Dan. \ipa{úper + bereti}).} v.
			\textbf{1.} to raise (active).
			\textbf{2.} to rise, to be raised (passive).
	

	
		\key{uperjeros.}  adj. intercalary, epagomenal.
	

	
		\key{uperslevjos.} {\footnotesize (From Dan. \ipa{uperslevos}).} adj. glorious.
	

	
		\key{uperslevos.} {\footnotesize (From Dan. \ipa{uper- + slevos}).} n. glory.
	

	
		\key{upo.} {\footnotesize (From PIE \ipa{*upó}).} prep.
			\textbf{1.} (+loc) under, beneath (stationary, answers the question "where?").
			\textbf{2.} (+acc) under, beneath (with movement, answers the question "where to?").
	

	
		\key{upoíti.} {\footnotesize (From Dan. \ipa{upo- + iti}).} v. to go down, to come down, to descend, to get off.
	

	
		\key{uposme.} {\footnotesize (From Dan. \ipa{upo}).} conj.
			\textbf{1.} as, while.
			\textbf{2.} since, seeing as, because.
	

	
		\key{uraniom.}  n.{\footnotesize \textit{ chemistry.}} uranium.
	

	
		\key{urusj.} {\footnotesize (From a Turkic language; compare Uzbek \ipa{urush} and Turkmen \ipa{uruş}).} m. war.
	

	\key{us,}
		\textit{auses.} {\footnotesize (From PIE \ipa{*h₂ṓws}).} n. ear.
	

	
		\key{usterom mori.}  n. high sea, open sea.
	

	
		\key{Usteromoris.} {\footnotesize (From Dan. \ipa{usterom + mori}).} m. Danetian personal name, meaning \textit{high seas}.
	

	
		\key{usteros.} {\footnotesize (From PIE \ipa{*údteros}).} adj. higher, outer.
	

	
		\key{vacantia.} {\footnotesize (From Fr. \ipa{vacance}, with a latinized ending).} f.
			\textbf{1.} vacation.
			\textbf{2.} (in plural) holidays.
			\textbf{3.} vacancy, free position.
	

	
		\key{Vadus.} {\footnotesize (From Ger. \ipa{Vaduz}).} m. Vaduz, capital of Liechtenstein.
	

	
		\key{valdus.} {\footnotesize (From PG \ipa{*walþuz}).} m. forest.
	

	
		\key{Valeria.} {\footnotesize (From Lat. \ipa{Valeria}).} f. Valerie.
	

	
		\key{Valerios.} {\footnotesize (From Lat. \ipa{Valerius}).} m. Valerius, Valerio.
	

	
		\key{Valéta.}  f. Valletta, capital of Malta.
	

	
		\key{Valpobolos.} {\footnotesize (From Dan. \ipa{valpos + bolos}).} m. Danetian personal name, meaning \textit{strong wolf}.
	

	
		\key{Valponer.} {\footnotesize (From Dan. \ipa{valpos + ner}).} m. Danetian personal name, meaning \textit{powerful wolf}.
	

	
		\key{valpos.} {\footnotesize (From PIE \ipa{*wĺ̥kʷos}).} m.
			\textbf{1.} wolf.
			\textbf{2.} thief.
	

	
		\key{vanadiom.}  n.{\footnotesize \textit{ chemistry.}} vanadium.
	

	
		\key{vardjeti.} {\footnotesize (From Dan. \ipa{verdom²}).} v. to work.
	

	
		\key{vargos.} {\footnotesize (From PG \ipa{*wargaz}).} m.
			\textbf{1.} enemy.
			\textbf{2.} outlaw, criminal.
	

	
		\key{Varsovia.} {\footnotesize (From Lat. \ipa{Varsovia}).} f. Warsaw, capital of Poland.
	

	
		\key{varstos.} {\footnotesize (From PIE \ipa{*wr̥ǵtós}).} adj. worked, fashioned.
	

	
		\key{vassal.} {\footnotesize (From Fr. \ipa{vassal}).} m. vassal.
	

	
		\key{Vaticánom.}  n. Vatican City.
	

	
		\key{ve.}  n. letter \textit{v}.
	

	
		\key{vedenjeti.} {\footnotesize (From Dan. \ipa{vedenjos}).} v. to water.
	

	
		\key{vedenjos.} {\footnotesize (From Dan. \ipa{vodan}).} m.
			\textbf{1.} water-bearer, water-pourer.
			\textbf{2.} Aquarius (constellation).
	

	
		\key{veksperos.} {\footnotesize (From PIE \ipa{*wek(ʷ)speros}).} m. evening.
	

	
		\key{velocipods.} {\footnotesize (From Fr. \ipa{vélocipède}, influenced by Dan. \ipa{pods}).} m. velocipede; bicycle.
	

	
		\key{Venés.}  f.
			\textbf{1.} Venus (the goddess).
			\textbf{2.} Venus (the planet).
	

	\key{vents,}
		\textit{vantes.} {\footnotesize (From PIE \ipa{*h₂wéh₁n̥ts}).} m. wind.
	

	
		\key{vepti.} {\footnotesize (From PIE \ipa{*wekʷ-}).} v.
			\textbf{1.} to speak (when used in the imperfective aspect).
			\textbf{2.} to say (when used in the perfective aspect).
	

	
		\key{verdom¹.} {\footnotesize (From PIE \ipa{*werdʰh₁om}).} n. word.
	

	
		\key{verdom².} {\footnotesize (From PIE \ipa{*wérǵom}).} n. work.
	

	
		\key{verna.} {\footnotesize (From PIE \ipa{*wern-}).} f. alder.
	

	
		\key{veros.} {\footnotesize (From PIE \ipa{*weh₁ros}).} adj. True.
	

	
		\key{Versaljai.} {\footnotesize (From Fr. \ipa{Versailles}).} f. Versailles.
	

	
		\key{verster.} {\footnotesize (From Dan. \ipa{verd- + ter}).} m. worker.
	

	
		\key{verteti.} {\footnotesize (From PIE \ipa{*wert-}).} v. to turn, to spin.
	

	
		\key{verus.} {\footnotesize (From PIE \ipa{*h₁wérus}).} adj. broad.
	

	
		\key{Vestina Germania.}  f. West Germany.
	

	
		\key{vestinos.} {\footnotesize (From Dan. \ipa{vestom}).} adj. western, west, westerly.
	

	\key{vestis,}
		\textit{usties.} {\footnotesize (From PIE \ipa{*wéstis}).} f.
			\textbf{1.} the act of dressing.
			\textbf{2.} clothing.
	

	
		\key{vestom.} {\footnotesize (From PG \ipa{*westrą}).} n. west.
	

	
		\key{vesus.} {\footnotesize (From PIE \ipa{*h₁wésus}).} adj. good, excellent.
	

	
		\key{vesverdíeti.}  v. to bless.
	

	
		\key{veterános.} {\footnotesize (From Lat. \ipa{veterānus}).} m. veteran.
	

	
		\key{vid.} {\footnotesize (From Dan. \ipa{voide}).} interj. lo, behold, look.
	

	
		\key{videti.} {\footnotesize (From PIE \ipa{*weyd-e-ti}).} v. to see.
	

	
		\key{videva.} {\footnotesize (From PIE \ipa{*h₁widʰéwh₂}).} f. widow.
	

	
		\key{vidia.} {\footnotesize (From PIE \ipa{*weyd-}).} f. knowledge, wisdom.
	

	
		\key{vidsant.} {\footnotesize (From PIE \ipa{*wídḱm̥ti}).} num. twenty.
	

	
		\key{vidsantos.}  adj. twentieth.
	

	
		\key{Vienna.} {\footnotesize (From It. \ipa{Vienna}).} f. Vienna.
	

	
		\key{Vilmos.} {\footnotesize (Ultimately from PG \ipa{*Wiljahelmaz}).} m. William, Wilhelm, Guillermo, Guillaume.
	

	
		\key{Vilmos Condrados Rointgen.}  m. Wilhelm Conrad Röntgen.
	

	
		\key{Vilnius.} {\footnotesize (From Lith. \ipa{Vilnius}).} m. Vilnius, capital of Lithuania.
	

	
		\key{viman.} {\footnotesize (From PIE \ipa{*wéh₁imn̥}).} n. that which is twisted, plaiting.
	

	
		\key{vinó.} {\footnotesize (From PIE \ipa{*wí-noh₁}).} prep. (+gen) except.
	

	
		\key{virikom.} {\footnotesize (From Dan. \ipa{viros}).} n. little man.
	

	
		\key{viros.} {\footnotesize (From PIE \ipa{*wiHrós}).} m.
			\textbf{1.} man.
			\textbf{2.} husband.
			\textbf{3.} warrior, hero.
	

	
		\key{visodi.} {\footnotesize (From Dan. \ipa{viso- + -di}. The suffix \ipa{-di} comes from PIE \ipa{*dʰi}).} proadv. (universal) everywhere.
	

	
		\key{visoimos.}  proadv. (universal) to everywhere.
	

	
		\key{visoives.} {\footnotesize (From Dan. \ipa{viso- + -ives}. The suffix \ipa{-ives} comes from \ipa{juves}, the genitive of \ipa{oju}).} proadv. (universal) at every time, always, forever.
	

	
		\key{visosmad.}  proadv. (universal) from everywhere.
	

	
		\key{visosme.}  proadv. (universal) in every way, in every manner.
	

	
		\key{visoti.}  proadv. (universal) by every amount.
	

	
		\key{vispotis.} {\footnotesize (From PIE \ipa{*wiḱpótis}).} m.
			\textbf{1.} clan chief.
			\textbf{2.} lord.
	

	
		\key{vistis.} {\footnotesize (From PIE \ipa{*weyd-}).} f. news.
	

	
		\key{vistor.} {\footnotesize (From PIE \ipa{*wéydtōr}).} m.
			\textbf{1.} one who has seen, one who knows.
			\textbf{2.} witness.
	

	
		\key{vistos.} {\footnotesize (From PIE \ipa{*widtós}).} adj. seen, known.
	

	
		\key{viteros.}  adj. second.
	

	
		\key{vitis.} {\footnotesize (From PIE \ipa{*wéh₁itis}).} f.
			\textbf{1.} that which twines or bends.
			\textbf{2.} branch, switch.
	

	\key{vodan,}
		\textit{vedens.} {\footnotesize (From PIE \ipa{*wódr̥}).} n. water.
	

	
		\key{voide.} {\footnotesize (From PIE \ipa{*weyd-}).} v. to know.
	

	
		\key{vojasj.} {\footnotesize (From Fr. \ipa{voyage}).} m. trip, voyage.
	

	
		\key{vojasjeti.} {\footnotesize (From Dan. \ipa{vojasj}).} v. to travel, to take a trip, to go on a trip.
	

	
		\key{volhoviskos.}  adj. pertaining to the Volkhov region in Russia.
	

	
		\key{vops.} {\footnotesize (From PIE \ipa{*wṓkʷs}).} f. voice, speech.
	

	\key{vor,}
		\textit{fures, varmos.} {\footnotesize (From PIE \ipa{*dʰwṓr}).} m. door.
	

	\key{vosan,}
		\textit{vesens.} {\footnotesize (From PIE \ipa{*wósr̥}).} n. spring.
	

	
		\key{vostu.} {\footnotesize (From PIE \ipa{*h₂wóstu}).} n. home, dwelling.
	

	
		\key{wagón.} {\footnotesize (From Fr. \ipa{wagon}, from En. \ipa{waggon}, from Dutch \ipa{wagen}).} m. railway carriage.
	

	
		\key{Werner.} {\footnotesize (From Ger. \ipa{Werner}).} m. Werner, Warner.
	

	
		\key{Werner Haisemberg.}  m. Werner Heisenberg.
	

	
		\key{wolframiom.}  n.{\footnotesize \textit{ chemistry.}} wolfram, wolframium, tungsten.
	

	
		\key{Xavier.} {\footnotesize (From Basque \ipa{etxeberria}).} m. Javier, Xavier, Saverio.
	

	
		\key{Xavier Castro.}  m. Javier Castro, creator of the Danetian language.
	

	
		\key{xenón.}  m.{\footnotesize \textit{ chemistry.}} xenon.
	

	
		\key{York.} {\footnotesize (From En. \ipa{York}).} f. York.
	

	
		\key{ypsilón.}  f. Greek letter upsilon.
	

	
		\key{ytterbiom.}  n.{\footnotesize \textit{ chemistry.}} ytterbium.
	

	
		\key{yttriom.}  n.{\footnotesize \textit{ chemistry.}} yttrium.
	

	
		\key{Zagrabia.} {\footnotesize (From Lat. \ipa{Zagrabia}).} f. Zagreb, capital of Croatia.
	

	
		\key{ze.}  n. letter \textit{z}.
	

	
		\key{zeta.}  f. Greek letter zeta.
	

	
		\key{zincom.}  n.{\footnotesize \textit{ chemistry.}} zinc.
	

	
		\key{zirconiom.}  n.{\footnotesize \textit{ chemistry.}} zirconium.

